### 2025-02-18 Draft1.qmd
Add initial report draft qmd

---
title: "Draft 1"
author: "Kevin, Marco, Bianca"
date: ''
format:
  pdf: default
  html: default
editor: visual
fontsize: 10pt
geometry: margin=1in
toc: true                   # add table of contents at the beginning
toc-depth: 2                # Only titles that start with # or ##
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

### 2025-02-19 Draft1.qmd
Added 2 project ideas + bibliography

@@
 ---
-title: "Draft 1"
-author: "Kevin, Marco, Bianca"
-date: ''
+title: "Homicide Increases in the COVID-19 Era"
+
+author: "Marco Stine, Kevin Hu, and Bianca Schutz"
+
 format:
+
  pdf: default
-  html: default
-editor: visual
+
 fontsize: 10pt
+
 geometry: margin=1in
-toc: true                   # add table of contents at the beginning
-toc-depth: 2                # Only titles that start with # or ##
+bibliography: references.bib
 ---
 
-## Quarto
+# Introduction - Idea #1
 
-Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.
+During the COVID-19 pandemic in 2021, homicide rates in Los Angeles increased substantially.
+A report by the Legal Defense Fund's Thurgood Marshall Institute found that this increase was associated with both pre-pandemic and pandemic-induced economic instability and inequalities.[@moore2022]
 
-## Running Code
+Using National Incident-Based Reporting System (NIBRS) data, this report will examine the impact of the pandemic and various socioeconomic factors on homicide rates in LA.
+We aim to compare the effect of these local characteristics on the incidence of homicide in pre-, mid-, and post-pandemic conditions.
 
-When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:
+We will utilize NIBRS data from 2018-2024 to examine incident-level data, allowing us to fully capture the circumstances of each event.
+Our data is obtained from the Los Angeles Police Department (LAPD), who regularly updates Crime Data on the Data.gov website.
+The data includes all crime data, which we will subset to focus on homicide, and details such as date, location, and offense type.
+
+# Loading in packages and dataset
 
 ```{r}
-1 + 1
+# Primary datasets: Crime Data 2020-Present and 2010-2019
+crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")
+
+crime_la_2020_present <- read.csv("Raw_Data/Crime_Data_from_2020_to_Present.csv")
+
+# Secondary datasets
+for (i in 2018:2023) {
+  obj_name <- paste0("california_median_income_", i)
+  assign(obj_name, read.csv(paste0("Raw_Data/ACSST5Y",i,".S1903-Data.csv")))
+}
 ```
 
-You can add options to executable code like this
+# Introduction - Idea #2
 
-```{r}
-#| echo: false
-2 * 2
-```
+During the COVID-19 pandemic in 2021, homicide rates in Los Angeles and Houston increased substantially, while rates slightly increased in New York and Chicago and decreased in Dallas and Seattle.
+A report by the Legal Defense Fund's Thurgood Marshall Institute found that this increase was associated with both pre-pandemic and pandemic-induced economic instability and inequalities.[@moore2022]
 
-The `echo: false` option disables the printing of code (only output is displayed).
+Using National Incident-Based Reporting System (NIBRS) data, this report will examine the impact of the pandemic and various socioeconomic factors on homicide rates in these six cities.
+We aim to compare the effect of these local characteristics on the incidence of homicide in pre-, mid-, and post-pandemic conditions.
+
+We will utilize NIBRS data from 2018-2024 to examine incident-level data, allowing us to fully capture the circumstances of each event.
+Our data is obtained from the FBI Crime Data Explorer Tool, who publishes NIBRS data by state and year.
+The data includes all crime data, which we will subset to focus on homicide and the incidents occurring in the relevant agencies (city police departments such as LAPD).

### 2025-02-20 Draft1.qmd
added my stuff

@@
 The data includes all crime data, which we will subset to focus on homicide and the incidents occurring in the relevant agencies (city police departments such as LAPD).
+
+# Plot 1 - Examining Crime Density from 2010–2019 and 2020-Present
+
+```{r}
+
+par(mfrow = c(1, 2), mar = c(4, 4, 2, 1))
+
+smoothScatter(
+  x    = crime_la_2010_2019$LON,
+  y    = crime_la_2010_2019$LAT,
+  xlim = c(-119, -118),
+  ylim = c(33.5, 34.5),
+  nbin = 300,
+  xlab = "Longitude",
+  ylab = "Latitude",
+  main = "Crime Density (2010–2019)",
+  colramp = colorRampPalette(c("white", "skyblue", "blue", "darkblue"))
+)
+
+smoothScatter(
+  x    = crime_la_2020_present$LON,
+  y    = crime_la_2020_present$LAT,
+  xlim = c(-119, -118),
+  ylim = c(33.5, 34.5),
+  nbin = 300,
+  xlab = "Longitude",
+  ylab = "Latitude",
+  main = "Crime Density (2020–Present)",
+  colramp = colorRampPalette(c("white", "pink", "red", "darkred"))
+)
+
+par(mfrow = c(1, 1))
+
+```
+Both time periods show major clusters of reported crimes in central LA. The 2020–present data appears slightly more dispersed to the north, but overall patterns remain similar looking.
+
+# Plot 2 - Examining Number Of Crimes Over Time
+
+```{r}
+
+crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
+
+crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
+                                  format = "%m/%d/%Y %I:%M:%S %p", 
+                                  tz = "America/Los_Angeles")
+
+crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
+
+weekly_counts <- aggregate(DR_NO ~ YearWeek, data = crime_la, FUN = length)
+names(weekly_counts)[2] <- "Crimes"
+
+weekly_counts <- weekly_counts[order(weekly_counts$YearWeek), ]
+
+weekly_counts$Index <- seq_len(nrow(weekly_counts))
+
+plot(
+  weekly_counts$Index,
+  weekly_counts$Crimes,
+  type = "l",
+  col  = "blue",
+  lwd  = 2,
+  xlab = "Weekly Index",
+  ylab = "Number of Crimes",
+  main = "Weekly Crime Counts in Los Angeles (2010–Present)"
+)
+
+abline(h = pretty(weekly_counts$Crimes), v = pretty(weekly_counts$Index), 
+       col = "gray", lty = "dotted")
+
+```
+The time series shows stable weekly crime counts with occasional large spikes.
+These surges indicate certain weeks where crime reporting jumped significantly, there is a recent interesting surge that seems to mean-revert in 2024.
+
+# Using Regression to Predict ____
+
+```{r}
+
+weekly_counts$CrimesLag <- c(NA, weekly_counts$Crimes[-nrow(weekly_counts)])
+
+ar1_trend_regression <- lm(Crimes ~ Index + CrimesLag, data = weekly_counts, na.action = na.omit)
+
+summary(ar1_trend_regression)
+
+```
+
+# Interpretation
+
+The lagged crime coefficient (0.81949) suggests that a high crime week is often followed by another high crime week.
+Since the trend term is not significant, there is no clear drift in weekly crime over time.

### 2025-02-20 Draft1.qmd
Update Draft1.qmd

@@
 
-```
 The time series shows stable weekly crime counts with occasional large spikes.
@@
 
+# Plot 3 Crime rate frequency by day of the week
+
+```{r}
+# Convert Date.Rptd to Date format (ignore time part)
+crime_la$Date.Rptd
+crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")
+
+# Set locale to English (for weekday names)
+Sys.setlocale("LC_TIME", "C")
+
+# Extract the day of the week
+crime_la$DayOfWeek <- weekdays(crime_la$Date)
+crime_la$DayOfWeek
+
+# Count occurrences of crimes per day
+crime_counts <- table(crime_la$DayOfWeek)
+crime_counts
+
+# Order the days correctly
+day_order <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
+crime_counts <- crime_counts[day_order]
+
+# Plot bar chart with horizontal labels
+barplot(crime_counts, col = "lightblue",
+        main = "Crime Occurrences by Day of the Week",
+        xlab = "Day of the Week", ylab = "Number of Crimes",
+        names.arg = names(crime_counts), las = 1, cex.names = 0.8, cex.axis = 0.6)
+```
+The number of crimes are rather evenly distributed, with the weekends having slightly lower crime compared to week days.
+
+# Plot 4 Gender Distribution for crimes over time
+
+```{r}
+
+# Combine the datasets
+crime_la <- rbind(crime_rate_ds, crime_old_ds)
+
+# Remove missing values and empty strings in Vict.Sex column
+crime_la <- subset(crime_la, Vict.Sex != "" & !is.na(Vict.Sex))
+
+# Replace all non-"F" and non-"M" values with "N"
+crime_la$Vict.Sex <- ifelse(crime_la$Vict.Sex %in% c("F", "M"), 
+                                  crime_la$Vict.Sex, "N")
+# Count occurrences of each gender
+gender_counts <- table(crime_la$Vict.Sex)
+
+# Create a pie chart using base R
+color <- c("F" = "pink", "M" = "lightblue", "N" = "purple")
+pie(
+  gender_counts, 
+  labels = paste(names(gender_counts), 
+                 round(gender_counts / sum(gender_counts) * 100, 1), "%"), 
+  col = color[names(gender_counts)],
+  main = "Distribution of Crime Victims by Gender"
+)
+```
+
+From the pie chart, we can see that around 50% of the crime is commited by male, 45% by female and 5% by non-binary.
+
 # Interpretation

### 2025-02-20 Draft1.qmd
Add new project member name
Updated introduction section

@@
 
-author: "Marco Stine, Kevin Hu, and Bianca Schutz"
+author: "Marco Stine, Kevin Hu, Bianca Schutz, and Kaitlyn Kirt"
 
@@
 
-# Introduction - Idea #1
+# Introduction
 
@@
 ```
-
-# Introduction - Idea #2
-
-During the COVID-19 pandemic in 2021, homicide rates in Los Angeles and Houston increased substantially, while rates slightly increased in New York and Chicago and decreased in Dallas and Seattle.
-A report by the Legal Defense Fund's Thurgood Marshall Institute found that this increase was associated with both pre-pandemic and pandemic-induced economic instability and inequalities.[@moore2022]
-
-Using National Incident-Based Reporting System (NIBRS) data, this report will examine the impact of the pandemic and various socioeconomic factors on homicide rates in these six cities.
-We aim to compare the effect of these local characteristics on the incidence of homicide in pre-, mid-, and post-pandemic conditions.
-
-We will utilize NIBRS data from 2018-2024 to examine incident-level data, allowing us to fully capture the circumstances of each event.
-Our data is obtained from the FBI Crime Data Explorer Tool, who publishes NIBRS data by state and year.
-The data includes all crime data, which we will subset to focus on homicide and the incidents occurring in the relevant agencies (city police departments such as LAPD).
 
@@
 
+# Plot 5 - Median Household Income by Zip Code in Los Angeles by Year
+
 # Interpretation

### 2025-02-20 Draft1.qmd
Added Plot 5 first draft

@@
 
+# Combine the two datasets
+crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
+
 # Secondary datasets
-for (i in 2018:2023) {
-  obj_name <- paste0("california_median_income_", i)
-  assign(obj_name, read.csv(paste0("Raw_Data/ACSST5Y",i,".S1903-Data.csv")))
+california_median_income_2018_2023 <- list()
+yrs <- 2018:2023
+for (i in 1:length(yrs)) {
+  obj <- read.csv(paste0("Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
+  obj <- obj[-1,]
+  obj$year <- yrs[[i]]
+  california_median_income_2018_2023[[i]] <- obj
 }
+
+cali_median_income <- do.call(rbind, california_median_income_2018_2023)
+
+# adding a variable that's just the ZCTA zip code
+cali_median_income$ZCTA <- stringr::str_sub(cali_median_income$NAME, -5)
 ```
@@
 ```{r}
-
-crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
-
 crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
@@
 
+```{r}
+library(sf)
+library(tidygeocoder)
+zcta <- read_sf(dsn = "cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
+LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs
+
+colnames(LA_zctas) <- c("X", "ZCTA5CE20")
+LA_zctas$ZCTA5CE20 <- as.character(LA_zctas$ZCTA5CE20)
+
+zcta_filtered <- zcta %>% inner_join(LA_zctas, by = "ZCTA5CE20")
+
+median_household_income <- cali_median_income %>% select(ZCTA, S1903_C03_001E, year)
+
+median_household_income$S1903_C03_001E <- as.integer(median_household_income$S1903_C03_001E)
+
+zcta_with_data <- zcta_filtered %>% inner_join(median_household_income, join_by("ZCTA5CE20" == "ZCTA"))
+
+plot(st_geometry(zcta_with_data), 
+     col = brewer.pal(9, "Blues")[as.numeric(cut(zcta_with_data$S1903_C03_001E, 
+                                                 breaks = 9))], 
+     border = 'grey', 
+     axes = TRUE, 
+     main = "ZCTA Data Colored by Numerical Column", 
+     cex.main = 1.5)
+```
+
 # Interpretation

### 2025-02-20 Draft1.qmd
fixed incomplete code block

@@
        col = "gray", lty = "dotted")
+```
 
@@
      axes = TRUE, 
-     main = "ZCTA Data Colored by Numerical Column", 
+     main = "Median Household Income by ", 
      cex.main = 1.5)

### 2025-02-20 Draft1.qmd
fixed error - crime_la is already combined in initial block now

@@
 ```{r}
-
-# Combine the datasets
-crime_la <- rbind(crime_rate_ds, crime_old_ds)
-
 # Remove missing values and empty strings in Vict.Sex column

### 2025-02-20 Draft1.qmd
Fixed destination

@@
 library(tidygeocoder)
-zcta <- read_sf(dsn = "cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
+zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
 LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs

### 2025-02-20 Draft1.qmd
Added viz

--- 2025-02-20 Draft1.qmd
@@
 library(tidygeocoder)
-zcta <- read_sf(dsn = "cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
+library(tidyverse)
+library(RColorBrewer)
+
+zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
 LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs
@@
 
+zcta_no_NAs <- zcta_with_data %>% filter(!is.na(S1903_C03_001E))
+
 plot(st_geometry(zcta_with_data), 
@@
      axes = TRUE, 
-     main = "Median Household Income by ", 
+     main = "Median Household Income by Year and Zip Code", 
      cex.main = 1.5)
 ```
+
+# Plot 6 - Frequency of Top 10 Crimes
 

### 2025-02-21 Draft1.qmd
Added references and began Plot 6
Added maps

@@
 The data includes all crime data, which we will subset to focus on homicide, and details such as date, location, and offense type.
+
+The codebook for this dataset can be found here: https://data.lacity.org/Public-Safety/Crime-Data-from-2020-to-Present/2nrs-mtv8/about_data.
 
@@
 ```
+# Interpretation
+
+The lagged crime coefficient (0.81949) suggests that a high crime week is often followed by another high crime week.
+Since the trend term is not significant, there is no clear drift in weekly crime over time.
 
@@
 
+# joining the zcta shape file with LA zcta reference file
 zcta_filtered <- zcta %>% inner_join(LA_zctas, by = "ZCTA5CE20")
 
+# getting LA median income using joined file
 median_household_income <- cali_median_income %>% select(ZCTA, S1903_C03_001E, year)
@@
 
-zcta_no_NAs <- zcta_with_data %>% filter(!is.na(S1903_C03_001E))
+# remove NAs
+zcta_no_NAs <- zcta_with_data %>% filter(!is.na(S1903_C03_001E), S1903_C03_001E != "-")
 
-plot(st_geometry(zcta_with_data), 
-     col = brewer.pal(9, "Blues")[as.numeric(cut(zcta_with_data$S1903_C03_001E, 
-                                                 breaks = 9))], 
+palette <- brewer.pal(7, "Greens")
+
+par(mfrow = c(2, 3), mar = c(2, 2, 2, 6), oma = c(0, 0, 4, 0))
+
+yrs <- 2018:2023
+
+for (yr in yrs) {
+  # Filter by year
+  subset_data <- zcta_no_NAs[zcta_no_NAs$year == yr, ]
+
+  # Divide it into bins
+  bins <- cut(subset_data$S1903_C03_001E, breaks = 7, labels = FALSE)
+
+  # Plot
+  plot(st_geometry(subset_data),
+       col = palette[bins],
        border = 'grey',
-     axes = TRUE, 
-     main = "Median Household Income by Year and Zip Code", 
+       axes = FALSE,
+       main = yr,
        cex.main = 1.5)
+
+  # Add legend
+  legend("right", inset = c(-0.5, 0), legend = round(quantile(subset_data$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0),
+         fill = palette, title = "Median Income ($)", bty = "n", cex = 0.8, xpd = TRUE)
+}
+mtext("Median Household Income by Year and Zip Code in LA", outer = TRUE, cex = 1, font = 2)
 ```
@@
 
-# Interpretation
-
-The lagged crime coefficient (0.81949) suggests that a high crime week is often followed by another high crime week.
-Since the trend term is not significant, there is no clear drift in weekly crime over time.
+```{r}
+crime_frequency <- data.frame(table(crime_la$Crm.Cd.Desc))
+top_10_crimes <- crime_frequency[order(-crime_frequency$Freq), ][1:10, ]
+```

### 2025-02-21 Draft1.qmd
Draft1.qmd

@@
 ```
+# Interpretation
+
 The number of crimes are rather evenly distributed, with the weekends having slightly lower crime compared to week days.
@@
                                   crime_la$Vict.Sex, "N")
+
 # Count occurrences of each gender
@@
 ```
+# Interpretation
 

### 2025-02-22 Draft1.qmd
Added TOC

@@
 fontsize: 10pt
-
+toc: true                   # add table of contents at the beginning
+toc-depth: 3               # Only titles that start with # or ##
 geometry: margin=1in
@@
 
-```{r}
+```{r, eval=FALSE}
+library(sf)
+library(tidygeocoder)
+library(tidyverse) # using only for joining datasets to build map visual
+library(RColorBrewer)
+
 # Primary datasets: Crime Data 2020-Present and 2010-2019
@@
 
-# Plot 1 - Examining Crime Density from 2010–2019 and 2020-Present
+# Plots
+
+## Examining Crime Density from 2010–2019 and 2020-Present
 
@@
 ```
+
 Both time periods show major clusters of reported crimes in central LA. The 2020–present data appears slightly more dispersed to the north, but overall patterns remain similar looking.
 
-# Plot 2 - Examining Number Of Crimes Over Time
+## Examining Number Of Crimes Over Time
 
@@
 
-# Using Regression to Predict ____
-
-```{r}
-
-weekly_counts$CrimesLag <- c(NA, weekly_counts$Crimes[-nrow(weekly_counts)])
-
-ar1_trend_regression <- lm(Crimes ~ Index + CrimesLag, data = weekly_counts, na.action = na.omit)
-
-summary(ar1_trend_regression)
-
-```
-# Interpretation
-
-The lagged crime coefficient (0.81949) suggests that a high crime week is often followed by another high crime week.
-Since the trend term is not significant, there is no clear drift in weekly crime over time.
-
-# Plot 3 Crime rate frequency by day of the week
+## Crime rate frequency by day of the week
 
@@
 ```
-# Interpretation
+
+### Interpretation
 
@@
 
-# Plot 4 Gender Distribution for crimes over time
+## Gender Distribution for crimes over time
 
@@
 
-# Plot 5 - Median Household Income by Zip Code in Los Angeles by Year
+## Median Household Income by Zip Code in Los Angeles by Year
 
 ```{r}
-library(sf)
-library(tidygeocoder)
-library(tidyverse)
-library(RColorBrewer)
-
 zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
@@
 
-median_household_income$S1903_C03_001E <- as.integer(median_household_income$S1903_C03_001E)
-
 zcta_with_data <- zcta_filtered %>% inner_join(median_household_income, join_by("ZCTA5CE20" == "ZCTA"))
@@
 
+zcta_no_NAs$S1903_C03_001E <- as.integer(zcta_no_NAs$S1903_C03_001E)
+
 palette <- brewer.pal(7, "Greens")
 
-par(mfrow = c(2, 3), mar = c(2, 2, 2, 6), oma = c(0, 0, 4, 0))
+# setting a layout for a 2x3 grid, margins adjusted to fit the plots, oma() for outer margin area.
+par(mfrow = c(2, 3), mar = c(1, 1, 1, 6), oma = c(0, 0, 3, 0), cex.main = 1.5)
 
@@
        border = 'grey',
-       axes = FALSE,
-       main = yr,
-       cex.main = 1.5)
+       axes = FALSE)
+  title(yr, line = -1)
 
@@
 }
-mtext("Median Household Income by Year and Zip Code in LA", outer = TRUE, cex = 1, font = 2)
+mtext("Median Household Income by Year and Zip Code in LA", outer = TRUE, cex = 1.5, font = 2)
 ```
 
-# Plot 6 - Frequency of Top 10 Crimes
+## Frequency of Top 10 Crimes
 
@@
 ```
+
+## Using Regression to Examine Weekly Crime Rates
+
+```{r}
+
+weekly_counts$CrimesLag <- c(NA, weekly_counts$Crimes[-nrow(weekly_counts)])
+
+ar1_trend_regression <- lm(Crimes ~ Index + CrimesLag, data = weekly_counts, na.action = na.omit)
+
+summary(ar1_trend_regression)
+
+```
+
+### Interpretation
+
+The lagged crime coefficient (0.81949) suggests that a high crime week is often followed by another high crime week.
+Since the trend term is not significant, there is no clear drift in weekly crime over time.

### 2025-02-22 Draft1.qmd
Fixed error
Updated notebook

@@
 author: "Marco Stine, Kevin Hu, Bianca Schutz, and Kaitlyn Kirt"
-
-format:
-
- pdf: default
-
 fontsize: 10pt
@@
 bibliography: references.bib
+format:
+  pdf: default
 ---
@@
 
-```{r, eval=FALSE}
+```{r, output=FALSE}
 library(sf)
@@
   main = "Crime Density (2010–2019)",
-  colramp = colorRampPalette(c("white", "skyblue", "blue", "darkblue"))
+  colramp = colorRampPalette(c("white", "skyblue", "blue", "darkblue")),
+  cex.main=0.8
 )
@@
   main = "Crime Density (2020–Present)",
-  colramp = colorRampPalette(c("white", "pink", "red", "darkred"))
+  colramp = colorRampPalette(c("white", "pink", "red", "darkred")),
+  cex.main=0.8
 )
@@
 # Convert Date.Rptd to Date format (ignore time part)
-crime_la$Date.Rptd
+#crime_la$Date.Rptd
 crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")
@@
 crime_la$DayOfWeek <- weekdays(crime_la$Date)
-crime_la$DayOfWeek
+# crime_la$DayOfWeek
 
@@
 crime_counts <- table(crime_la$DayOfWeek)
-crime_counts
+# crime_counts
 
@@
 zcta_no_NAs$S1903_C03_001E <- as.integer(zcta_no_NAs$S1903_C03_001E)
+```
 
+Now, using this joined dataset, create six maps, one for each year.
+
+```{r}
+# Calculate quantiles for the entire dataset (from 2018 to 2023)
+quantiles <- round(quantile(zcta_no_NAs$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0)
+
+# Setting a layout for a 2x3 grid
+par(mfrow = c(2, 3), mar = c(1, 1, 1, 5), oma = c(0, 0, 3, 0), cex.main = 1.5)
+
+# Define color palette
 palette <- brewer.pal(7, "Greens")
 
-# setting a layout for a 2x3 grid, margins adjusted to fit the plots, oma() for outer margin area.
-par(mfrow = c(2, 3), mar = c(1, 1, 1, 6), oma = c(0, 0, 3, 0), cex.main = 1.5)
-
+# Plot for each year
 yrs <- 2018:2023
-
 for (yr in yrs) {
@@
 
-  # Divide it into bins
-  bins <- cut(subset_data$S1903_C03_001E, breaks = 7, labels = FALSE)
+  # Divide into consistent bins based on the global quantiles
+  bins <- cut(subset_data$S1903_C03_001E, breaks = quantiles, labels = FALSE, include.lowest = TRUE)
 
@@
        border = 'grey',
-       axes = FALSE)
+       axes = FALSE,
+       lwd = 0.5)
   title(yr, line = -1)
 
-  # Add legend
-  legend("right", inset = c(-0.5, 0), legend = round(quantile(subset_data$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0),
-         fill = palette, title = "Median Income ($)", bty = "n", cex = 0.8, xpd = TRUE)
+  # Add legend (use the global quantiles)
+  legend("right", inset = c(-.75, 0), legend = quantiles, fill = palette, title = "Median Income ($)", bty = "n", cex = 0.8, xpd = TRUE)
 }
-mtext("Median Household Income by Year and Zip Code in LA", outer = TRUE, cex = 1.5, font = 2)
+
+# Outer title for the whole set of plots
+mtext("Median Household Income by Year and Zip Code in LA", outer = TRUE, cex = 1, font = 2)
+
 ```
+
+### Interpretation
+The maps demonstrate that median income in certain LA zip codes (particularly those to the west) have a much higher median household income than the rest of the city, but over the last six years, the median income in most other LA zip codes has increased.
+However, there are some that a typical family is at or below the poverty line.
+We intend to investigate if there is a spatial correlation between these poorer zip codes and the incidence of crimes, particularly homicide, and whether this is affected by pandemic conditions.
 
@@
 ```
+
+## Homicide
 

### 2025-02-22 Draft1.qmd
added subtitle for bib

@@
 ```{r, output=FALSE}
-library(sf)
-library(tidygeocoder)
+library(sf) # for map
 library(tidyverse) # using only for joining datasets to build map visual
-library(RColorBrewer)
+library(RColorBrewer) # for map color scheme
+library(tidygeocoder) # to get zip codes of addresses
 
@@
 ```{r}
-# Calculate quantiles for the entire dataset (from 2018 to 2023)
-quantiles <- round(quantile(zcta_no_NAs$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0)
+# quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
+qntls <- round(quantile(zcta_no_NAs$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0)
 
-# Setting a layout for a 2x3 grid
+# 2x3 plot layout for years 2018-2023, setting margins and outer margins, main title sizes
 par(mfrow = c(2, 3), mar = c(1, 1, 1, 5), oma = c(0, 0, 3, 0), cex.main = 1.5)
 
-# Define color palette
+# color palette for sequential data
 palette <- brewer.pal(7, "Greens")
 
-# Plot for each year
 yrs <- 2018:2023
+
 for (yr in yrs) {
-  # Filter by year
+  # get just that year's data
   subset_data <- zcta_no_NAs[zcta_no_NAs$year == yr, ]
 
-  # Divide into consistent bins based on the global quantiles
-  bins <- cut(subset_data$S1903_C03_001E, breaks = quantiles, labels = FALSE, include.lowest = TRUE)
+  # creating bins based on the qtls
+  bins <- cut(subset_data$S1903_C03_001E, breaks = qntls, labels = FALSE, include.lowest = TRUE)
 
@@
 
-  # Add legend (use the global quantiles)
-  legend("right", inset = c(-.75, 0), legend = quantiles, fill = palette, title = "Median Income ($)", bty = "n", cex = 0.8, xpd = TRUE)
+  # adding legend
+  legend("right", inset = c(-.75, 0), legend = qntls, fill = palette, title = "Median Income ($)", bty = "n", cex = 0.8, xpd = TRUE)
 }
 
-# Outer title for the whole set of plots
+# main title for all plots
 mtext("Median Household Income by Year and Zip Code in LA", outer = TRUE, cex = 1, font = 2)
@@
 
-## Frequency of Top 10 Crimes
+## Crime Rates Compared to Median Household Income
 
 ```{r}
-crime_frequency <- data.frame(table(crime_la$Crm.Cd.Desc))
-top_10_crimes <- crime_frequency[order(-crime_frequency$Freq), ][1:10, ]
+# using latitude and longitude to obtain the zip codes of the crime locations
+# crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
+# id_lat_long_crime_la <- subset(crime_la, select = c("LAT", "LON", "DR_NO", "year"))
+
+# taking a random sample to reduce load on model
+#set.seed(123)
+#sampled_ids <- id_lat_long_crime_la[sample(1:nrow(id_lat_long_crime_la), 500),]
+#rownames(sampled_ids) <- NULL
+#zip_codes <- reverse_geocode(sampled_ids, lat = LAT, long = LON, verbose = TRUE)
+#write.csv(zip_codes, "sampled_crimes.csv")
+zip_codes <- read.csv("sampled_crimes.csv")
+
+zip_codes$ZCTA5CE20 <- stringr::str_extract(zip_codes$address, "\\d{5}")
+
+crime_income_sample <- zip_codes %>% inner_join(unique(subset(zcta_no_NAs, select = c("ZCTA5CE20", "S1903_C03_001E", "year"))), by = c("ZCTA5CE20", "year"))
+
+# this takes a while, so loading in this csv instead
+
+crime_income_sample <- zip_codes %>% inner_join(unique(subset(zcta_no_NAs, select = c("ZCTA5CE20", "S1903_C03_001E", "year"))), by = c("ZCTA5CE20", "year"))
+
+crime_income_per_zip <- crime_income_sample %>% 
+  group_by(ZCTA5CE20) %>% 
+  summarize(number_of_crimes = n(),
+            cali_median_income = mean(S1903_C03_001E)) %>%
+  ungroup()
+
+plot(x = crime_income_per_zip$cali_median_income, 
+     y = crime_income_per_zip$number_of_crimes, 
+     type = "p",
+     ylab = "Number of Crimes",  
+     xlab = "Median Income ($)",  
+     main = "Crime vs. Median Income by ZIP Code",
+     bty = "l")
 ```
 
-## Homicide
+### Based on this plot, we do not observe a clear trend between the number of crimes committed and the median income in that zip code.
+However, this scatterplot only accounts for a small portion of the dataset and averages across pre- and post-pandemic conditions.
+Future visualizations and analyses could consider pandemic vs non-pandemic years as variables.
+
+# Regression Analyses
 
@@
 ```{r}
-
 weekly_counts$CrimesLag <- c(NA, weekly_counts$Crimes[-nrow(weekly_counts)])
@@
 summary(ar1_trend_regression)
-
 ```
-
 ### Interpretation
@@
 Since the trend term is not significant, there is no clear drift in weekly crime over time.
+
+## Basic Regression Between Median Income and Crimes per Zip Code
+
+```{r}
+lm_income_crimes <- lm(number_of_crimes ~ cali_median_income, data = crime_income_per_zip)
+
+summary(lm_income_crimes)
+```
+### Interpretation
+
+Based on the p-value, we observe that median income is not significant at a 0.05 level, meaning it does not significantly decrease the amount of crimes committed.
+The r^2 value additionally tells us that the income only explains a small portion of the variation in the number of crimes.
+As stated with the scatterplot, a larger sample and considering other factors might yield different results.
+
+# Bibliography

### 2025-02-24 Draft1.qmd
Updated draft based on report feedback about moving code blocks to the end

@@
   pdf: default
+number-sections: true
 ---
@@
 
-# Loading in packages and dataset
-
-```{r, output=FALSE}
+```{r, output=FALSE, echo=FALSE}
 library(sf) # for map
@@
 
-```{r}
+```{r, echo=FALSE}
 
@@
 
-```{r}
+```{r, echo=FALSE}
 crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
@@
 
-```{r}
+```{r, echo=FALSE, fig.cap = "Crime Occurrences by Day of the Week"}
 # Convert Date.Rptd to Date format (ignore time part)
@@
 barplot(crime_counts, col = "lightblue",
-        main = "Crime Occurrences by Day of the Week",
         xlab = "Day of the Week", ylab = "Number of Crimes",
@@
 
-```{r}
+```{r, echo=FALSE}
 # Remove missing values and empty strings in Vict.Sex column
@@
 
-```{r}
+```{r, echo=FALSE}
 zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
@@
 
-```{r}
+```{r, echo=FALSE}
 # quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
@@
 ### Interpretation
+
 The maps demonstrate that median income in certain LA zip codes (particularly those to the west) have a much higher median household income than the rest of the city, but over the last six years, the median income in most other LA zip codes has increased.
@@
 
-```{r}
+```{r, echo=FALSE}
 # using latitude and longitude to obtain the zip codes of the crime locations
@@
 
-### Based on this plot, we do not observe a clear trend between the number of crimes committed and the median income in that zip code.
+### Interpretation
+
+Based on this plot, we do not observe a clear trend between the number of crimes committed and the median income in that zip code.
 However, this scatterplot only accounts for a small portion of the dataset and averages across pre- and post-pandemic conditions.
@@
 
-## Using Regression to Examine Weekly Crime Rates
+## Regression to Examine Weekly Crime Rates
 
-```{r}
+```{r, echo=FALSE}
+library(broom)
+
 weekly_counts$CrimesLag <- c(NA, weekly_counts$Crimes[-nrow(weekly_counts)])
@@
 
-summary(ar1_trend_regression)
+ar1_trend_regression %>%
+tidy() %>%
+  mutate(
+    p.value = scales::pvalue(p.value)
+  ) %>%
+  knitr::kable(
+    caption = "Linear Regression Model Results Estimating Crimes Per Week",
+    col.names = c("Predictor", "B", "SE", "t", "p"),
+    digits = c(0, 2, 3, 2, 3),
+    align = c("l", "r", "r", "r", "r")
+  )
 ```
+
 ### Interpretation
@@
 
-```{r}
+```{r, echo=FALSE}
 lm_income_crimes <- lm(number_of_crimes ~ cali_median_income, data = crime_income_per_zip)
@@
 summary(lm_income_crimes)
+
+lm_income_crimes %>%
+tidy() %>%
+  mutate(
+    p.value = scales::pvalue(p.value)
+  ) %>%
+  knitr::kable(
+    caption = "Linear Regression Model Results Estimating Crimes Based on Median Income",
+    col.names = c("Predictor", "B", "SE", "t", "p"),
+    digits = c(0, 2, 3, 2, 3),
+    align = c("l", "r", "r", "r", "r")
+  )
 ```
+
 ### Interpretation
@@
 Based on the p-value, we observe that median income is not significant at a 0.05 level, meaning it does not significantly decrease the amount of crimes committed.
-The r^2 value additionally tells us that the income only explains a small portion of the variation in the number of crimes.
+The r\^2 value additionally tells us that the income only explains a small portion of the variation in the number of crimes.
 As stated with the scatterplot, a larger sample and considering other factors might yield different results.
 
-# Bibliography
+# References {.unnumbered}
+
+::: {#refs}
+:::
+
+# Appendix {.unnumbered}
+
+```{r, eval=FALSE}
+library(sf) # for map
+library(tidyverse) # using only for joining datasets to build map visual
+library(RColorBrewer) # for map color scheme
+library(tidygeocoder) # to get zip codes of addresses
+
+# Primary datasets: Crime Data 2020-Present and 2010-2019
+crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")
+
+crime_la_2020_present <- read.csv("Raw_Data/Crime_Data_from_2020_to_Present.csv")
+
+# Combine the two datasets
+crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
+
+# Secondary datasets
+california_median_income_2018_2023 <- list()
+yrs <- 2018:2023
+for (i in 1:length(yrs)) {
+  obj <- read.csv(paste0("Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
+  obj <- obj[-1,]
+  obj$year <- yrs[[i]]
+  california_median_income_2018_2023[[i]] <- obj
+}
+
+cali_median_income <- do.call(rbind, california_median_income_2018_2023)
+
+# adding a variable that's just the ZCTA zip code
+cali_median_income$ZCTA <- stringr::str_sub(cali_median_income$NAME, -5)
+```
+
+## Figure 1 - Examining Crime Density from 2010–2019 and 2020-Present {.unnumbered}
+
+```{r, eval=FALSE}
+
+par(mfrow = c(1, 2), mar = c(4, 4, 2, 1))
+
+smoothScatter(
+  x    = crime_la_2010_2019$LON,
+  y    = crime_la_2010_2019$LAT,
+  xlim = c(-119, -118),
+  ylim = c(33.5, 34.5),
+  nbin = 300,
+  xlab = "Longitude",
+  ylab = "Latitude",
+  main = "Crime Density (2010–2019)",
+  colramp = colorRampPalette(c("white", "skyblue", "blue", "darkblue")),
+  cex.main=0.8
+)
+
+smoothScatter(
+  x    = crime_la_2020_present$LON,
+  y    = crime_la_2020_present$LAT,
+  xlim = c(-119, -118),
+  ylim = c(33.5, 34.5),
+  nbin = 300,
+  xlab = "Longitude",
+  ylab = "Latitude",
+  main = "Crime Density (2020–Present)",
+  colramp = colorRampPalette(c("white", "pink", "red", "darkred")),
+  cex.main=0.8
+)
+
+par(mfrow = c(1, 1))
+
+```
+
+## Figure 2 - Examining Number Of Crimes Over Time {.unnumbered}
+
+```{r, eval=FALSE}
+crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
+                                  format = "%m/%d/%Y %I:%M:%S %p", 
+                                  tz = "America/Los_Angeles")
+
+crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
+
+weekly_counts <- aggregate(DR_NO ~ YearWeek, data = crime_la, FUN = length)
+names(weekly_counts)[2] <- "Crimes"
+
+weekly_counts <- weekly_counts[order(weekly_counts$YearWeek), ]
+
+weekly_counts$Index <- seq_len(nrow(weekly_counts))
+
+plot(
+  weekly_counts$Index,
+  weekly_counts$Crimes,
+  type = "l",
+  col  = "blue",
+  lwd  = 2,
+  xlab = "Weekly Index",
+  ylab = "Number of Crimes",
+  main = "Weekly Crime Counts in Los Angeles (2010–Present)"
+)
+
+abline(h = pretty(weekly_counts$Crimes), v = pretty(weekly_counts$Index), 
+       col = "gray", lty = "dotted")
+```
+
+## Figure 3 - Crime rate frequency by day of the week {.unnumbered}
+
+```{r, eval=FALSE}
+# Convert Date.Rptd to Date format (ignore time part)
+#crime_la$Date.Rptd
+crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")
+
+# Set locale to English (for weekday names)
+Sys.setlocale("LC_TIME", "C")
+
+# Extract the day of the week
+crime_la$DayOfWeek <- weekdays(crime_la$Date)
+# crime_la$DayOfWeek
+
+# Count occurrences of crimes per day
+crime_counts <- table(crime_la$DayOfWeek)
+# crime_counts
+
+# Order the days correctly
+day_order <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
+crime_counts <- crime_counts[day_order]
+
+# Plot bar chart with horizontal labels
+barplot(crime_counts, col = "lightblue",
+        xlab = "Day of the Week", ylab = "Number of Crimes",
+        names.arg = names(crime_counts), las = 1, cex.names = 0.8, cex.axis = 0.6)
+```
+
+## Figure 4 - Gender Distribution for crimes over time {.unnumbered}
+
+```{r, eval=FALSE}
+# Remove missing values and empty strings in Vict.Sex column
+crime_la <- subset(crime_la, Vict.Sex != "" & !is.na(Vict.Sex))
+
+# Replace all non-"F" and non-"M" values with "N"
+crime_la$Vict.Sex <- ifelse(crime_la$Vict.Sex %in% c("F", "M"), 
+                                  crime_la$Vict.Sex, "N")
+
+# Count occurrences of each gender
+gender_counts <- table(crime_la$Vict.Sex)
+
+# Create a pie chart using base R
+color <- c("F" = "pink", "M" = "lightblue", "N" = "purple")
+pie(
+  gender_counts, 
+  labels = paste(names(gender_counts), 
+                 round(gender_counts / sum(gender_counts) * 100, 1), "%"), 
+  col = color[names(gender_counts)],
+  main = "Distribution of Crime Victims by Gender"
+)
+```
+
+## Figure 5 - Median Household Income by Zip Code in Los Angeles by Year {.unnumbered}
+
+```{r, eval=FALSE}
+zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
+LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs
+
+colnames(LA_zctas) <- c("X", "ZCTA5CE20")
+LA_zctas$ZCTA5CE20 <- as.character(LA_zctas$ZCTA5CE20)
+
+# joining the zcta shape file with LA zcta reference file
+zcta_filtered <- zcta %>% inner_join(LA_zctas, by = "ZCTA5CE20")
+
+# getting LA median income using joined file
+median_household_income <- cali_median_income %>% select(ZCTA, S1903_C03_001E, year)
+
+zcta_with_data <- zcta_filtered %>% inner_join(median_household_income, join_by("ZCTA5CE20" == "ZCTA"))
+
+# remove NAs
+zcta_no_NAs <- zcta_with_data %>% filter(!is.na(S1903_C03_001E), S1903_C03_001E != "-")
+
+zcta_no_NAs$S1903_C03_001E <- as.integer(zcta_no_NAs$S1903_C03_001E)
+```
+
+```{r, eval=FALSE}
+# quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
+qntls <- round(quantile(zcta_no_NAs$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0)
+
+# 2x3 plot layout for years 2018-2023, setting margins and outer margins, main title sizes
+par(mfrow = c(2, 3), mar = c(1, 1, 1, 5), oma = c(0, 0, 3, 0), cex.main = 1.5)
+
+# color palette for sequential data
+palette <- brewer.pal(7, "Greens")
+
+yrs <- 2018:2023
+
+for (yr in yrs) {
+  # get just that year's data
+  subset_data <- zcta_no_NAs[zcta_no_NAs$year == yr, ]
+
+  # creating bins based on the qtls
+  bins <- cut(subset_data$S1903_C03_001E, breaks = qntls, labels = FALSE, include.lowest = TRUE)
+
+  # Plot
+  plot(st_geometry(subset_data),
+       col = palette[bins],
+       border = 'grey',
+       axes = FALSE,
+       lwd = 0.5)
+  title(yr, line = -1)
+
+  # adding legend
+  legend("right", inset = c(-.75, 0), legend = qntls, fill = palette, title = "Median Income ($)", bty = "n", cex = 0.8, xpd = TRUE)
+}
+
+# main title for all plots
+mtext("Median Household Income by Year and Zip Code in LA", outer = TRUE, cex = 1, font = 2)
+
+```
+
+## Figure 6 - Crime Rates Compared to Median Household Income {.unnumbered}
+
+```{r, eval=FALSE}
+# using latitude and longitude to obtain the zip codes of the crime locations
+# crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
+# id_lat_long_crime_la <- subset(crime_la, select = c("LAT", "LON", "DR_NO", "year"))
+
+# taking a random sample to reduce load on model
+#set.seed(123)
+#sampled_ids <- id_lat_long_crime_la[sample(1:nrow(id_lat_long_crime_la), 500),]
+#rownames(sampled_ids) <- NULL
+#zip_codes <- reverse_geocode(sampled_ids, lat = LAT, long = LON, verbose = TRUE)
+#write.csv(zip_codes, "sampled_crimes.csv")
+zip_codes <- read.csv("sampled_crimes.csv")
+
+zip_codes$ZCTA5CE20 <- stringr::str_extract(zip_codes$address, "\\d{5}")
+
+crime_income_sample <- zip_codes %>% inner_join(unique(subset(zcta_no_NAs, select = c("ZCTA5CE20", "S1903_C03_001E", "year"))), by = c("ZCTA5CE20", "year"))
+
+# this takes a while, so loading in this csv instead
+
+crime_income_sample <- zip_codes %>% inner_join(unique(subset(zcta_no_NAs, select = c("ZCTA5CE20", "S1903_C03_001E", "year"))), by = c("ZCTA5CE20", "year"))
+
+crime_income_per_zip <- crime_income_sample %>% 
+  group_by(ZCTA5CE20) %>% 
+  summarize(number_of_crimes = n(),
+            cali_median_income = mean(S1903_C03_001E)) %>%
+  ungroup()
+
+plot(x = crime_income_per_zip$cali_median_income, 
+     y = crime_income_per_zip$number_of_crimes, 
+     type = "p",
+     ylab = "Number of Crimes",  
+     xlab = "Median Income ($)",  
+     main = "Crime vs. Median Income by ZIP Code",
+     bty = "l")
+```
+
+## Crime \~ Week Regression Code {.unnumbered}
+
+```{r, eval=FALSE}
+library(broom)
+
+weekly_counts$CrimesLag <- c(NA, weekly_counts$Crimes[-nrow(weekly_counts)])
+
+ar1_trend_regression <- lm(Crimes ~ Index + CrimesLag, data = weekly_counts, na.action = na.omit)
+
+ar1_trend_regression %>%
+tidy() %>%
+  mutate(
+    p.value = scales::pvalue(p.value)
+  ) %>%
+  knitr::kable(
+    caption = "Linear Regression Model Results Estimating Crimes Per Week",
+    col.names = c("Predictor", "B", "SE", "t", "p"),
+    digits = c(0, 2, 3, 2, 3),
+    align = c("l", "r", "r", "r", "r")
+  )
+```
+
+## Crime \~ Income Regression Code {.unnumbered}
+
+```{r, eval=FALSE}
+lm_income_crimes <- lm(number_of_crimes ~ cali_median_income, data = crime_income_per_zip)
+
+summary(lm_income_crimes)
+
+lm_income_crimes %>%
+tidy() %>%
+  mutate(
+    p.value = scales::pvalue(p.value)
+  ) %>%
+  knitr::kable(
+    caption = "Linear Regression Model Results Estimating Crimes Based on Median Income",
+    col.names = c("Predictor", "B", "SE", "t", "p"),
+    digits = c(0, 2, 3, 2, 3),
+    align = c("l", "r", "r", "r", "r")
+  )
+```

### 2025-02-25 Draft1.qmd
fixed appendix

@@
 
-# Plots
+# Exploratory Data Analysis
 
@@
 
+### Interpretation
+
 Both time periods show major clusters of reported crimes in central LA. The 2020–present data appears slightly more dispersed to the north, but overall patterns remain similar looking.
@@
 ```
+### Interpretation
 
@@
 ```
-
-Now, using this joined dataset, create six maps, one for each year.
 
@@
 
-## Basic Regression Between Median Income and Crimes per Zip Code
+## Regression Between Median Income and Crimes per Zip Code
 
@@
 
-summary(lm_income_crimes)
+# summary(lm_income_crimes)
 
@@
 
+## ggplot2 ideas
+
+* histogram (we did a barplot)
+* boxplots!
+* geom_tile - correlations
+*
+
 # References {.unnumbered}
@@
 
-# Appendix {.unnumbered}
+# Appendices {.appendix}
 
@@
 
-## Figure 1 - Examining Crime Density from 2010–2019 and 2020-Present {.unnumbered}
+# Figure 1 - Examining Crime Density from 2010–2019 and 2020-Present {.unnumbered}
 
@@
 
-## Figure 2 - Examining Number Of Crimes Over Time {.unnumbered}
+# Figure 2 - Examining Number Of Crimes Over Time {.unnumbered}
 
@@
 
-## Figure 3 - Crime rate frequency by day of the week {.unnumbered}
+# Figure 3 - Crime rate frequency by day of the week {.unnumbered}
 
@@
 
-## Figure 5 - Median Household Income by Zip Code in Los Angeles by Year {.unnumbered}
+# Figure 5 - Median Household Income by Zip Code in Los Angeles by Year {.unnumbered}
 
@@
 
-## Figure 6 - Crime Rates Compared to Median Household Income {.unnumbered}
+# Figure 6 - Crime Rates Compared to Median Household Income {.unnumbered}
 

### 2025-02-25 Draft1.qmd
added ggplot ideas

@@
 crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
+```
 
-# Secondary datasets
+```{r}
+# Secondary dataset - Median Income by ZCTA
 california_median_income_2018_2023 <- list()
@@
 cali_median_income$ZCTA <- stringr::str_sub(cali_median_income$NAME, -5)
+```
+
+```{r}
+# Secondary dataset - COVID cases in Los Angeles County
+covid_14day <- read.csv("Raw_Data/LA_County_Covid19_CSA_14day_case_death_table.csv")
+
+covid_la <- covid_14day %>% filter(grepl("Los Angeles", geo_merge))
+covid_la$area <- gsub("Los Angeles - ", "", covid_la$geo_merge)
 ```
@@
 * histogram (we did a barplot)
-* boxplots!
-* geom_tile - correlations
-*
+* boxplots! - 
+* geom_tile - correlations - Bianca
+* geom_jitter
 
@@
 
-## Figure 4 - Gender Distribution for crimes over time {.unnumbered}
+# Figure 4 - Gender Distribution for crimes over time {.unnumbered}
 

### 2025-02-25 Draft1.qmd
ggplot code

@@
 
-## ggplot2 ideas
+# ggplot2 ideas
 
-* histogram (we did a barplot)
-* boxplots! - 
-* geom_tile - correlations - Bianca
-* geom_jitter
+## Histogram
+
+## Boxplot - Kaitlyn
+
+```{r}
+crime_by_area <- crime_la %>% group_by(AREA.NAME) %>% summarize(numberOfCrimes = n()) %>% ungroup()
+
+bureaus <- read.csv("Raw_Data/Division_Bureau_Ref.csv")
+
+bureaus$Division <- gsub("77th", "77th Street", bureaus$Division)
+bureaus$Division <- gsub("North", "N", bureaus$Division)
+bureaus$Division <- gsub("Neast", "Northeast", bureaus$Division)
+
+crime_by_area <- crime_by_area %>% left_join(bureaus, join_by("AREA.NAME" == "Division"))
+```
+
+## geom_tile for correlations - Bianca
+
+## geom_jitter
 

### 2025-02-26 Draft1.qmd
added code used to obtain all but 3,000 ZCTA codes/addresses

@@
 crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
+```
+
+```{r}
+crimes_with_zips <- read.csv("crimes_with_zips.csv") # just inner_join with the crime_la dataset
 ```
@@
 
+## Data Wrangling {.unnumbered}
+
+# Reading in Raw Datasets
+
 ```{r, eval=FALSE}
@@
 cali_median_income$ZCTA <- stringr::str_sub(cali_median_income$NAME, -5)
+
+write.csv(cali_median_income, "cali_median_income.csv")
+```
+
+### Zip Codes for Crimes {.unnumbered}
+```{r, eval=FALSE}
+library(sf)
+library(dplyr)
+
+crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")
+
+crime_la_2020_present <- read.csv("Raw_Data/Crime_Data_from_2020_to_Present.csv")
+
+# Combine the two datasets
+crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
+
+crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
+                                  format = "%m/%d/%Y %I:%M:%S %p", 
+                                  tz = "America/Los_Angeles")
+
+crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
+
+crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
+
+coords <- crime_la %>% select(LAT, LON, DR_NO, year)
+
+zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
+
+  points <- st_as_sf(coords, coords = c("LON", "LAT"), crs = 4326)  # WGS84
+
+  zcta <- st_transform(zcta, 4326)  # Convert ZCTA to WGS84 if needed
+  cat("Starting Zip Code Retrieval \n")
+
+  result <- st_join(points, zcta)
+
+  cat("Finished Zip Code Retrieval \n")
+
+  write.csv(result, "crime_zip_files/all_zips.csv")
+
+crime_zctas <- read.csv("Raw_Data/all_zips.csv")
+colnames(crime_zctas) <- c(colnames(crime_zctas)[-c(1, length(colnames(crime_zctas)))],"LON", "LAT")
+
+crime_zctas$LON <- as.numeric(gsub('c(', "", crime_zctas$LON, fixed = TRUE))
+crime_zctas$LAT <- as.numeric(gsub(')', "", crime_zctas$LAT, fixed = TRUE))
+```
+
+```{r}
+
+duplicates <- as.integer(as.character(data.frame(table(crime_zctas$universal_idx)[table(crime_zctas$universal_idx) > 1]) %>% pull(Var1)))
+
+# there are approximately 200 rows that are duplicated (100 incidents) as the area lies on the intersection between two zctas. 
+duplicate_data <- crime_zctas %>% filter(universal_idx %in% duplicates)
+
+rev_geo_duplicates <- reverse_geocode(duplicates2, lat = LAT, long = LON)
+# the above query identifies the address of these coordinates as KFC, 8644, Balboa Boulevard, Northridge South Neighborhood Council District, Los Angeles, Los Angeles County, California, 91325, United States.
+# therefore, we will only keep the 91325 zip code rows.
+
+crime_zctas_no_duplicates <- crime_zctas %>% mutate(duplicate = ifelse(universal_idx %in% duplicate_data$universal_idx & ZCTA5CE20 != 91325, "Drop", "Keep")) %>% filter(duplicate != "Drop") %>% select(-duplicate)
+
+# it's empty now 
+# duplicates <- table(crime_zctas_no_duplicates$universal_idx)[table(crime_zctas_no_duplicates$universal_idx) > 1]
+
+# next, check for NAs
+sum(is.na(crime_zctas_no_duplicates$ZCTA5CE20)) # there are 3724 NAs. Let's find them a zip code
+
+# some also have lat long = 0. We'll find those separately
+no_zipcode_not0 <- crime_zctas_no_duplicates %>% filter(is.na(ZCTA5CE20) & LON != 0)
+
+no_zipcode_not0_zips <- reverse_geocode(no_zipcode_not0, lat = LAT, long = LON)
+
+no_zipcode_not0_zips$ZCTA5CE20 <- stringr::str_extract(no_zipcode_not0_zips$address, "\\d{5}")
+
+no_zipcode_not0_zips <- no_zipcode_not0_zips %>% select(-address)
+
+crime_zctas2 <- crime_zctas_no_duplicates %>% filter(!is.na(ZCTA5CE20))
+
+crime_zctas3 <- rbind(crime_zctas2, no_zipcode_not0_zips)
+
+crime_zctas4 <- crime_zctas3 %>% filter(!is.na(ZCTA5CE20))
+
+write.csv(crime_zctas4, "crimes_with_zips.csv")
+
+no_zipcode_0s <- crime_zctas_no_duplicates %>% filter(is.na(ZCTA5CE20) & LON == 0) %>% pull(DR_NO)
+
+main_dataset_0s <- crime_la %>% filter(DR_NO %in% no_zipcode_0s)
+
+summary(main_dataset_0s)
+
+summary(crime_la)
 ```

### 2025-02-27 Draft1.qmd
cleaned up beginning data, moved messier raw data wrangling to the end. Cleaned data csvs are in the cleaned_data folder in Drive.

@@
 
-# Primary datasets: Crime Data 2020-Present and 2010-2019
-crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")
+# primary dataset
+crime_la <- read.csv("Cleaned_Data/all_crimes_2010_present.csv")
 
-crime_la_2020_present <- read.csv("Raw_Data/Crime_Data_from_2020_to_Present.csv")
+# primary dataset zip code and geometry data
+crimes_with_zips <- read.csv("Cleaned_Data/crimes_with_zips.csv") # just inner_join with the crime_la dataset
 
-# Combine the two datasets
-crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
-```
-
-```{r}
-crimes_with_zips <- read.csv("crimes_with_zips.csv") # just inner_join with the crime_la dataset
-```
-
-```{r}
 # Secondary dataset - Median Income by ZCTA
@@
 
-cali_median_income <- do.call(rbind, california_median_income_2018_2023)
+cali_median_income <- read.csv("Cleaned_Data/cali_median_income.csv")
 
@@
 cali_median_income$ZCTA <- stringr::str_sub(cali_median_income$NAME, -5)
-```
 
-```{r}
 # Secondary dataset - COVID cases in Los Angeles County
@@
 
-```{r, eval=FALSE}
+```{r Wrangling Main, eval=FALSE}
 library(sf) # for map
@@
 crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
+```
 
+```{r Combining Median Income, eval=FALSE}
 # Secondary datasets
@@
 ### Zip Codes for Crimes {.unnumbered}
-```{r, eval=FALSE}
+```{r Get Zip Codes, eval=FALSE}
 library(sf)
@@
 
-```{r}
+```{r Get Zip Codes 2, eval=FALSE}
 
@@
 
-```{r, eval=FALSE}
+```{r Figure 1 Code, eval=FALSE}
 
@@
 
-```{r, eval=FALSE}
+```{r Figure 2 Code, eval=FALSE}
 crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
@@
 
-```{r, eval=FALSE}
+```{r Figure 3 Code, eval=FALSE}
 # Convert Date.Rptd to Date format (ignore time part)
@@
 
-```{r, eval=FALSE}
+```{r Figure 4 Code, eval=FALSE}
 # Remove missing values and empty strings in Vict.Sex column
@@
 
-```{r, eval=FALSE}
+```{r Figure 5 Wrangling, eval=FALSE}
 zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
@@
 
-```{r, eval=FALSE}
+```{r Figure 5 Code, eval=FALSE}
 # quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
@@
 
-```{r, eval=FALSE}
+```{r Figure 6 Code, eval=FALSE}
 # using latitude and longitude to obtain the zip codes of the crime locations
@@
 
-```{r, eval=FALSE}
+```{r Regression 1, eval=FALSE}
 library(broom)
@@
 
-```{r, eval=FALSE}
+```{r Regression 2, eval=FALSE}
 lm_income_crimes <- lm(number_of_crimes ~ cali_median_income, data = crime_income_per_zip)

### 2025-02-27 Draft1.qmd
updated code

@@
 # Secondary dataset - Median Income by ZCTA
-california_median_income_2018_2023 <- list()
-yrs <- 2018:2023
-for (i in 1:length(yrs)) {
-  obj <- read.csv(paste0("Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
-  obj <- obj[-1,]
-  obj$year <- yrs[[i]]
-  california_median_income_2018_2023[[i]] <- obj
-}
-
 cali_median_income <- read.csv("Cleaned_Data/cali_median_income.csv")

### 2025-02-27 Draft1.qmd
histogram

@@
 
-```{r, output=FALSE, echo=FALSE}
+```{r DATA, output=FALSE, echo=FALSE}
 library(sf) # for map
@@
 covid_la$area <- gsub("Los Angeles - ", "", covid_la$geo_merge)
+
+evictions_la <- read.csv("Raw_Data/LA_evictions.csv")
 ```
@@
 
+# Exploratory Data Analysis With ggplot2
+
+## Histogram
+```{r}
+# fix mistake
+evictions_la$Date.Received <- gsub("2205", "2025", evictions_la$Date.Received)
+
+crime_la_monthly <- crime_la %>% group_by(Year)
+# convert to date
+evictions_la$month_year <- format(as.Date(evictions_la$Date.Received, format="%m-%d-%Y"), "%Y-%m")
+
+# get unique months
+unique_months <- data.frame(month_year = sort(unique(evictions_la$month_year)), index = 1:length(unique(evictions_la$month_year)))
+
+evictions_la_indexed <- evictions_la %>% left_join(unique_months, by = "month_year")
+
+ggplot(data = evictions_la_indexed, aes(x = index)) + 
+  geom_histogram(bins = nrow(unique_months)) +
+  scale_x_continuous(breaks=unique_months$index,
+        labels=unique_months$month_year) + 
+  theme(axis.text.x = element_text(angle = 285, vjust = 0.5, hjust=1),
+        axis.title.x = element_text(vjust = -.75),
+        plot.title = element_text(size=12)) + 
+  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", title = "Monthly Frequency of Evictions in the City of Los Angeles \nFebruary 2023 to February 2025") +
+
+```
+
+## Boxplot - Kaitlyn
+
+```{r}
+crime_by_area <- crime_la %>% group_by(AREA.NAME) %>% summarize(numberOfCrimes = n()) %>% ungroup()
+
+bureaus <- read.csv("Raw_Data/Division_Bureau_Ref.csv")
+
+bureaus$Division <- gsub("77th", "77th Street", bureaus$Division)
+bureaus$Division <- gsub("North", "N", bureaus$Division)
+bureaus$Division <- gsub("Neast", "Northeast", bureaus$Division)
+
+crime_by_area <- crime_by_area %>% left_join(bureaus, join_by("AREA.NAME" == "Division"))
+```
+
+## geom_tile for correlations - Bianca
+
+```{r, echo=FALSE}
+# creating dataset for correlations
+
+```
+
+## geom_jitter
+
+## Plot 5
+
+## Plot 6
+
 # Regression Analyses
@@
 As stated with the scatterplot, a larger sample and considering other factors might yield different results.
-
-# ggplot2 ideas
-
-## Histogram
-
-## Boxplot - Kaitlyn
-
-```{r}
-crime_by_area <- crime_la %>% group_by(AREA.NAME) %>% summarize(numberOfCrimes = n()) %>% ungroup()
-
-bureaus <- read.csv("Raw_Data/Division_Bureau_Ref.csv")
-
-bureaus$Division <- gsub("77th", "77th Street", bureaus$Division)
-bureaus$Division <- gsub("North", "N", bureaus$Division)
-bureaus$Division <- gsub("Neast", "Northeast", bureaus$Division)
-
-crime_by_area <- crime_by_area %>% left_join(bureaus, join_by("AREA.NAME" == "Division"))
-```
-
-## geom_tile for correlations - Bianca
-
-## geom_jitter
 

### 2025-02-27 Draft1.qmd
completed histogram (plot 1 for draft 2)

@@
 
-crime_la_monthly <- crime_la %>% group_by(Year)
 # convert to date
@@
 
-ggplot(data = evictions_la_indexed, aes(x = index)) + 
-  geom_histogram(bins = nrow(unique_months)) +
+# create color scheme
+crime_la$month_year <- format(crime_la$DateParsed, "%Y-%m")
+crime_la_monthly <- crime_la %>% 
+  group_by(month_year) %>% 
+  summarize(num_crimes = n())
+
+avg_crimes <- mean(evictions_la_indexed_crimes$num_crimes)
+
+crime_la_monthly$high_low_crime <- as.factor(ifelse(crime_la_monthly$num_crimes > avg_crimes, "Above Average", "Below Average"))
+
+evictions_la_indexed_crimes <- evictions_la_indexed %>% left_join(crime_la_monthly, by = "month_year")
+
+ggplot(data = evictions_la_indexed_crimes, aes(x = index)) + 
+  geom_histogram(aes(fill = high_low_crime), bins = nrow(unique_months)) +
   scale_x_continuous(breaks=unique_months$index,
         labels=unique_months$month_year) + 
-  theme(axis.text.x = element_text(angle = 285, vjust = 0.5, hjust=1),
+  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", title = "Monthly Frequency of Evictions and Crimes in the City of Los Angeles \nFebruary 2023 to February 2025", fill = "Crime Level") +
+  ggthemes::scale_fill_fivethirtyeight() + 
+  ggthemes::theme_stata(scheme = "s1mono") +
+  theme(axis.text.x = element_text(angle = 300, vjust = 0.5, hjust = -.1),
         axis.title.x = element_text(vjust = -.75),
-        plot.title = element_text(size=12)) + 
-  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", title = "Monthly Frequency of Evictions in the City of Los Angeles \nFebruary 2023 to February 2025") +
-
+        plot.title = element_text(size=12))
 ```
+### Interpretation
 

### 2025-02-27 Draft1.qmd
added interpretation

@@
 
-## Histogram
+## Number of Evictions and Crime Levels Per Month
 ```{r}
@@
 
-avg_crimes <- mean(evictions_la_indexed_crimes$num_crimes)
-
 crime_la_monthly$high_low_crime <- as.factor(ifelse(crime_la_monthly$num_crimes > avg_crimes, "Above Average", "Below Average"))
@@
 evictions_la_indexed_crimes <- evictions_la_indexed %>% left_join(crime_la_monthly, by = "month_year")
+
+avg_crimes <- mean(evictions_la_indexed_crimes$num_crimes)
+
+evictions_la_indexed_crimes$high_low_crime <- as.factor(ifelse(evictions_la_indexed_crimes$num_crimes > avg_crimes, "Above Average", "Below Average"))
 
@@
 ### Interpretation
+This histogram plots the frequency of evictions in Los Angeles from February 2023 to February 2025, where the City of Los Angeles Housing Department has published case data.
+By summarizing both the crime and the evictions datasets by month, we see that the number of evictions was highest in early-mid 2023, and since, there have typically been slightly lower frequencies of evictions.
+Additionally, we split the number of crimes per month based on the average number of crimes of 14,439 to denote above and below average number of crimes per month.
+There is a clear split between March and April of 2024, as all months before that had above average crime, and all after had below average.
+Interestingly, the months where there is below average crime, seem to be the ones with slightly lower numbers of evictions.
 

### 2025-02-27 Draft1.qmd
echo False

@@
 ## Number of Evictions and Crime Levels Per Month
-```{r}
+```{r, echo=FALSE}
 # fix mistake

### 2025-02-27 Draft1.qmd
fixed data load

@@
 
-# primary dataset
+# primary datasets
+# Primary datasets: Crime Data 2020-Present and 2010-2019
+crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")
+
+crime_la_2020_present <- read.csv("Raw_Data/Crime_Data_from_2020_to_Present.csv")
+
 crime_la <- read.csv("Cleaned_Data/all_crimes_2010_present.csv")
@@
 
+# since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions. 
 ```
@@
 :::
+
+# Data Sources {.unnumbered}
+## Primary Datasets {.unnumbered}
+Crime Data From 2010-2019: https://data.lacity.org/Public-Safety/Crime-Data-from-2010-to-2019/63jg-8b9z/about_data
+
+Crime Data From 2020-Present: https://catalog.data.gov/dataset/crime-data-from-2020-to-present
+
+## Secondary Datasets {.unnumbered}
+Los Angeles County Eviction Data: https://calmatters.org/housing/homelessness/2023/11/california-evictions-post-pandemic/
+
+City of Los Angeles Eviction Data: https://housing.lacity.gov/residents/renters/eviction-notices-filed
+
+Median Income Data: https://data.census.gov/table/ACSST1Y2023.S1903?q=income
 

### 2025-02-27 Draft1.qmd
updated
regular save

@@
 
-crime_la_monthly$high_low_crime <- as.factor(ifelse(crime_la_monthly$num_crimes > avg_crimes, "Above Average", "Below Average"))
-
 evictions_la_indexed_crimes <- evictions_la_indexed %>% left_join(crime_la_monthly, by = "month_year")
@@
 
-```{r}
+```{r, echo=FALSE}
 crime_by_area <- crime_la %>% group_by(AREA.NAME) %>% summarize(numberOfCrimes = n()) %>% ungroup()
@@
 ```{r, echo=FALSE}
+library(lubridate)
 # creating dataset for correlations
@@
 # since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions. 
+LA_county_evictions <- read.csv("Raw_Data/LA_county_evictions.csv")
+
+# covid data
+covid_la$date <- ymd(covid_la$ep_date)
+covid_la$month <- month(covid_la$date)
+covid_la$year <- year(covid_la$date)
+covid_la$date <- day(covid_la$date)
+# since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
+covid_la$lagged <- covid_la$date + 7
+
+covid_la_lagged <- covid_la %>% filter(lagged %in% c(21, 35)) %>% arrange(month, year)
+
+covid_la_monthly <- covid_la_lagged %>% group_by(month, year) %>% summarize(cases = sum(cases_14day)) %>% ungroup() %>% arrange(year, month)
+
+covid_la_monthly$index <- 1:nrow(covid_la_monthly)
+# base plot shows that this does in fact mirror LA county so the case numbers should be accurate
+# plot(y = covid_la_monthly$cases, x = covid_la_monthly$index, type = "l")
+
+crime_la_monthly$date <- ymd(paste0(crime_la_monthly$month_year, "-01"))
+crime_la_monthly$month <- month(crime_la_monthly$date)
+crime_la_monthly$year <- year(crime_la_monthly$date)
 ```

### 2025-02-27 Draft1.qmd
Added Boxplot
updated code

@@
 covid_14day <- read.csv("Raw_Data/LA_County_Covid19_CSA_14day_case_death_table.csv")
-
-covid_la <- covid_14day %>% filter(grepl("Los Angeles", geo_merge))
-covid_la$area <- gsub("Los Angeles - ", "", covid_la$geo_merge)
 
@@
 crime_by_area <- crime_by_area %>% left_join(bureaus, join_by("AREA.NAME" == "Division"))
+
+ggplot(crime_by_area, aes(y = numberOfCrimes, group = Bureau, x = Bureau)) + 
+        geom_boxplot(aes(fill = Bureau))
 ```
@@
 # covid data
-covid_la$date <- ymd(covid_la$ep_date)
-covid_la$month <- month(covid_la$date)
-covid_la$year <- year(covid_la$date)
-covid_la$date <- day(covid_la$date)
+covid_14day$date <- ymd(covid_la$ep_date)
+covid_14day$month <- month(covid_la$date)
+covid_14day$year <- year(covid_la$date)
+covid_14day$date <- day(covid_la$date)
 # since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
-covid_la$lagged <- covid_la$date + 7
+covid_14day$lagged <- covid_la$date + 7
 
-covid_la_lagged <- covid_la %>% filter(lagged %in% c(21, 35)) %>% arrange(month, year)
+covid_la_lagged <- covid_14day %>% filter(lagged %in% c(21, 35)) %>% arrange(month, year)
 

### 2025-02-28 Draft2.qmd
Creating Draft2 version to submit

--- 2025-02-27 Draft1.qmd
@@
   ggthemes::theme_stata(scheme = "s1mono") +
-  theme(axis.text.x = element_text(angle = 300, vjust = 0.5, hjust = -.1),
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
         axis.title.x = element_text(vjust = -.75),
@@
 # covid data
-covid_14day$date <- ymd(covid_la$ep_date)
-covid_14day$month <- month(covid_la$date)
-covid_14day$year <- year(covid_la$date)
-covid_14day$date <- day(covid_la$date)
+covid_14day$date <- ymd(covid_14day$ep_date)
+covid_14day$month <- month(covid_14day$date)
+covid_14day$year <- year(covid_14day$date)
+covid_14day$date <- day(covid_14day$date)
 # since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
-covid_14day$lagged <- covid_la$date + 7
+covid_14day$lagged <- covid_14day$date + 7
 
@@
 covid_la_monthly$index <- 1:nrow(covid_la_monthly)
-# base plot shows that this does in fact mirror LA county so the case numbers should be accurate
-# plot(y = covid_la_monthly$cases, x = covid_la_monthly$index, type = "l")
+
+# next, unemployment rate
+unemployment <- read.csv("Raw_Data/la_county_unemployment_rate.csv")
+unemployment$month <- as.numeric(gsub("M", "", unemployment$Period))
+
+# lapd budget
+lapd_budget <- read.csv("Raw_Data/lapd_budget.csv")
 
@@
 crime_la_monthly$year <- year(crime_la_monthly$date)
+
+correlations_df <- crime_la_monthly %>% inner_join(unemployment, join_by("year" == "Year", "month" == "month")) %>% select(month, year, Value, num_crimes) %>% rename("unemployment_rate" = "Value")
+
+correlations_df2 <- correlations_df %>% inner_join(covid_la_monthly, by = c("month", "year"))
+
+correlations_df3 <- correlations_df2 %>% inner_join(LA_county_evictions, join_by("year" == "Year", "month" == "Month")) %>% rename("num_evictions" = "Count")
+
+correlations_df4 <- correlations_df3 %>% inner_join(lapd_budget, by = "year")
+
+correlations <- correlations_df4 %>% select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)
+
+Cor_mat <- round(cor(correlations), 2)
+
+long_Cor <- reshape2::melt(Cor_mat)
+
+long_Cor <- long_Cor %>%
+    mutate(Var1 = recode(Var1, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget ($ billions)"),
+           Var2 = recode(Var2, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget \n ($ billions)"))
+
+ggplot(data = long_Cor, aes(x = Var1, y = Var2, fill = value)) +
+  geom_tile(color = "white") +
+  scale_fill_viridis_c(direction = -1) +
+  theme(axis.title.x =  element_blank(),
+        axis.title.y = element_blank()) +
+  geom_text(aes(label = value), color = "black", size = 4) + 
+   theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.title = element_text(size=12)) + 
+  ggtitle("Correlations Between the Number of Crimes and \n Different Law Enforcement and Socioeconomic Factors")
 ```

### 2025-02-28 Draft2.qmd
Added Marco's charts to the second draft

@@
 ## geom_jitter
+## geom_jitter
 
-## Plot 5
+```{r, echo=FALSE}
+ggplot(crime_la, aes(x = DateParsed, y = Vict.Sex, color = Vict.Sex)) +
+  geom_jitter(alpha = 0.4, width = 0.2) +
+  labs(title = "Crime Over Time by Gender",
+       x = "Date",
+       y = "Gender") +
+  theme_minimal()
+
+```
+
+## Density Plot
+
+```{r, echo=FALSE}
+
+crime_la$Hour <- as.numeric(format(crime_la$DateParsed, "%H"))
+
+ggplot(crime_la, aes(x = Hour, fill = ..density..)) +
+  geom_density(alpha = 0.5, color = "black") +
+  scale_fill_gradient(low = "lightblue", high = "darkblue") +
+  labs(title = "Crime Density Per Hour",
+       x = "Hour",
+       y = "Density") +
+  theme_minimal()
+
+```
 

### 2025-02-28 1_FirstDraft/Draft1.qmd
Organization

--- 2025-02-27 Draft1.qmd
@@
 
+```{r, echo=FALSE}
+ggplot(crime_la, aes(x = DateParsed, y = Vict.Sex, color = Vict.Sex)) +
+  geom_jitter(alpha = 0.4, width = 0.2) +
+  labs(title = "Crime Over Time by Gender",
+       x = "Date",
+       y = "Gender") +
+  theme_minimal()
+
+```
+
 ## Plot 5
+
+```{r, echo=FALSE}
+
+crime_la$Hour <- as.numeric(format(crime_la$DateParsed, "%H"))
+
+ggplot(crime_la, aes(x = Hour, fill = ..density..)) +
+  geom_density(alpha = 0.5, color = "black") +
+  scale_fill_gradient(low = "lightblue", high = "darkblue") +
+  labs(title = "Crime Density Per Hour",
+       x = "Hour",
+       y = "Density") +
+  theme_minimal()
+
+```
 

### 2025-02-28 2_SecondDraft/Draft2.qmd
Organization

--- 2025-02-28 Draft2.qmd
@@
 ```
-
-## geom_jitter
 ## geom_jitter

### 2025-02-28 1_FirstDraft/Draft1.qmd
Organization

--- 2025-02-27 Draft1.qmd
@@
   pdf: default
-number-sections: true
 ---
@@
 
-```{r DATA, output=FALSE, echo=FALSE}
+# Loading in packages and dataset
+
+```{r, output=FALSE}
 library(sf) # for map
@@
 
-# primary datasets
 # Primary datasets: Crime Data 2020-Present and 2010-2019
@@
 
-crime_la <- read.csv("Cleaned_Data/all_crimes_2010_present.csv")
+# Combine the two datasets
+crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
 
-# primary dataset zip code and geometry data
-crimes_with_zips <- read.csv("Cleaned_Data/crimes_with_zips.csv") # just inner_join with the crime_la dataset
+# Secondary datasets
+california_median_income_2018_2023 <- list()
+yrs <- 2018:2023
+for (i in 1:length(yrs)) {
+  obj <- read.csv(paste0("Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
+  obj <- obj[-1,]
+  obj$year <- yrs[[i]]
+  california_median_income_2018_2023[[i]] <- obj
+}
 
-# Secondary dataset - Median Income by ZCTA
-cali_median_income <- read.csv("Cleaned_Data/cali_median_income.csv")
+cali_median_income <- do.call(rbind, california_median_income_2018_2023)
 
@@
 cali_median_income$ZCTA <- stringr::str_sub(cali_median_income$NAME, -5)
-
-# Secondary dataset - COVID cases in Los Angeles County
-covid_14day <- read.csv("Raw_Data/LA_County_Covid19_CSA_14day_case_death_table.csv")
-
-evictions_la <- read.csv("Raw_Data/LA_evictions.csv")
 ```
 
-# Exploratory Data Analysis
+# Plots
 
@@
 
-```{r, echo=FALSE}
+```{r}
 
@@
 
-### Interpretation
-
 Both time periods show major clusters of reported crimes in central LA. The 2020–present data appears slightly more dispersed to the north, but overall patterns remain similar looking.
@@
 
-```{r, echo=FALSE}
+```{r}
 crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
@@
 ```
-### Interpretation
 
@@
 
-```{r, echo=FALSE, fig.cap = "Crime Occurrences by Day of the Week"}
+```{r}
 # Convert Date.Rptd to Date format (ignore time part)
@@
 barplot(crime_counts, col = "lightblue",
+        main = "Crime Occurrences by Day of the Week",
         xlab = "Day of the Week", ylab = "Number of Crimes",
@@
 
-```{r, echo=FALSE}
+```{r}
 # Remove missing values and empty strings in Vict.Sex column
@@
 
-```{r, echo=FALSE}
+```{r}
 zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
@@
 
-```{r, echo=FALSE}
+Now, using this joined dataset, create six maps, one for each year.
+
+```{r}
 # quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
@@
 ### Interpretation
-
 The maps demonstrate that median income in certain LA zip codes (particularly those to the west) have a much higher median household income than the rest of the city, but over the last six years, the median income in most other LA zip codes has increased.
@@
 
-```{r, echo=FALSE}
+```{r}
 # using latitude and longitude to obtain the zip codes of the crime locations
@@
 
-### Interpretation
-
-Based on this plot, we do not observe a clear trend between the number of crimes committed and the median income in that zip code.
+### Based on this plot, we do not observe a clear trend between the number of crimes committed and the median income in that zip code.
 However, this scatterplot only accounts for a small portion of the dataset and averages across pre- and post-pandemic conditions.
@@
 
-# Exploratory Data Analysis With ggplot2
-
-## Number of Evictions and Crime Levels Per Month
-```{r, echo=FALSE}
-# fix mistake
-evictions_la$Date.Received <- gsub("2205", "2025", evictions_la$Date.Received)
-
-# convert to date
-evictions_la$month_year <- format(as.Date(evictions_la$Date.Received, format="%m-%d-%Y"), "%Y-%m")
-
-# get unique months
-unique_months <- data.frame(month_year = sort(unique(evictions_la$month_year)), index = 1:length(unique(evictions_la$month_year)))
-
-evictions_la_indexed <- evictions_la %>% left_join(unique_months, by = "month_year")
-
-# create color scheme
-crime_la$month_year <- format(crime_la$DateParsed, "%Y-%m")
-crime_la_monthly <- crime_la %>% 
-  group_by(month_year) %>% 
-  summarize(num_crimes = n())
-
-evictions_la_indexed_crimes <- evictions_la_indexed %>% left_join(crime_la_monthly, by = "month_year")
-
-avg_crimes <- mean(evictions_la_indexed_crimes$num_crimes)
-
-evictions_la_indexed_crimes$high_low_crime <- as.factor(ifelse(evictions_la_indexed_crimes$num_crimes > avg_crimes, "Above Average", "Below Average"))
-
-ggplot(data = evictions_la_indexed_crimes, aes(x = index)) + 
-  geom_histogram(aes(fill = high_low_crime), bins = nrow(unique_months)) +
-  scale_x_continuous(breaks=unique_months$index,
-        labels=unique_months$month_year) + 
-  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", title = "Monthly Frequency of Evictions and Crimes in the City of Los Angeles \nFebruary 2023 to February 2025", fill = "Crime Level") +
-  ggthemes::scale_fill_fivethirtyeight() + 
-  ggthemes::theme_stata(scheme = "s1mono") +
-  theme(axis.text.x = element_text(angle = 300, vjust = 0.5, hjust = -.1),
-        axis.title.x = element_text(vjust = -.75),
-        plot.title = element_text(size=12))
-```
-### Interpretation
-This histogram plots the frequency of evictions in Los Angeles from February 2023 to February 2025, where the City of Los Angeles Housing Department has published case data.
-By summarizing both the crime and the evictions datasets by month, we see that the number of evictions was highest in early-mid 2023, and since, there have typically been slightly lower frequencies of evictions.
-Additionally, we split the number of crimes per month based on the average number of crimes of 14,439 to denote above and below average number of crimes per month.
-There is a clear split between March and April of 2024, as all months before that had above average crime, and all after had below average.
-Interestingly, the months where there is below average crime, seem to be the ones with slightly lower numbers of evictions.
-
-## Boxplot - Kaitlyn
-
-```{r, echo=FALSE}
-crime_by_area <- crime_la %>% group_by(AREA.NAME) %>% summarize(numberOfCrimes = n()) %>% ungroup()
-
-bureaus <- read.csv("Raw_Data/Division_Bureau_Ref.csv")
-
-bureaus$Division <- gsub("77th", "77th Street", bureaus$Division)
-bureaus$Division <- gsub("North", "N", bureaus$Division)
-bureaus$Division <- gsub("Neast", "Northeast", bureaus$Division)
-
-crime_by_area <- crime_by_area %>% left_join(bureaus, join_by("AREA.NAME" == "Division"))
-
-ggplot(crime_by_area, aes(y = numberOfCrimes, group = Bureau, x = Bureau)) + 
-        geom_boxplot(aes(fill = Bureau))
-```
-
-## geom_tile for correlations - Bianca
-
-```{r, echo=FALSE}
-library(lubridate)
-# creating dataset for correlations
-
-# since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions. 
-LA_county_evictions <- read.csv("Raw_Data/LA_county_evictions.csv")
-
-# covid data
-covid_14day$date <- ymd(covid_la$ep_date)
-covid_14day$month <- month(covid_la$date)
-covid_14day$year <- year(covid_la$date)
-covid_14day$date <- day(covid_la$date)
-# since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
-covid_14day$lagged <- covid_la$date + 7
-
-covid_la_lagged <- covid_14day %>% filter(lagged %in% c(21, 35)) %>% arrange(month, year)
-
-covid_la_monthly <- covid_la_lagged %>% group_by(month, year) %>% summarize(cases = sum(cases_14day)) %>% ungroup() %>% arrange(year, month)
-
-covid_la_monthly$index <- 1:nrow(covid_la_monthly)
-# base plot shows that this does in fact mirror LA county so the case numbers should be accurate
-# plot(y = covid_la_monthly$cases, x = covid_la_monthly$index, type = "l")
-
-crime_la_monthly$date <- ymd(paste0(crime_la_monthly$month_year, "-01"))
-crime_la_monthly$month <- month(crime_la_monthly$date)
-crime_la_monthly$year <- year(crime_la_monthly$date)
-```
-
-## geom_jitter
-
-## Plot 5
-
-## Plot 6
-
 # Regression Analyses
 
-## Regression to Examine Weekly Crime Rates
+## Using Regression to Examine Weekly Crime Rates
 
-```{r, echo=FALSE}
-library(broom)
-
+```{r}
 weekly_counts$CrimesLag <- c(NA, weekly_counts$Crimes[-nrow(weekly_counts)])
@@
 
-ar1_trend_regression %>%
-tidy() %>%
-  mutate(
-    p.value = scales::pvalue(p.value)
-  ) %>%
-  knitr::kable(
-    caption = "Linear Regression Model Results Estimating Crimes Per Week",
-    col.names = c("Predictor", "B", "SE", "t", "p"),
-    digits = c(0, 2, 3, 2, 3),
-    align = c("l", "r", "r", "r", "r")
-  )
+summary(ar1_trend_regression)
 ```
-
 ### Interpretation
@@
 
-## Regression Between Median Income and Crimes per Zip Code
+## Basic Regression Between Median Income and Crimes per Zip Code
 
-```{r, echo=FALSE}
+```{r}
 lm_income_crimes <- lm(number_of_crimes ~ cali_median_income, data = crime_income_per_zip)
 
-# summary(lm_income_crimes)
-
-lm_income_crimes %>%
-tidy() %>%
-  mutate(
-    p.value = scales::pvalue(p.value)
-  ) %>%
-  knitr::kable(
-    caption = "Linear Regression Model Results Estimating Crimes Based on Median Income",
-    col.names = c("Predictor", "B", "SE", "t", "p"),
-    digits = c(0, 2, 3, 2, 3),
-    align = c("l", "r", "r", "r", "r")
-  )
+summary(lm_income_crimes)
 ```
-
 ### Interpretation
@@
 Based on the p-value, we observe that median income is not significant at a 0.05 level, meaning it does not significantly decrease the amount of crimes committed.
-The r\^2 value additionally tells us that the income only explains a small portion of the variation in the number of crimes.
+The r^2 value additionally tells us that the income only explains a small portion of the variation in the number of crimes.
 As stated with the scatterplot, a larger sample and considering other factors might yield different results.
-
-# References {.unnumbered}
-
-::: {#refs}
-:::
-
-# Data Sources {.unnumbered}
-## Primary Datasets {.unnumbered}
-Crime Data From 2010-2019: https://data.lacity.org/Public-Safety/Crime-Data-from-2010-to-2019/63jg-8b9z/about_data
-
-Crime Data From 2020-Present: https://catalog.data.gov/dataset/crime-data-from-2020-to-present
-
-## Secondary Datasets {.unnumbered}
-Los Angeles County Eviction Data: https://calmatters.org/housing/homelessness/2023/11/california-evictions-post-pandemic/
-
-City of Los Angeles Eviction Data: https://housing.lacity.gov/residents/renters/eviction-notices-filed
-
-Median Income Data: https://data.census.gov/table/ACSST1Y2023.S1903?q=income
-
-# Appendices {.appendix}
-
-## Data Wrangling {.unnumbered}
-
-# Reading in Raw Datasets
-
-```{r Wrangling Main, eval=FALSE}
-library(sf) # for map
-library(tidyverse) # using only for joining datasets to build map visual
-library(RColorBrewer) # for map color scheme
-library(tidygeocoder) # to get zip codes of addresses
-
-# Primary datasets: Crime Data 2020-Present and 2010-2019
-crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")
-
-crime_la_2020_present <- read.csv("Raw_Data/Crime_Data_from_2020_to_Present.csv")
-
-# Combine the two datasets
-crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
-```
-
-```{r Combining Median Income, eval=FALSE}
-# Secondary datasets
-california_median_income_2018_2023 <- list()
-yrs <- 2018:2023
-for (i in 1:length(yrs)) {
-  obj <- read.csv(paste0("Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
-  obj <- obj[-1,]
-  obj$year <- yrs[[i]]
-  california_median_income_2018_2023[[i]] <- obj
-}
-
-cali_median_income <- do.call(rbind, california_median_income_2018_2023)
-
-# adding a variable that's just the ZCTA zip code
-cali_median_income$ZCTA <- stringr::str_sub(cali_median_income$NAME, -5)
-
-write.csv(cali_median_income, "cali_median_income.csv")
-```
-
-### Zip Codes for Crimes {.unnumbered}
-```{r Get Zip Codes, eval=FALSE}
-library(sf)
-library(dplyr)
-
-crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")
-
-crime_la_2020_present <- read.csv("Raw_Data/Crime_Data_from_2020_to_Present.csv")
-
-# Combine the two datasets
-crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
-
-crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
-                                  format = "%m/%d/%Y %I:%M:%S %p", 
-                                  tz = "America/Los_Angeles")
-
-crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
-
-crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
-
-coords <- crime_la %>% select(LAT, LON, DR_NO, year)
-
-zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
-
-  points <- st_as_sf(coords, coords = c("LON", "LAT"), crs = 4326)  # WGS84
-
-  zcta <- st_transform(zcta, 4326)  # Convert ZCTA to WGS84 if needed
-  cat("Starting Zip Code Retrieval \n")
-
-  result <- st_join(points, zcta)
-
-  cat("Finished Zip Code Retrieval \n")
-
-  write.csv(result, "crime_zip_files/all_zips.csv")
-
-crime_zctas <- read.csv("Raw_Data/all_zips.csv")
-colnames(crime_zctas) <- c(colnames(crime_zctas)[-c(1, length(colnames(crime_zctas)))],"LON", "LAT")
-
-crime_zctas$LON <- as.numeric(gsub('c(', "", crime_zctas$LON, fixed = TRUE))
-crime_zctas$LAT <- as.numeric(gsub(')', "", crime_zctas$LAT, fixed = TRUE))
-```
-
-```{r Get Zip Codes 2, eval=FALSE}
-
-duplicates <- as.integer(as.character(data.frame(table(crime_zctas$universal_idx)[table(crime_zctas$universal_idx) > 1]) %>% pull(Var1)))
-
-# there are approximately 200 rows that are duplicated (100 incidents) as the area lies on the intersection between two zctas. 
-duplicate_data <- crime_zctas %>% filter(universal_idx %in% duplicates)
-
-rev_geo_duplicates <- reverse_geocode(duplicates2, lat = LAT, long = LON)
-# the above query identifies the address of these coordinates as KFC, 8644, Balboa Boulevard, Northridge South Neighborhood Council District, Los Angeles, Los Angeles County, California, 91325, United States.
-# therefore, we will only keep the 91325 zip code rows.
-
-crime_zctas_no_duplicates <- crime_zctas %>% mutate(duplicate = ifelse(universal_idx %in% duplicate_data$universal_idx & ZCTA5CE20 != 91325, "Drop", "Keep")) %>% filter(duplicate != "Drop") %>% select(-duplicate)
-
-# it's empty now 
-# duplicates <- table(crime_zctas_no_duplicates$universal_idx)[table(crime_zctas_no_duplicates$universal_idx) > 1]
-
-# next, check for NAs
-sum(is.na(crime_zctas_no_duplicates$ZCTA5CE20)) # there are 3724 NAs. Let's find them a zip code
-
-# some also have lat long = 0. We'll find those separately
-no_zipcode_not0 <- crime_zctas_no_duplicates %>% filter(is.na(ZCTA5CE20) & LON != 0)
-
-no_zipcode_not0_zips <- reverse_geocode(no_zipcode_not0, lat = LAT, long = LON)
-
-no_zipcode_not0_zips$ZCTA5CE20 <- stringr::str_extract(no_zipcode_not0_zips$address, "\\d{5}")
-
-no_zipcode_not0_zips <- no_zipcode_not0_zips %>% select(-address)
-
-crime_zctas2 <- crime_zctas_no_duplicates %>% filter(!is.na(ZCTA5CE20))
-
-crime_zctas3 <- rbind(crime_zctas2, no_zipcode_not0_zips)
-
-crime_zctas4 <- crime_zctas3 %>% filter(!is.na(ZCTA5CE20))
-
-write.csv(crime_zctas4, "crimes_with_zips.csv")
-
-no_zipcode_0s <- crime_zctas_no_duplicates %>% filter(is.na(ZCTA5CE20) & LON == 0) %>% pull(DR_NO)
-
-main_dataset_0s <- crime_la %>% filter(DR_NO %in% no_zipcode_0s)
-
-summary(main_dataset_0s)
-
-summary(crime_la)
-```
-
-# Figure 1 - Examining Crime Density from 2010–2019 and 2020-Present {.unnumbered}
-
-```{r Figure 1 Code, eval=FALSE}
-
-par(mfrow = c(1, 2), mar = c(4, 4, 2, 1))
-
-smoothScatter(
-  x    = crime_la_2010_2019$LON,
-  y    = crime_la_2010_2019$LAT,
-  xlim = c(-119, -118),
-  ylim = c(33.5, 34.5),
-  nbin = 300,
-  xlab = "Longitude",
-  ylab = "Latitude",
-  main = "Crime Density (2010–2019)",
-  colramp = colorRampPalette(c("white", "skyblue", "blue", "darkblue")),
-  cex.main=0.8
-)
-
-smoothScatter(
-  x    = crime_la_2020_present$LON,
-  y    = crime_la_2020_present$LAT,
-  xlim = c(-119, -118),
-  ylim = c(33.5, 34.5),
-  nbin = 300,
-  xlab = "Longitude",
-  ylab = "Latitude",
-  main = "Crime Density (2020–Present)",
-  colramp = colorRampPalette(c("white", "pink", "red", "darkred")),
-  cex.main=0.8
-)
-
-par(mfrow = c(1, 1))
-
-```
-
-# Figure 2 - Examining Number Of Crimes Over Time {.unnumbered}
-
-```{r Figure 2 Code, eval=FALSE}
-crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
-                                  format = "%m/%d/%Y %I:%M:%S %p", 
-                                  tz = "America/Los_Angeles")
-
-crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
-
-weekly_counts <- aggregate(DR_NO ~ YearWeek, data = crime_la, FUN = length)
-names(weekly_counts)[2] <- "Crimes"
-
-weekly_counts <- weekly_counts[order(weekly_counts$YearWeek), ]
-
-weekly_counts$Index <- seq_len(nrow(weekly_counts))
-
-plot(
-  weekly_counts$Index,
-  weekly_counts$Crimes,
-  type = "l",
-  col  = "blue",
-  lwd  = 2,
-  xlab = "Weekly Index",
-  ylab = "Number of Crimes",
-  main = "Weekly Crime Counts in Los Angeles (2010–Present)"
-)
-
-abline(h = pretty(weekly_counts$Crimes), v = pretty(weekly_counts$Index), 
-       col = "gray", lty = "dotted")
-```
-
-# Figure 3 - Crime rate frequency by day of the week {.unnumbered}
-
-```{r Figure 3 Code, eval=FALSE}
-# Convert Date.Rptd to Date format (ignore time part)
-#crime_la$Date.Rptd
-crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")
-
-# Set locale to English (for weekday names)
-Sys.setlocale("LC_TIME", "C")
-
-# Extract the day of the week
-crime_la$DayOfWeek <- weekdays(crime_la$Date)
-# crime_la$DayOfWeek
-
-# Count occurrences of crimes per day
-crime_counts <- table(crime_la$DayOfWeek)
-# crime_counts
-
-# Order the days correctly
-day_order <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
-crime_counts <- crime_counts[day_order]
-
-# Plot bar chart with horizontal labels
-barplot(crime_counts, col = "lightblue",
-        xlab = "Day of the Week", ylab = "Number of Crimes",
-        names.arg = names(crime_counts), las = 1, cex.names = 0.8, cex.axis = 0.6)
-```
-
-# Figure 4 - Gender Distribution for crimes over time {.unnumbered}
-
-```{r Figure 4 Code, eval=FALSE}
-# Remove missing values and empty strings in Vict.Sex column
-crime_la <- subset(crime_la, Vict.Sex != "" & !is.na(Vict.Sex))
-
-# Replace all non-"F" and non-"M" values with "N"
-crime_la$Vict.Sex <- ifelse(crime_la$Vict.Sex %in% c("F", "M"), 
-                                  crime_la$Vict.Sex, "N")
-
-# Count occurrences of each gender
-gender_counts <- table(crime_la$Vict.Sex)
-
-# Create a pie chart using base R
-color <- c("F" = "pink", "M" = "lightblue", "N" = "purple")
-pie(
-  gender_counts, 
-  labels = paste(names(gender_counts), 
-                 round(gender_counts / sum(gender_counts) * 100, 1), "%"), 
-  col = color[names(gender_counts)],
-  main = "Distribution of Crime Victims by Gender"
-)
-```
-
-# Figure 5 - Median Household Income by Zip Code in Los Angeles by Year {.unnumbered}
-
-```{r Figure 5 Wrangling, eval=FALSE}
-zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
-LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs
-
-colnames(LA_zctas) <- c("X", "ZCTA5CE20")
-LA_zctas$ZCTA5CE20 <- as.character(LA_zctas$ZCTA5CE20)
-
-# joining the zcta shape file with LA zcta reference file
-zcta_filtered <- zcta %>% inner_join(LA_zctas, by = "ZCTA5CE20")
-
-# getting LA median income using joined file
-median_household_income <- cali_median_income %>% select(ZCTA, S1903_C03_001E, year)
-
-zcta_with_data <- zcta_filtered %>% inner_join(median_household_income, join_by("ZCTA5CE20" == "ZCTA"))
-
-# remove NAs
-zcta_no_NAs <- zcta_with_data %>% filter(!is.na(S1903_C03_001E), S1903_C03_001E != "-")
-
-zcta_no_NAs$S1903_C03_001E <- as.integer(zcta_no_NAs$S1903_C03_001E)
-```
-
-```{r Figure 5 Code, eval=FALSE}
-# quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
-qntls <- round(quantile(zcta_no_NAs$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0)
-
-# 2x3 plot layout for years 2018-2023, setting margins and outer margins, main title sizes
-par(mfrow = c(2, 3), mar = c(1, 1, 1, 5), oma = c(0, 0, 3, 0), cex.main = 1.5)
-
-# color palette for sequential data
-palette <- brewer.pal(7, "Greens")
-
-yrs <- 2018:2023
-
-for (yr in yrs) {
-  # get just that year's data
-  subset_data <- zcta_no_NAs[zcta_no_NAs$year == yr, ]
-
-  # creating bins based on the qtls
-  bins <- cut(subset_data$S1903_C03_001E, breaks = qntls, labels = FALSE, include.lowest = TRUE)
-
-  # Plot
-  plot(st_geometry(subset_data),
-       col = palette[bins],
-       border = 'grey',
-       axes = FALSE,
-       lwd = 0.5)
-  title(yr, line = -1)
-
-  # adding legend
-  legend("right", inset = c(-.75, 0), legend = qntls, fill = palette, title = "Median Income ($)", bty = "n", cex = 0.8, xpd = TRUE)
-}
-
-# main title for all plots
-mtext("Median Household Income by Year and Zip Code in LA", outer = TRUE, cex = 1, font = 2)
-
-```
-
-# Figure 6 - Crime Rates Compared to Median Household Income {.unnumbered}
-
-```{r Figure 6 Code, eval=FALSE}
-# using latitude and longitude to obtain the zip codes of the crime locations
-# crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
-# id_lat_long_crime_la <- subset(crime_la, select = c("LAT", "LON", "DR_NO", "year"))
-
-# taking a random sample to reduce load on model
-#set.seed(123)
-#sampled_ids <- id_lat_long_crime_la[sample(1:nrow(id_lat_long_crime_la), 500),]
-#rownames(sampled_ids) <- NULL
-#zip_codes <- reverse_geocode(sampled_ids, lat = LAT, long = LON, verbose = TRUE)
-#write.csv(zip_codes, "sampled_crimes.csv")
-zip_codes <- read.csv("sampled_crimes.csv")
-
-zip_codes$ZCTA5CE20 <- stringr::str_extract(zip_codes$address, "\\d{5}")
-
-crime_income_sample <- zip_codes %>% inner_join(unique(subset(zcta_no_NAs, select = c("ZCTA5CE20", "S1903_C03_001E", "year"))), by = c("ZCTA5CE20", "year"))
-
-# this takes a while, so loading in this csv instead
-
-crime_income_sample <- zip_codes %>% inner_join(unique(subset(zcta_no_NAs, select = c("ZCTA5CE20", "S1903_C03_001E", "year"))), by = c("ZCTA5CE20", "year"))
-
-crime_income_per_zip <- crime_income_sample %>% 
-  group_by(ZCTA5CE20) %>% 
-  summarize(number_of_crimes = n(),
-            cali_median_income = mean(S1903_C03_001E)) %>%
-  ungroup()
-
-plot(x = crime_income_per_zip$cali_median_income, 
-     y = crime_income_per_zip$number_of_crimes, 
-     type = "p",
-     ylab = "Number of Crimes",  
-     xlab = "Median Income ($)",  
-     main = "Crime vs. Median Income by ZIP Code",
-     bty = "l")
-```
-
-## Crime \~ Week Regression Code {.unnumbered}
-
-```{r Regression 1, eval=FALSE}
-library(broom)
-
-weekly_counts$CrimesLag <- c(NA, weekly_counts$Crimes[-nrow(weekly_counts)])
-
-ar1_trend_regression <- lm(Crimes ~ Index + CrimesLag, data = weekly_counts, na.action = na.omit)
-
-ar1_trend_regression %>%
-tidy() %>%
-  mutate(
-    p.value = scales::pvalue(p.value)
-  ) %>%
-  knitr::kable(
-    caption = "Linear Regression Model Results Estimating Crimes Per Week",
-    col.names = c("Predictor", "B", "SE", "t", "p"),
-    digits = c(0, 2, 3, 2, 3),
-    align = c("l", "r", "r", "r", "r")
-  )
-```
-
-## Crime \~ Income Regression Code {.unnumbered}
-
-```{r Regression 2, eval=FALSE}
-lm_income_crimes <- lm(number_of_crimes ~ cali_median_income, data = crime_income_per_zip)
-
-summary(lm_income_crimes)
-
-lm_income_crimes %>%
-tidy() %>%
-  mutate(
-    p.value = scales::pvalue(p.value)
-  ) %>%
-  knitr::kable(
-    caption = "Linear Regression Model Results Estimating Crimes Based on Median Income",
-    col.names = c("Predictor", "B", "SE", "t", "p"),
-    digits = c(0, 2, 3, 2, 3),
-    align = c("l", "r", "r", "r", "r")
-  )
-```

### 2025-03-01 Draft2.qmd
Update Draft2.qmd

@@
 
+```
+
+## Time series analyzing homicide rate trends over time by socioeconomic quartiles
+
+```{r, echo=FALSE}
+# Filter only homicide cases
+homicides <- crime_la %>%
+  inner_join(crimes_with_zips, by = "crime_id") %>%  
+  filter(offense_type == "HOMICIDE") %>%
+  mutate(date = as.Date(date))
+
+# Join homicide data with income data
+homicides <- homicides %>%
+  left_join(cali_median_income, by = c("zip_code" = "ZCTA"))
+
+# Convert median income to numeric if needed
+homicides$median_income <- as.numeric(homicides$median_income)
+
+# Create income quartiles
+homicides <- homicides %>%
+  mutate(income_quartile = ntile(median_income, 4))  # Divides into 4 quartiles
+
+# Aggregate homicide counts by year and income quartile
+homicide_trends <- homicides %>%
+  mutate(year = lubridate::year(date)) %>%
+  group_by(year, income_quartile) %>%
+  summarize(homicide_count = n(), .groups = "drop")
+
+# Plot homicide trends over time by income quartile
+ggplot(homicide_trends, aes(x = year, y = homicide_count, color = factor(income_quartile))) +
+  geom_line(size = 1.2) +
+  geom_point(size = 2) +
+  scale_color_brewer(palette = "Set1") +
+  labs(
+    title = "Homicide Rate Trends Over Time by Socioeconomic Quartiles",
+    x = "Year",
+    y = "Number of Homicides",
+    color = "Income Quartile"
+  ) +
+  theme_minimal() +
+  theme(
+    legend.position = "top",
+    text = element_text(size = 14)
+  )
 ```

### 2025-03-01 Draft2.qmd
Update Draft2.qmd

--- 2025-02-28 Draft2.qmd
@@
 
+```
+
+## Time series analyzing homicide rate trends over time by socioeconomic quartiles
+
+```{r, echo=FALSE}
+# Data Preparation
+homicides <- crime_la %>% inner_join(crimes_with_zips, by = "crime_id") %>% filter(offense_type == "HOMICIDE") %>% mutate(date = as.Date(date)) 
+homicides <- homicides %>% left_join(cali_median_income, by = c("zip_code" = "ZCTA"))
+homicides$median_income <- as.numeric(homicides$median_income)
+
+# Form income quartiles
+homicides <- homicides %>% mutate(income_quartile = ntile(median_income, 4)) 
+homicide_trends <- homicides %>% mutate(year = lubridate::year(date)) %>% group_by(year, income_quartile) %>% summarize(homicide_count = n(), .groups = "drop")
+
+# Plot homicide trends over time by income quartile
+ggplot(homicide_trends, aes(x = year, y = homicide_count, color = factor(income_quartile))) +
+  geom_line(size = 1.2) +
+  geom_point(size = 2) +
+  labs(
+    x = "Year",
+    y = "Number of Homicides",
+    color = "Income Quartile"
+  ) +
+  theme_minimal()
 ```

### 2025-03-01 Draft2.qmd
Added boxplot interpretation

@@
 ```
+###Interpretation
+
+This boxplot shows the number of crimes in different bureaus of Los Angeles.
+We categorized them into Central, South, Valley, and West.
+The South bureau has the highest median number of crimes, as well as the largest variability, while the Central, Valley, and West bureaus have relatively similar distributions with lower crime counts.The outliers suggests some months had significantly lower crime counts than others.
+This visualization helps us understand how crime is distributed across different regions, with the South experiencing the highest fluctuation and overall crime levels.
 

### 2025-03-01 Draft2.qmd
Update Draft2.qmd

@@
 ```
-###Interpretation
+## Interpretation
 
@@
 ```
+
+## Interpretation
+Homicide rates were low pre-pandemic and had a sharp increase since pandemic due to a reduction in income.
+Homicide rate reduces again after pandemic likely due to the economic recovery and increasing wages.
 

### 2025-03-04 2_SecondDraft/Draft2.qmd
Submitted draft version

--- 2025-02-28 Draft2.qmd
@@
 toc: true                   # add table of contents at the beginning
-toc-depth: 3               # Only titles that start with # or ##
+toc-depth: 3               
 geometry: margin=1in
@@
 # Primary datasets: Crime Data 2020-Present and 2010-2019
-crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")
-
-crime_la_2020_present <- read.csv("Raw_Data/Crime_Data_from_2020_to_Present.csv")
-
-crime_la <- read.csv("Cleaned_Data/all_crimes_2010_present.csv")
+crime_la_2010_2019 <- read.csv("0_Data/Raw_Data/Crime_Data_from_2010_to_2019.csv")
+
+crime_la_2020_present <- read.csv("0_Data/Raw_Data/Crime_Data_from_2020_to_Present.csv")
+
+crime_la <- read.csv("0_Data/Cleaned_Data/all_crimes_2010_present.csv")
 
 # primary dataset zip code and geometry data
-crimes_with_zips <- read.csv("Cleaned_Data/crimes_with_zips.csv") # just inner_join with the crime_la dataset
+crimes_with_zips <- read.csv("0_Data/Cleaned_Data/crimes_with_zips.csv") # just inner_join with the crime_la dataset
 
 # Secondary dataset - Median Income by ZCTA
-cali_median_income <- read.csv("Cleaned_Data/cali_median_income.csv")
+cali_median_income <- read.csv("0_Data/Cleaned_Data/cali_median_income.csv")
 
@@
 # Secondary dataset - COVID cases in Los Angeles County
-covid_14day <- read.csv("Raw_Data/LA_County_Covid19_CSA_14day_case_death_table.csv")
-
-evictions_la <- read.csv("Raw_Data/LA_evictions.csv")
+covid_14day <- read.csv("0_Data/Raw_Data/LA_County_Covid19_CSA_14day_case_death_table.csv")
+
+evictions_la <- read.csv("0_Data/Raw_Data/LA_evictions.csv")
 ```
@@
 
-```{r, echo=FALSE, fig.cap = "Crime Occurrences by Day of the Week"}
+```{r, echo=FALSE, fig.cap = "Crime Occurrences by Day of the Week", message=FALSE}
 # Convert Date.Rptd to Date format (ignore time part)
@@
 
-## Median Household Income by Zip Code in Los Angeles by Year
-
-```{r, echo=FALSE}
-zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
-LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs
+## Median Household Income by Year and Zip Code in Los Angeles
+
+```{r, echo=FALSE}
+zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
+LA_zctas <- read.csv("0_Data/Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs
 
@@
 # color palette for sequential data
-palette <- brewer.pal(7, "Greens")
+palette <- brewer.pal(7, "Reds")
 
@@
 
-# main title for all plots
-mtext("Median Household Income by Year and Zip Code in LA", outer = TRUE, cex = 1, font = 2)
-
 ```
@@
 #write.csv(zip_codes, "sampled_crimes.csv")
-zip_codes <- read.csv("sampled_crimes.csv")
+zip_codes <- read.csv("0_Data/Cleaned_Data/sampled_crimes.csv")
 
@@
      xlab = "Median Income ($)",  
-     main = "Crime vs. Median Income by ZIP Code",
      bty = "l")
@@
 
-## Number of Evictions and Crime Levels Per Month
+## Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025
 ```{r, echo=FALSE}
@@
         labels=unique_months$month_year) + 
-  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", title = "Monthly Frequency of Evictions and Crimes in the City of Los Angeles \nFebruary 2023 to February 2025", fill = "Crime Level") +
+  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", fill = "Crime Level") +
   ggthemes::scale_fill_fivethirtyeight() + 
@@
 
-## Boxplot - Kaitlyn
-
-```{r, echo=FALSE}
-crime_by_area <- crime_la %>% group_by(AREA.NAME) %>% summarize(numberOfCrimes = n()) %>% ungroup()
-
-bureaus <- read.csv("Raw_Data/Division_Bureau_Ref.csv")
+## Annual Number of Crimes by Los Angeles Police Department Bureau
+
+```{r, echo=FALSE}
+crime_la$year <- lubridate::year(crime_la$DateParsed)
+crime_by_area <- crime_la %>% group_by(AREA.NAME, year) %>% summarize(numberOfCrimes = n()) %>% ungroup()
+
+bureaus <- read.csv("0_Data/Raw_Data/Division_Bureau_Ref.csv")
 
@@
 ggplot(crime_by_area, aes(y = numberOfCrimes, group = Bureau, x = Bureau)) + 
-        geom_boxplot(aes(fill = Bureau))
-```
-
-## geom_tile for correlations - Bianca
+        geom_boxplot(aes(fill = Bureau)) +
+  labs(y = "Annual Number of Crimes")
+```
+### Interpretation
+
+This boxplot shows the number of crimes in the four different bureaus of Los Angeles annually.
+We categorized the areas into their respective Central, South, Valley, and West Bureaus.
+The South Bureau has the highest median number of crimes, as well as the largest variability, while the Central, Valley, and West bureaus have relatively similar distributions with lower crime counts.The outliers suggests some years had significantly lower crime counts than others.
+This visualization helps us understand how crime is distributed across different regions, with the South experiencing the highest fluctuation and overall crime levels.
+
+## Correlations Between the Number of Crimes and Different Law Enforcement and Socioeconomic Factors
 
@@
 # since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions. 
-LA_county_evictions <- read.csv("Raw_Data/LA_county_evictions.csv")
+LA_county_evictions <- read.csv("0_Data/Raw_Data/LA_county_evictions.csv")
 
@@
 # next, unemployment rate
-unemployment <- read.csv("Raw_Data/la_county_unemployment_rate.csv")
+unemployment <- read.csv("0_Data/Raw_Data/la_county_unemployment_rate.csv")
 unemployment$month <- as.numeric(gsub("M", "", unemployment$Period))
@@
 # lapd budget
-lapd_budget <- read.csv("Raw_Data/lapd_budget.csv")
+lapd_budget <- read.csv("0_Data/Raw_Data/lapd_budget.csv")
 
@@
   geom_tile(color = "white") +
-  scale_fill_viridis_c(direction = -1) +
+  scale_fill_viridis_c(direction = -1,limit = c(-1, 1)) +
   theme(axis.title.x =  element_blank(),
@@
         plot.title = element_text(size=12)) +
-  ggtitle("Correlations Between the Number of Crimes and \n Different Law Enforcement and Socioeconomic Factors")
-```
-
-## geom_jitter
-
-## Plot 5
-
-## Plot 6
+  labs(fill = "Correlation")
+```
+### Interpretation
+This heat map shows the correlations between different variables that could affect the number of crimes committed.
+Due to difficulties in obtaining city-level data, the unemployment rate, COVID cases, and number of evictions are for Los Angeles County, while the crime data is for the City of Los Angeles.
+The city comprises about a third of the county's population and the greater county area largely reflects the trends in the city itself.
+
+Here, we see that the most significant correlations for the number of crimes are the unemployment rate, year, and number of evictions.
+Years is positively correlated with number of crimes, meaning that as time passed, the number of crimes increased.
+Similarly, as the number of evictions increases, so does the number of crimes.
+Interestingly, as the unemployment rate increases, the number of crimes decreases, which is the opposite effect of what was expected (one would anticipate that higher unemployment would result in more crime).
+The LAPD budget interestingly had a medium-strength positive correlation with number of crimes, suggesting we should further examine the increase in the LAPD budget compared to the increase in crimes.
+
+## Homicide Victims by Age and Gender in Los Angeles 2010-Present
+
+```{r, echo=FALSE}
+homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex))
+
+levels(homicide_by_gender$gender) <- c("Female", "Male", "Other")
+ggplot(homicide_by_gender, aes(x = gender, y = Vict.Age, color = gender)) +
+  geom_jitter(alpha = 0.4, width = 0.2) +
+  labs(x = "Gender",
+       y = "Age") +
+  theme_minimal() +
+  scale_color_manual(c("M", "F", "N"), values = c("pink2", "blue", "purple")) + 
+  theme(legend.position = 'none')
+```
+### Interpretation
+
+This jitter plot shows us both the frequencies and the demographics of homicide victims in Los Angeles from 2010 to present.
+Based on the density of the points, we see that most homicide victims in the last 15 years have been men between the ages of around 15 to 65.
+Additionally, we see that there are not many victims who are toddlers and younger children, but there is a good number of babies who have been killed.
+This is an intricacy in our data that might be interesting to look into further, particularly into the circumstances of those deaths and potential correlations between reports of child abuse and/or domestic violence and infanticide.
+
+## Frequency of Crimes in Los Angeles Throughout the Day
+
+```{r, echo=FALSE}
+crime_la$time <- sprintf("%04d",crime_la$TIME.OCC)
+crime_la$Hour <- as.numeric(substr(as.character(crime_la$time), start = 1, stop = 2))
+
+ggplot(crime_la, aes(x = Hour, y = 0, fill = after_stat(density))) +
+  ggridges::geom_density_ridges_gradient(scale = 1, alpha = 0.5) +
+  scale_fill_gradient(low = "lightblue", high = "darkblue") +
+  labs(x = "Hour (Military Time)",
+       y = "Density") +
+  theme_minimal() +
+  theme(legend.position = 'none')
+```
+### Interpretation
+
+This density plot shows the frequency of crimes depending on the hour (in military time) that they occurred.
+Based on the plot, we see that fewer crimes occur early in the morning (with the lowest at 5 AM).
+The data peaks at noon, meaning the most crimes occur in that hour of the day.
+Crimes are more frequent after that, until the early morning.
+The second most common crime occurrence is 6 PM, or 18 hours.
+
+## Time series analyzing homicide rate trends over time by socioeconomic quartiles
+
+```{r, echo=FALSE}
+# Data Preparation
+homicides <- crime_la %>% inner_join(crimes_with_zips, by = c("DR_NO", "LAT", "LON","X", "year")) %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% 
+  mutate(date = as.Date(DateParsed),
+         zip = as.character(ZCTA5CE20)) %>% select(-X)
+
+homicides <- homicides %>% inner_join(cali_median_income, join_by("zip" == "ZCTA", "year" == "year"))
+homicides$S1903_C03_001E <- as.numeric(homicides$S1903_C03_001E)
+
+# Form income quartiles
+homicides <- homicides %>% mutate(income_quartile = ntile(S1903_C03_001E, 4)) 
+homicide_trends <- homicides %>% group_by(year, income_quartile) %>% summarize(homicide_count = n(), 
+          ref = mean(S1903_C03_001E),
+          .groups = "drop")
+
+homicide_trends$income_quartile <- as.factor(homicide_trends$income_quartile)
+# removing NAs
+homicide_trends <- homicide_trends %>% filter(!is.na(homicide_count) & !is.na(income_quartile))
+
+# creating labels
+levels(homicide_trends$income_quartile) <- c("Low Income", "Lower-Middle Income", "Higher-Middle Income", "High Income")
+# Plot homicide trends over time by income quartile
+ggplot(homicide_trends, aes(x = year, y = homicide_count)) +
+  geom_area(aes(fill = as.factor(income_quartile)), alpha = 0.6, position = "identity") +
+  labs(
+    x = "Year",
+    y = "Number of Homicides",
+    fill = "Income Quartile",
+    title = "Homicide Rates in Los Angeles By Median Income Level"
+  ) +
+  theme_minimal() +
+  scale_fill_viridis_d(option = "plasma", direction = -1)
+```
+
+### Interpretation
+Homicide rates were low pre-pandemic and had a sharp increase since the pandemic, except in areas of low income, where homicide rates were high before the pandemic.
+In lower income areas, the rates of homicide have decreased significantly since the end of the pandemic.
+On the other hand, areas of higher-middle and high income have seen large increases in homicide in the six years shown.
 
@@
 
-## Data Wrangling {.unnumbered}
-
-# Reading in Raw Datasets
+#### Data Wrangling {.unnumbered}
+
+##### Reading in Raw Datasets
 
@@
 # Primary datasets: Crime Data 2020-Present and 2010-2019
-crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")
-
-crime_la_2020_present <- read.csv("Raw_Data/Crime_Data_from_2020_to_Present.csv")
+crime_la_2010_2019 <- read.csv("0_Data/Raw_Data/Crime_Data_from_2010_to_2019.csv")
+
+crime_la_2020_present <- read.csv("0_Data/Raw_Data/Crime_Data_from_2020_to_Present.csv")
 
@@
 for (i in 1:length(yrs)) {
-  obj <- read.csv(paste0("Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
+  obj <- read.csv(paste0("0_Data/Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
   obj <- obj[-1,]
@@
 
-### Zip Codes for Crimes {.unnumbered}
+##### Zip Codes for Crimes
 ```{r Get Zip Codes, eval=FALSE}
@@
 
-crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")
-
-crime_la_2020_present <- read.csv("Raw_Data/Crime_Data_from_2020_to_Present.csv")
+crime_la_2010_2019 <- read.csv("0_Data/Raw_Data/Crime_Data_from_2010_to_2019.csv")
+
+crime_la_2020_present <- read.csv("0_Data/Raw_Data/Crime_Data_from_2020_to_Present.csv")
 
@@
 
-zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
+zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
 
@@
 
-  write.csv(result, "crime_zip_files/all_zips.csv")
-
-crime_zctas <- read.csv("Raw_Data/all_zips.csv")
+  write.csv(result, "0_Data/Raw_Data/all_zips.csv")
+
+crime_zctas <- read.csv("0_Data/Raw_Data/all_zips.csv")
 colnames(crime_zctas) <- c(colnames(crime_zctas)[-c(1, length(colnames(crime_zctas)))],"LON", "LAT")
@@
 
-# Figure 1 - Examining Crime Density from 2010–2019 and 2020-Present {.unnumbered}
+#### Figure 1 - Examining Crime Density from 2010–2019 and 2020-Present {.unnumbered}
 
@@
 
-# Figure 2 - Examining Number Of Crimes Over Time {.unnumbered}
+#### Figure 2 - Examining Number Of Crimes Over Time {.unnumbered}
 
@@
 
-# Figure 3 - Crime rate frequency by day of the week {.unnumbered}
+#### Figure 3 - Crime rate frequency by day of the week {.unnumbered}
 
@@
 
-# Figure 4 - Gender Distribution for crimes over time {.unnumbered}
+#### Figure 4 - Gender Distribution for crimes over time {.unnumbered}
 
@@
 
-# Figure 5 - Median Household Income by Zip Code in Los Angeles by Year {.unnumbered}
+#### Figure 5 - Median Household Income by Zip Code in Los Angeles by Year {.unnumbered}
 
 ```{r Figure 5 Wrangling, eval=FALSE}
-zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
-LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs
+zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
+LA_zctas <- read.csv("0_Data/Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs
 
@@
 
-# Figure 6 - Crime Rates Compared to Median Household Income {.unnumbered}
+#### Figure 6 - Crime Rates Compared to Median Household Income {.unnumbered}
 
@@
 #write.csv(zip_codes, "sampled_crimes.csv")
-zip_codes <- read.csv("sampled_crimes.csv")
+zip_codes <- read.csv("0_Data/sampled_crimes.csv")
 
@@
 
-## Crime \~ Week Regression Code {.unnumbered}
+####Figure 7 - Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025 {.unnumbered}
+
+```{r, eval=FALSE}
+# fix mistake
+evictions_la$Date.Received <- gsub("2205", "2025", evictions_la$Date.Received)
+
+# convert to date
+evictions_la$month_year <- format(as.Date(evictions_la$Date.Received, format="%m-%d-%Y"), "%Y-%m")
+
+# get unique months
+unique_months <- data.frame(month_year = sort(unique(evictions_la$month_year)), index = 1:length(unique(evictions_la$month_year)))
+
+evictions_la_indexed <- evictions_la %>% left_join(unique_months, by = "month_year")
+
+# create color scheme
+crime_la$month_year <- format(crime_la$DateParsed, "%Y-%m")
+crime_la_monthly <- crime_la %>% 
+  group_by(month_year) %>% 
+  summarize(num_crimes = n())
+
+evictions_la_indexed_crimes <- evictions_la_indexed %>% left_join(crime_la_monthly, by = "month_year")
+
+avg_crimes <- mean(evictions_la_indexed_crimes$num_crimes)
+
+evictions_la_indexed_crimes$high_low_crime <- as.factor(ifelse(evictions_la_indexed_crimes$num_crimes > avg_crimes, "Above Average", "Below Average"))
+
+ggplot(data = evictions_la_indexed_crimes, aes(x = index)) + 
+  geom_histogram(aes(fill = high_low_crime), bins = nrow(unique_months)) +
+  scale_x_continuous(breaks=unique_months$index,
+        labels=unique_months$month_year) + 
+  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", fill = "Crime Level") +
+  ggthemes::scale_fill_fivethirtyeight() + 
+  ggthemes::theme_stata(scheme = "s1mono") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
+        axis.title.x = element_text(vjust = -.75),
+        plot.title = element_text(size=12))
+```
+
+#### Figure 8 - Annual Number of Crimes by Los Angeles Police Department Bureau {.unnumbered}
+
+```{r, eval=FALSE}
+crime_la$year <- lubridate::year(crime_la$DateParsed)
+crime_by_area <- crime_la %>% group_by(AREA.NAME, year) %>% summarize(numberOfCrimes = n()) %>% ungroup()
+
+bureaus <- read.csv("0_Data/Raw_Data/Division_Bureau_Ref.csv")
+
+bureaus$Division <- gsub("77th", "77th Street", bureaus$Division)
+bureaus$Division <- gsub("North", "N", bureaus$Division)
+bureaus$Division <- gsub("Neast", "Northeast", bureaus$Division)
+
+crime_by_area <- crime_by_area %>% left_join(bureaus, join_by("AREA.NAME" == "Division"))
+
+ggplot(crime_by_area, aes(y = numberOfCrimes, group = Bureau, x = Bureau)) + 
+        geom_boxplot(aes(fill = Bureau)) +
+  labs(y = "Annual Number of Crimes")
+```
+
+#### Figure 9 - Correlations Between the Number of Crimes and Different Law Enforcement and Socioeconomic Factors {.unnumbered}
+
+```{r, eval=FALSE}
+library(lubridate)
+# creating dataset for correlations
+
+# since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions. 
+LA_county_evictions <- read.csv("0_Data/Raw_Data/LA_county_evictions.csv")
+
+# covid data
+covid_14day$date <- ymd(covid_14day$ep_date)
+covid_14day$month <- month(covid_14day$date)
+covid_14day$year <- year(covid_14day$date)
+covid_14day$date <- day(covid_14day$date)
+# since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
+covid_14day$lagged <- covid_14day$date + 7
+
+covid_la_lagged <- covid_14day %>% filter(lagged %in% c(21, 35)) %>% arrange(month, year)
+
+covid_la_monthly <- covid_la_lagged %>% group_by(month, year) %>% summarize(cases = sum(cases_14day)) %>% ungroup() %>% arrange(year, month)
+
+covid_la_monthly$index <- 1:nrow(covid_la_monthly)
+
+# next, unemployment rate
+unemployment <- read.csv("0_Data/Raw_Data/la_county_unemployment_rate.csv")
+unemployment$month <- as.numeric(gsub("M", "", unemployment$Period))
+
+# lapd budget
+lapd_budget <- read.csv("0_Data/Raw_Data/lapd_budget.csv")
+
+crime_la_monthly$date <- ymd(paste0(crime_la_monthly$month_year, "-01"))
+crime_la_monthly$month <- month(crime_la_monthly$date)
+crime_la_monthly$year <- year(crime_la_monthly$date)
+
+correlations_df <- crime_la_monthly %>% inner_join(unemployment, join_by("year" == "Year", "month" == "month")) %>% select(month, year, Value, num_crimes) %>% rename("unemployment_rate" = "Value")
+
+correlations_df2 <- correlations_df %>% inner_join(covid_la_monthly, by = c("month", "year"))
+
+correlations_df3 <- correlations_df2 %>% inner_join(LA_county_evictions, join_by("year" == "Year", "month" == "Month")) %>% rename("num_evictions" = "Count")
+
+correlations_df4 <- correlations_df3 %>% inner_join(lapd_budget, by = "year")
+
+correlations <- correlations_df4 %>% select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)
+
+Cor_mat <- round(cor(correlations), 2)
+
+long_Cor <- reshape2::melt(Cor_mat)
+
+long_Cor <- long_Cor %>%
+    mutate(Var1 = recode(Var1, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget ($ billions)"),
+           Var2 = recode(Var2, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget \n ($ billions)"))
+
+ggplot(data = long_Cor, aes(x = Var1, y = Var2, fill = value)) +
+  geom_tile(color = "white") +
+  scale_fill_viridis_c(direction = -1,limit = c(-1, 1)) +
+  theme(axis.title.x =  element_blank(),
+        axis.title.y = element_blank()) +
+  geom_text(aes(label = value), color = "black", size = 4) + 
+   theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.title = element_text(size=12)) +
+  labs(fill = "Correlation")
+```
+
+#### Figure 10 - Homicide Victims by Age and Gender in Los Angeles 2010-Present {.unnumbered}
+
+```{r, eval=FALSE}
+homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex))
+
+levels(homicide_by_gender$gender) <- c("Female", "Male", "Other")
+ggplot(homicide_by_gender, aes(x = gender, y = Vict.Age, color = gender)) +
+  geom_jitter(alpha = 0.4, width = 0.2) +
+  labs(x = "Gender",
+       y = "Age") +
+  theme_minimal() +
+  scale_color_manual(c("M", "F", "N"), values = c("pink2", "blue", "purple")) + 
+  theme(legend.position = 'none')
+```
+
+#### Figure 11 - Frequency of Crimes in Los Angeles Throughout the Day {.unnumbered}
+
+```{r, eval=FALSE}
+crime_la$time <- sprintf("%04d",crime_la$TIME.OCC)
+crime_la$Hour <- as.numeric(substr(as.character(crime_la$time), start = 1, stop = 2))
+
+ggplot(crime_la, aes(x = Hour, y = 0, fill = after_stat(density))) +
+  ggridges::geom_density_ridges_gradient(scale = 1, alpha = 0.5) +
+  scale_fill_gradient(low = "lightblue", high = "darkblue") +
+  labs(x = "Hour (Military Time)",
+       y = "Density") +
+  theme_minimal() +
+  theme(legend.position = 'none')
+```
+
+#### Figure 12 - Time series analyzing homicide rate trends over time by socioeconomic quartiles {.unnumbered}
+
+```{r, eval=FALSE}
+# Data Preparation
+homicides <- crime_la %>% inner_join(crimes_with_zips, by = c("DR_NO", "LAT", "LON","X", "year")) %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% 
+  mutate(date = as.Date(DateParsed),
+         zip = as.character(ZCTA5CE20)) %>% select(-X)
+
+homicides <- homicides %>% inner_join(cali_median_income, join_by("zip" == "ZCTA", "year" == "year"))
+homicides$S1903_C03_001E <- as.numeric(homicides$S1903_C03_001E)
+
+# Form income quartiles
+homicides <- homicides %>% mutate(income_quartile = ntile(S1903_C03_001E, 4)) 
+homicide_trends <- homicides %>% group_by(year, income_quartile) %>% summarize(homicide_count = n(), 
+          ref = mean(S1903_C03_001E),
+          .groups = "drop")
+
+homicide_trends$income_quartile <- as.factor(homicide_trends$income_quartile)
+# removing NAs
+homicide_trends <- homicide_trends %>% filter(!is.na(homicide_count) & !is.na(income_quartile))
+
+# creating labels
+levels(homicide_trends$income_quartile) <- c("Low Income", "Lower-Middle Income", "Higher-Middle Income", "High Income")
+# Plot homicide trends over time by income quartile
+ggplot(homicide_trends, aes(x = year, y = homicide_count)) +
+  geom_area(aes(fill = as.factor(income_quartile)), alpha = 0.6, position = "identity") +
+  labs(
+    x = "Year",
+    y = "Number of Homicides",
+    fill = "Income Quartile",
+    title = "Homicide Rates in Los Angeles By Median Income Level"
+  ) +
+  theme_minimal() +
+  scale_fill_viridis_d(option = "plasma", direction = -1)
+```
+#### Crime \~ Week Regression Code {.unnumbered}
 
@@
 
-## Crime \~ Income Regression Code {.unnumbered}
+#### Crime \~ Income Regression Code {.unnumbered}
 

### 2025-03-04 zipcodes.R
Updated repo

@@ -1,0 +1,41 @@
+library(sf)
+library(dplyr)
+
+crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")
+
+crime_la_2020_present <- read.csv("Raw_Data/Crime_Data_from_2020_to_Present.csv")
+
+# Combine the two datasets
+crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
+
+crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
+                                  format = "%m/%d/%Y %I:%M:%S %p", 
+                                  tz = "America/Los_Angeles")
+
+crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
+
+crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
+
+coords <- crime_la %>% select(LAT, LON, DR_NO, year)
+
+coords$universal_idx <- 1:nrow(coords)
+
+print(head(coords))
+
+zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
+
+  points <- st_as_sf(coords, coords = c("LON", "LAT"), crs = 4326)  # WGS84
+  
+  print(head(points))
+  
+  zcta <- st_transform(zcta, 4326)  # Convert ZCTA to WGS84 if needed
+  
+  cat("Starting Zip Code Retrieval \n")
+  
+  result <- st_join(points, zcta)
+  
+  print(head(result))
+
+  cat("Finished Zip Code Retrieval \n")
+  
+  write.csv(result, "crime_zip_files/all_zips.csv")

### 2025-03-04 zipcodes.R
Delete zipcodes.R

@@ -1,41 +1,0 @@
-library(sf)
-library(dplyr)
-
-crime_la_2010_2019 <- read.csv("Raw_Data/Crime_Data_from_2010_to_2019.csv")
-
-crime_la_2020_present <- read.csv("Raw_Data/Crime_Data_from_2020_to_Present.csv")
-
-# Combine the two datasets
-crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
-
-crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
-                                  format = "%m/%d/%Y %I:%M:%S %p", 
-                                  tz = "America/Los_Angeles")
-
-crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
-
-crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
-
-coords <- crime_la %>% select(LAT, LON, DR_NO, year)
-
-coords$universal_idx <- 1:nrow(coords)
-
-print(head(coords))
-
-zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
-
-  points <- st_as_sf(coords, coords = c("LON", "LAT"), crs = 4326)  # WGS84
-  
-  print(head(points))
-  
-  zcta <- st_transform(zcta, 4326)  # Convert ZCTA to WGS84 if needed
-  
-  cat("Starting Zip Code Retrieval \n")
-  
-  result <- st_join(points, zcta)
-  
-  print(head(result))
-
-  cat("Finished Zip Code Retrieval \n")
-  
-  write.csv(result, "crime_zip_files/all_zips.csv")

### 2025-03-04 Draft2.qmd
organization

--- 2025-03-01 Draft2.qmd
@@
 
+```{r, setup, include=FALSE}
+knitr::opts_knit$set(root.dir = "..")
+```
+
 ```{r DATA, output=FALSE, echo=FALSE}
@@
         plot.title = element_text(size=12)) +
-  ggtitle("Correlations Between the Number of Crimes and \n Different Law Enforcement and Socioeconomic Factors")
-```
-
-## geom_jitter
-## geom_jitter
-
-```{r, echo=FALSE}
-ggplot(crime_la, aes(x = DateParsed, y = Vict.Sex, color = Vict.Sex)) +
+  labs(fill = "Correlation")
+```
+### Interpretation
+This heat map shows the correlations between different variables that could affect the number of crimes committed.
+Due to difficulties in obtaining city-level data, the unemployment rate, COVID cases, and number of evictions are for Los Angeles County, while the crime data is for the City of Los Angeles.
+The city comprises about a third of the county's population and the greater county area largely reflects the trends in the city itself.
+
+Here, we see that the most significant correlations for the number of crimes are the unemployment rate, year, and number of evictions.
+Years is positively correlated with number of crimes, meaning that as time passed, the number of crimes increased.
+Similarly, as the number of evictions increases, so does the number of crimes.
+Interestingly, as the unemployment rate increases, the number of crimes decreases, which is the opposite effect of what was expected (one would anticipate that higher unemployment would result in more crime).
+The LAPD budget interestingly had a medium-strength positive correlation with number of crimes, suggesting we should further examine the increase in the LAPD budget compared to the increase in crimes.
+
+## Homicide Victims by Age and Gender in Los Angeles 2010-Present
+
+```{r, echo=FALSE}
+homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex))
+
+levels(homicide_by_gender$gender) <- c("Female", "Male", "Other")
+ggplot(homicide_by_gender, aes(x = gender, y = Vict.Age, color = gender)) +
   geom_jitter(alpha = 0.4, width = 0.2) +
-  labs(title = "Crime Over Time by Gender",
-       x = "Date",
-       y = "Gender") +
-  theme_minimal()
-
-```
-
-## Density Plot
-
-```{r, echo=FALSE}
-
-crime_la$Hour <- as.numeric(format(crime_la$DateParsed, "%H"))
-
-ggplot(crime_la, aes(x = Hour, fill = ..density..)) +
-  geom_density(alpha = 0.5, color = "black") +
+  labs(x = "Gender",
+       y = "Age") +
+  theme_minimal() +
+  scale_color_manual(c("M", "F", "N"), values = c("pink2", "blue", "purple")) + 
+  theme(legend.position = 'none')
+```
+### Interpretation
+
+This jitter plot shows us both the frequencies and the demographics of homicide victims in Los Angeles from 2010 to present.
+Based on the density of the points, we see that most homicide victims in the last 15 years have been men between the ages of around 15 to 65.
+Additionally, we see that there are not many victims who are toddlers and younger children, but there is a good number of babies who have been killed.
+This is an intricacy in our data that might be interesting to look into further, particularly into the circumstances of those deaths and potential correlations between reports of child abuse and/or domestic violence and infanticide.
+
+## Frequency of Crimes in Los Angeles Throughout the Day
+
+```{r, echo=FALSE}
+crime_la$time <- sprintf("%04d",crime_la$TIME.OCC)
+crime_la$Hour <- as.numeric(substr(as.character(crime_la$time), start = 1, stop = 2))
+
+ggplot(crime_la, aes(x = Hour, y = 0, fill = after_stat(density))) +
+  ggridges::geom_density_ridges_gradient(scale = 1, alpha = 0.5) +
   scale_fill_gradient(low = "lightblue", high = "darkblue") +
-  labs(title = "Crime Density Per Hour",
-       x = "Hour",
+  labs(x = "Hour (Military Time)",
        y = "Density") +
-  theme_minimal()
-
-```
+  theme_minimal() +
+  theme(legend.position = 'none')
+```
+### Interpretation
+
+This density plot shows the frequency of crimes depending on the hour (in military time) that they occurred.
+Based on the plot, we see that fewer crimes occur early in the morning (with the lowest at 5 AM).
+The data peaks at noon, meaning the most crimes occur in that hour of the day.
+Crimes are more frequent after that, until the early morning.
+The second most common crime occurrence is 6 PM, or 18 hours.
 
@@
 # Data Preparation
-homicides <- crime_la %>% inner_join(crimes_with_zips, by = "crime_id") %>% filter(offense_type == "HOMICIDE") %>% mutate(date = as.Date(date)) 
-homicides <- homicides %>% left_join(cali_median_income, by = c("zip_code" = "ZCTA"))
-homicides$median_income <- as.numeric(homicides$median_income)
+homicides <- crime_la %>% inner_join(crimes_with_zips, by = c("DR_NO", "LAT", "LON","X", "year")) %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% 
+  mutate(date = as.Date(DateParsed),
+         zip = as.character(ZCTA5CE20)) %>% select(-X)
+
+homicides <- homicides %>% inner_join(cali_median_income, join_by("zip" == "ZCTA", "year" == "year"))
+homicides$S1903_C03_001E <- as.numeric(homicides$S1903_C03_001E)
 
 # Form income quartiles
-homicides <- homicides %>% mutate(income_quartile = ntile(median_income, 4)) 
-homicide_trends <- homicides %>% mutate(year = lubridate::year(date)) %>% group_by(year, income_quartile) %>% summarize(homicide_count = n(), .groups = "drop")
-
+homicides <- homicides %>% mutate(income_quartile = ntile(S1903_C03_001E, 4)) 
+homicide_trends <- homicides %>% group_by(year, income_quartile) %>% summarize(homicide_count = n(), 
+          ref = mean(S1903_C03_001E),
+          .groups = "drop")
+
+homicide_trends$income_quartile <- as.factor(homicide_trends$income_quartile)
+# removing NAs
+homicide_trends <- homicide_trends %>% filter(!is.na(homicide_count) & !is.na(income_quartile))
+
+# creating labels
+levels(homicide_trends$income_quartile) <- c("Low Income", "Lower-Middle Income", "Higher-Middle Income", "High Income")
 # Plot homicide trends over time by income quartile
-ggplot(homicide_trends, aes(x = year, y = homicide_count, color = factor(income_quartile))) +
-  geom_line(size = 1.2) +
-  geom_point(size = 2) +
+ggplot(homicide_trends, aes(x = year, y = homicide_count)) +
+  geom_area(aes(fill = as.factor(income_quartile)), alpha = 0.6, position = "identity") +
   labs(
@@
     y = "Number of Homicides",
-    color = "Income Quartile"
+    fill = "Income Quartile",
+    title = "Homicide Rates in Los Angeles By Median Income Level"
   ) +
-  theme_minimal()
-```
-
-## Plot 6
+  theme_minimal() +
+  scale_fill_viridis_d(option = "plasma", direction = -1)
+```
+
+### Interpretation
+Homicide rates were low pre-pandemic and had a sharp increase since the pandemic, except in areas of low income, where homicide rates were high before the pandemic.
+In lower income areas, the rates of homicide have decreased significantly since the end of the pandemic.
+On the other hand, areas of higher-middle and high income have seen large increases in homicide in the six years shown.
 
@@
 
+# Spatio-Temporal Analysis of Crime in Los Angeles
+
+```{r}
+crimes_with_zips <- read.csv("0_Data/Cleaned_Data/crimes_with_zips.csv")
+
+```
+
 # References {.unnumbered}

### 2025-03-04 Draft2.qmd
hyperlinks

@@
 
-The codebook for this dataset can be found here: https://data.lacity.org/Public-Safety/Crime-Data-from-2020-to-Present/2nrs-mtv8/about_data.
+The codebook for this dataset can be found here: <https://data.lacity.org/Public-Safety/Crime-Data-from-2020-to-Present/2nrs-mtv8/about_data>.
 
@@
 ```
+
 ### Interpretation
@@
 ## Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025
+
 ```{r, echo=FALSE}
@@
 ```
-### Interpretation
+
+### Interpretation
+
 This histogram plots the frequency of evictions in Los Angeles from February 2023 to February 2025, where the City of Los Angeles Housing Department has published case data.
@@
 ```
-### Interpretation
+
+### Interpretation
+
 This heat map shows the correlations between different variables that could affect the number of crimes committed.
@@
 ### Interpretation
+
 Homicide rates were low pre-pandemic and had a sharp increase since the pandemic, except in areas of low income, where homicide rates were high before the pandemic.
@@
 # Data Sources {.unnumbered}
+
 ## Primary Datasets {.unnumbered}
-Crime Data From 2010-2019: https://data.lacity.org/Public-Safety/Crime-Data-from-2010-to-2019/63jg-8b9z/about_data
-
-Crime Data From 2020-Present: https://catalog.data.gov/dataset/crime-data-from-2020-to-present
+
+Crime Data From 2010-2019: <https://data.lacity.org/Public-Safety/Crime-Data-from-2010-to-2019/63jg-8b9z/about_data>
+
+Crime Data From 2020-Present: <https://catalog.data.gov/dataset/crime-data-from-2020-to-present>
 
 ## Secondary Datasets {.unnumbered}
-Los Angeles County Eviction Data: https://calmatters.org/housing/homelessness/2023/11/california-evictions-post-pandemic/
-
-City of Los Angeles Eviction Data: https://housing.lacity.gov/residents/renters/eviction-notices-filed
-
-Median Income Data: https://data.census.gov/table/ACSST1Y2023.S1903?q=income
+
+Los Angeles County Eviction Data: <https://calmatters.org/housing/homelessness/2023/11/california-evictions-post-pandemic/>
+
+City of Los Angeles Eviction Data: <https://housing.lacity.gov/residents/renters/eviction-notices-filed>
+
+Median Income Data: <https://data.census.gov/table/ACSST1Y2023.S1903?q=income>
 
@@
 ##### Zip Codes for Crimes
+
 ```{r Get Zip Codes, eval=FALSE}
@@
 ```
+
 #### Crime \~ Week Regression Code {.unnumbered}

### 2025-03-05 Draft3.qmd
Adding Draft 3 WIP

--- 2025-03-04 Draft2.qmd
@@
 
+crime_la$month <- lubridate::month(crime_la$DateParsed)
+
 crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
@@
 
-# Set locale to English (for weekday names)
-Sys.setlocale("LC_TIME", "C")
-
 # Extract the day of the week
@@
         xlab = "Day of the Week", ylab = "Number of Crimes",
-        names.arg = names(crime_counts), las = 1, cex.names = 0.8, cex.axis = 0.6)
+        names.arg = names(crime_counts), las = 1, cex.names = 0.4, cex.axis = 0.6)
 ```
@@
 
-## Regression Between Median Income and Crimes per Zip Code
+## Regression Between Median Income and Crimes per Zip Code - DROP THIS
 
@@
 
-# Spatio-Temporal Analysis of Crime in Los Angeles
+## Chi-Square for Zip Code
 
 ```{r}
-crimes_with_zips <- read.csv("0_Data/Cleaned_Data/crimes_with_zips.csv")
-
+homicides_by_zip <- homicides %>% group_by(ZCTA5CE20) %>% summarize(HomicideCount = n()) %>% ungroup()
+chisq.test(x = as.character(homicides_by_zip$ZCTA5CE20), y = homicides_by_zip$HomicideCount)
+```
+
+## Model Progression
+### Initial Spatial Analysis of Crime in Los Angeles
+
+```{r, echo=FALSE, warning=FALSE}
+homicides_latlon_rounded <- homicides %>% mutate(LAT_rounded = round(LAT, 1),
+                                                 LON_rounded = round(LON, 1))
+homicides_lat_long <- homicides_latlon_rounded %>% group_by(LAT_rounded, LON_rounded) %>% summarize(HomicideCount = n()) %>% ungroup()
+
+linear_model1 <- lm(HomicideCount ~ LAT_rounded + LON_rounded, data = homicides_by_year_lat_long)
+
+summary(linear_model1) # longitude rounded is significant, LAT_rounded is not
+```
+
+### By Year
+```{r}
+homicides_lat_long <- homicides_latlon_rounded %>% group_by(LAT_rounded, LON_rounded) %>% summarize(HomicideCount = n()) %>% ungroup()
+
+linear_model2 <- lm(HomicideCount ~ LAT_rounded + LON_rounded + year, data = homicides_by_year_lat_long)
+
+summary(linear_model2) # longitude rounded is significant, LAT_rounded is not. Neither is year
 ```

### 2025-03-06 Draft3.qmd
Updated draft

@@
 
-# Regression Analyses
+# Statistical Analyses
+
+To examine the relationship between the number of homicides and various socioeconomic and enforcement factors, we will use a variety of statistical analyses to see if certain factors are more influential in causing increases or decreases in homicide rates.
+
+For the purposes of this analysis, homicide rate is quantified as the annual number of homicides in a particular entity, whether that is ZIP Code Tabulation Area (ZCTA) or Census Tract.
+Additionally, we will examine the influence of general geographic location through the latitude and longitude where the homicide was committed.
 
@@
 
-### Interpretation
+### Discussion
 
@@
 
-## Chi-Square for Zip Code
-
-```{r}
-homicides_by_zip <- homicides %>% group_by(ZCTA5CE20) %>% summarize(HomicideCount = n()) %>% ungroup()
-chisq.test(x = as.character(homicides_by_zip$ZCTA5CE20), y = homicides_by_zip$HomicideCount)
-```
+## Negative Binomial for Census Block
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE}
+library(kableExtra)
+crimes_by_cb <- read.csv("0_Data/Cleaned_Data/crimes_by_census_block.csv")
+colnames(crimes_by_cb) <- c(colnames(crimes_by_cb)[-c(1, length(colnames(crimes_by_cb)))],"LON", "LAT")
+
+crimes_by_cb$LON <- as.numeric(gsub('c(', "", crimes_by_cb$LON, fixed = TRUE))
+crimes_by_cb$LAT <- as.numeric(gsub(')', "", crimes_by_cb$LAT, fixed = TRUE))
+
+crimes_by_cb$DateParsed <- as.POSIXct(crimes_by_cb$Date.Rptd, 
+                                  format = "%m/%d/%Y %I:%M:%S %p", 
+                                  tz = "America/Los_Angeles")
+
+crimes_by_cb$YearWeek <- format(crimes_by_cb$DateParsed, "%Y-%U")
+
+crimes_by_cb$year <- as.integer(stringr::str_sub(crimes_by_cb$YearWeek, 1, 4))
+
+homicides_by_cb <- crimes_by_cb %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% group_by(BLKGRPCE, year) %>% summarize(HomicideCount = n()) %>% ungroup()
+
+# shapiro.test(homicides_by_cb$HomicideCount) # cannot use parametric tests. Strong evidence against normality.
+
+mean_homicides <- mean(homicides_by_cb$HomicideCount)
+var_homicides <- var(homicides_by_cb$HomicideCount)
+# since variance > mean, use negative binomial 
+library(MASS)
+model <- glm.nb(HomicideCount ~ as.factor(BLKGRPCE), data = homicides_by_cb)
+
+summary(model)
+```
+### Discussion
 
@@
 crime_zctas$LAT <- as.numeric(gsub(')', "", crime_zctas$LAT, fixed = TRUE))
+
+# Repeat same process for Census Blocks
 ```

### 2025-03-07 Draft3.qmd
Added background map to Plot 1

@@
 evictions_la <- read.csv("0_Data/Raw_Data/LA_evictions.csv")
+
+crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
+                                  format = "%m/%d/%Y %I:%M:%S %p", 
+                                  tz = "America/Los_Angeles")
+
+crime_la$month <- lubridate::month(crime_la$DateParsed)
+
+crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
+
+crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
 ```
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+la_shape <- read_sf(dsn = "0_Data/Raw_Data/ca_counties", layer = "ca_counties")
+
+la_shape <- st_transform(la_shape, crs = 4326)
 
 par(mfrow = c(1, 2), mar = c(4, 4, 2, 1))
+
+plot(la_shape$geometry,xlim = c(-119, -118), ylim = c(33.5, 34.5),
+     xlab = "Longitude",
+  ylab = "Latitude",
+  main = "Crime Density (2010-2019)",
+  cex.main=0.8)
 
@@
   colramp = colorRampPalette(c("white", "skyblue", "blue", "darkblue")),
-  cex.main=0.8
+  cex.main=0.8,
+  add = TRUE
 )
+
+plot(la_shape$geometry,xlim = c(-119, -118), ylim = c(33.5, 34.5),
+     xlab = "Longitude",
+  ylab = "Latitude",
+  main = "Crime Density (2020–Present)",
+  cex.main=0.8)
 
@@
   colramp = colorRampPalette(c("white", "pink", "red", "darkred")),
-  cex.main=0.8
+  cex.main=0.8, 
+  add = TRUE
 )

### 2025-03-07 Draft3.qmd
Added background map to Plot 1

--- 2025-03-06 Draft3.qmd
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+la_shape <- read_sf(dsn = "0_Data/Raw_Data/Cities_and_Communities", layer = "Cities_and_Communities")
+
+la_shape <- st_transform(la_shape, crs = 4326)
+
+cali <- read_sf(dsn = "0_Data/Raw_Data/ca_counties", layer = "ca_counties")
+
+cali <- st_transform(cali, crs = 4326)
 
@@
 
+plot(la_shape$geometry, lwd=.5, add=TRUE)
+
+plot(cali$geometry, add = TRUE, lwd=.5)
+
 smoothScatter(
@@
 )
+
+plot(la_shape$geometry, lwd=.5, add=TRUE)
+
+plot(cali$geometry, add = TRUE, lwd=.5)
 

### 2025-03-07 Draft3.qmd
Got rid of warnings and messages in embedded code blocks w warning and message FALSE

@@
 # Exploratory Data Analysis
+
+[ADD MAIN EDA TAKEAWAYS HERE]
 
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
@@
 
-```{r, echo=FALSE, fig.cap = "Crime Occurrences by Day of the Week", message=FALSE}
+```{r, echo=FALSE, fig.cap = "Crime Occurrences by Day of the Week", warning=FALSE, message=FALSE}
 # Convert Date.Rptd to Date format (ignore time part)
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 # Remove missing values and empty strings in Vict.Sex column
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 # quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 # using latitude and longitude to obtain the zip codes of the crime locations
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 # fix mistake
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 crime_la$year <- lubridate::year(crime_la$DateParsed)
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 library(lubridate)
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex))
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 crime_la$time <- sprintf("%04d",crime_la$TIME.OCC)
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 # Data Preparation
@@
 
+[BASED ON THESE PLOTS, WE WILL DO XYZ TO EXAMINE ABC]
+
 # Statistical Analyses
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 library(broom)
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 lm_income_crimes <- lm(number_of_crimes ~ cali_median_income, data = crime_income_per_zip)
@@
 
-```{r, echo=FALSE, warning=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 homicides_latlon_rounded <- homicides %>% mutate(LAT_rounded = round(LAT, 1),
@@
 ### By Year
-```{r}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 homicides_lat_long <- homicides_latlon_rounded %>% group_by(LAT_rounded, LON_rounded) %>% summarize(HomicideCount = n()) %>% ungroup()

### 2025-03-07 Draft3.qmd
Updated code and intro

@@
 
+## Data Description
+
+### Primary Data: Los Angeles Crime Data
+
 We will utilize NIBRS data from 2018-2024 to examine incident-level data, allowing us to fully capture the circumstances of each event.
@@
 
-```{r, setup, include=FALSE}
-knitr::opts_knit$set(root.dir = "..")
-```
+To group this data geographically, we will organize the crimes by their latitude and longitude into the appropriate Census Blocks and ZIP Code Tabulation Areas (ZCTAs).
+Additionally, we will examine the impact of latitude and longitude themselves, as humans do not always behave by administrative unit boundaries and latitude and longitude by themselves might suggest a different pattern.[@krieger2002]
+
+### Secondary Data: Geographic and Socioeconomic Data
+
+The NIBRS data includes two main geographic identifiers: address and coordinates.
+Some addresses are poorly coded, such as just listing a street and no number, so we instead used latitude and longitude to examine geospatial relationships.
+We also used these coordinates to group our data into administrative units (Census Blocks and ZCTAs) to perform different levels of geospatial analysis.
+
+We used shapefiles for the state of California, Los Angeles Police Department Divisions, and the Census units to categorize each crime event and to visualize the spatial distribution of the crimes through maps.
+To categorize the crime coordinate data, we performed a spatial join where we converted the crime coordinates to a spatial object and then joined them based on the geometries of the shape files.
+
+We additionally used publicly available socioeconomic data from the US Census Bureau, the City of Los Angeles Housing Department (LAHD), CalMatters, and Los Angeles County Public Health.
+We used median household income data from the American Community Survey 5-Year estimates to get an annual measure of income in LA ZCTAs.
+We used evictions data from both Calmatters and LAHD to gather longer-term data and more recent data, as LAHD only started tracking evictions data in 2023, while Calmatters has tracked LA County data since 2012.
+
+Due to difficulties in obtaining city-level data, the unemployment rate, COVID cases, and number of evictions are for Los Angeles County, while the crime data is for the City of Los Angeles.
+The city comprises about a third of the county's population and the greater county area largely reflects the trends in the city itself.
+
+To maximize the amount of data available for analysis, we used LA County socioeconomic data from 2019-
+
+## Hypothesis
+
+Based on the 2022 Thurgood Marshall Institute report, we anticipate that the homicide rates have decreased since the end of the pandemic as society has returned to a more "normal" state.
+We aim to see how the rate of homicide compares to overall crime levels in the city and identify any patterns between socioeconomic factors such as eviction rates and the homicide count.
+
+Our hypothesis is two-fold; first, we predict that since the pandemic, that socioeconomic conditions have improved in Los Angeles over the past few years.
+Second, we predict that as these socioeconomic measures have bounced back, the rate of homicide has returned to a more "normal" rate.
+
+Additionally, we believe that we will observe a similar relationship between homicide and the overall crime rate; as the number of crimes decreases, so will the number of homicides.
+
+We will also investigate the relationship between homicide and enforcement factors.
+In particular, we will look at the LAPD budget in comparison to the homicide rate to determine if increased police funding is effective in reducing the number of homicides.
 
@@
 library(tidygeocoder) # to get zip codes of addresses
+library(dplyr)
 
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
-la_shape <- read_sf(dsn = "0_Data/Raw_Data/Cities_and_Communities", layer = "Cities_and_Communities")
-
-la_shape <- st_transform(la_shape, crs = 4326)
+lapd_shape <- read_sf(dsn = "0_Data/Raw_Data/LAPD_Divisions", layer = "LAPD_Divisions")
+
+lapd_shape <- st_transform(lapd_shape, crs = 4326)
 
@@
   ylab = "Latitude",
-  main = "Crime Density (2010–2019)",
+  main = "Crime Density by LAPD Division (2010–2019)",
   colramp = colorRampPalette(c("white", "skyblue", "blue", "darkblue")),
@@
 
-plot(la_shape$geometry, lwd=.5, add=TRUE)
+plot(lapd_shape$geometry, lwd=.5, add=TRUE)
 
@@
   ylab = "Latitude",
-  main = "Crime Density (2020–Present)",
+  main = "Crime Density by LAPD Division (2020–Present)",
   colramp = colorRampPalette(c("white", "pink", "red", "darkred")),
@@
 # getting LA median income using joined file
-median_household_income <- cali_median_income %>% select(ZCTA, S1903_C03_001E, year)
+median_household_income <- cali_median_income %>% dplyr::select(ZCTA, S1903_C03_001E, year)
 
@@
 
-correlations_df <- crime_la_monthly %>% inner_join(unemployment, join_by("year" == "Year", "month" == "month")) %>% select(month, year, Value, num_crimes) %>% rename("unemployment_rate" = "Value")
+correlations_df <- crime_la_monthly %>% inner_join(unemployment, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, Value, num_crimes) %>% rename("unemployment_rate" = "Value")
 
@@
 
-correlations <- correlations_df4 %>% select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)
+correlations <- correlations_df4 %>% dplyr::select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)
 
@@
 This heat map shows the correlations between different variables that could affect the number of crimes committed.
-Due to difficulties in obtaining city-level data, the unemployment rate, COVID cases, and number of evictions are for Los Angeles County, while the crime data is for the City of Los Angeles.
-The city comprises about a third of the county's population and the greater county area largely reflects the trends in the city itself.
-
-Here, we see that the most significant correlations for the number of crimes are the unemployment rate, year, and number of evictions.
+We see that the most significant correlations for the number of crimes are the unemployment rate, year, and number of evictions.
 Years is positively correlated with number of crimes, meaning that as time passed, the number of crimes increased.
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
-homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex))
+homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% dplyr::select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex))
 
@@
   mutate(date = as.Date(DateParsed),
-         zip = as.character(ZCTA5CE20)) %>% select(-X)
+         zip = as.character(ZCTA5CE20)) %>% dplyr::select(-X)
 
@@
 
-linear_model1 <- lm(HomicideCount ~ LAT_rounded + LON_rounded, data = homicides_by_year_lat_long)
+linear_model1 <- lm(HomicideCount ~ LAT_rounded + LON_rounded, data = homicides_lat_long)
 
@@
 ### By Year
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
 homicides_lat_long <- homicides_latlon_rounded %>% group_by(LAT_rounded, LON_rounded) %>% summarize(HomicideCount = n()) %>% ungroup()
@@
 
-coords <- crime_la %>% select(LAT, LON, DR_NO, year)
+coords <- crime_la %>% dplyr::select(LAT, LON, DR_NO, year)
 
@@
 
-crime_zctas_no_duplicates <- crime_zctas %>% mutate(duplicate = ifelse(universal_idx %in% duplicate_data$universal_idx & ZCTA5CE20 != 91325, "Drop", "Keep")) %>% filter(duplicate != "Drop") %>% select(-duplicate)
+crime_zctas_no_duplicates <- crime_zctas %>% mutate(duplicate = ifelse(universal_idx %in% duplicate_data$universal_idx & ZCTA5CE20 != 91325, "Drop", "Keep")) %>% filter(duplicate != "Drop") %>% dplyr::select(-duplicate)
 
@@
 
-no_zipcode_not0_zips <- no_zipcode_not0_zips %>% select(-address)
+no_zipcode_not0_zips <- no_zipcode_not0_zips %>% dplyr::select(-address)
 
@@
 # getting LA median income using joined file
-median_household_income <- cali_median_income %>% select(ZCTA, S1903_C03_001E, year)
+median_household_income <- cali_median_income %>% dplyr::select(ZCTA5CE20, S1903_C03_001E, year)
 

### 2025-03-07 Draft3.qmd
Update Draft3.qmd

@@
 
+The plot analyzes the number of crimes over time using time series.
 The time series shows stable weekly crime counts with occasional large spikes.
@@
 
-The number of crimes are rather evenly distributed, with the weekends having slightly lower crime compared to week days.
+The plot analyzes the crime rate frequency frequency by day of the week using a bar graph.
+The number of crimes are rather evenly distributed, with the weekends having slightly lower crime compared to week days, which may be a result from more people around during weekends which indirectly suprevises people's behavior.
 
@@
 
-From the pie chart, we can see that around 50% of the crime is commited by male, 45% by female and 5% by non-binary.
+The pie chart analyzes the gender distribution of crime activities.
+From the chart, we can infer that around 50% of the crime is commited by male, 45% by female and 5% by non-binary.
+The crime distribution by gender is pretty even
 

### 2025-03-08 Draft3.qmd
Submitted version

--- 2025-03-07 Draft3.qmd
@@
 
-author: "Marco Stine, Kevin Hu, Bianca Schutz, and Kaitlyn Kirt"
+author: "Bianca Schutz, Kaitlyn Kirt, Kevin Hu, and Marco Stine"
 fontsize: 10pt
@@
 
-[ADD MAIN EDA TAKEAWAYS HERE]
-
-## Examining Crime Density from 2010–2019 and 2020-Present
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
+```{r echo=FALSE, fig.cap="Crime densities in Los Angeles by LAPD Division, message=FALSE, warning=FALSE, 2010.2019 and 2020.Present", fig.align="center"}
 lapd_shape <- read_sf(dsn = "0_Data/Raw_Data/LAPD_Divisions", layer = "LAPD_Divisions")
@@
 cali <- st_transform(cali, crs = 4326)
-
-par(mfrow = c(1, 2), mar = c(4, 4, 2, 1))
-
-smoothScatter(
-  x    = crime_la_2010_2019$LON,
-  y    = crime_la_2010_2019$LAT,
-  xlim = c(-119, -118),
-  ylim = c(33.5, 34.5),
-  nbin = 300,
-  xlab = "Longitude",
-  ylab = "Latitude",
-  main = "Crime Density by LAPD Division (2010–2019)",
-  colramp = colorRampPalette(c("white", "skyblue", "blue", "darkblue")),
-  cex.main=0.8
-)
-
-plot(lapd_shape$geometry, lwd=.5, add=TRUE)
-
-plot(cali$geometry, add = TRUE, lwd=.5)
-
-smoothScatter(
-  x    = crime_la_2020_present$LON,
-  y    = crime_la_2020_present$LAT,
-  xlim = c(-119, -118),
-  ylim = c(33.5, 34.5),
-  nbin = 300,
-  xlab = "Longitude",
-  ylab = "Latitude",
-  main = "Crime Density by LAPD Division (2020–Present)",
-  colramp = colorRampPalette(c("white", "pink", "red", "darkred")),
-  cex.main=0.8
-)
-
-plot(lapd_shape$geometry, lwd=.5, add=TRUE)
-
-plot(cali$geometry, add = TRUE, lwd=.5)
-
-par(mfrow = c(1, 1))
-
-```
-
-### Interpretation
-
-Both time periods show major clusters of reported crimes in central LA. The 2020–present data appears slightly more dispersed to the north, but overall patterns remain similar looking.
-
-## Examining Number Of Crimes Over Time
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
-                                  format = "%m/%d/%Y %I:%M:%S %p", 
-                                  tz = "America/Los_Angeles")
-
-crime_la$month <- lubridate::month(crime_la$DateParsed)
-
-crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
-
-weekly_counts <- aggregate(DR_NO ~ YearWeek, data = crime_la, FUN = length)
-names(weekly_counts)[2] <- "Crimes"
-
-weekly_counts <- weekly_counts[order(weekly_counts$YearWeek), ]
-
-weekly_counts$Index <- seq_len(nrow(weekly_counts))
-
-plot(
-  weekly_counts$Index,
-  weekly_counts$Crimes,
-  type = "l",
-  col  = "blue",
-  lwd  = 2,
-  xlab = "Weekly Index",
-  ylab = "Number of Crimes",
-  main = "Weekly Crime Counts in Los Angeles (2010–Present)"
-)
-
-abline(h = pretty(weekly_counts$Crimes), v = pretty(weekly_counts$Index), 
-       col = "gray", lty = "dotted")
-```
-
-### Interpretation
-
-The time series shows stable weekly crime counts with occasional large spikes.
-These surges indicate certain weeks where crime reporting jumped significantly, there is a recent interesting surge that seems to mean-revert in 2024.
-
-## Crime rate frequency by day of the week
-
-```{r, echo=FALSE, fig.cap = "Crime Occurrences by Day of the Week", warning=FALSE, message=FALSE}
-# Convert Date.Rptd to Date format (ignore time part)
-#crime_la$Date.Rptd
-crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")
-
-# Extract the day of the week
-crime_la$DayOfWeek <- weekdays(crime_la$Date)
-# crime_la$DayOfWeek
-
-# Count occurrences of crimes per day
-crime_counts <- table(crime_la$DayOfWeek)
-# crime_counts
-
-# Order the days correctly
-day_order <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
-crime_counts <- crime_counts[day_order]
-
-# Plot bar chart with horizontal labels
-barplot(crime_counts, col = "lightblue",
-        xlab = "Day of the Week", ylab = "Number of Crimes",
-        names.arg = names(crime_counts), las = 1, cex.names = 0.4, cex.axis = 0.6)
-```
-
-### Interpretation
-
-The number of crimes are rather evenly distributed, with the weekends having slightly lower crime compared to week days.
-
-## Gender Distribution for crimes over time
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-# Remove missing values and empty strings in Vict.Sex column
-crime_la <- subset(crime_la, Vict.Sex != "" & !is.na(Vict.Sex))
-
-# Replace all non-"F" and non-"M" values with "N"
-crime_la$Vict.Sex <- ifelse(crime_la$Vict.Sex %in% c("F", "M"), 
-                                  crime_la$Vict.Sex, "N")
-
-# Count occurrences of each gender
-gender_counts <- table(crime_la$Vict.Sex)
-
-# Create a pie chart using base R
-color <- c("F" = "pink", "M" = "lightblue", "N" = "purple")
-pie(
-  gender_counts, 
-  labels = paste(names(gender_counts), 
-                 round(gender_counts / sum(gender_counts) * 100, 1), "%"), 
-  col = color[names(gender_counts)],
-  main = "Distribution of Crime Victims by Gender"
-)
-```
-
-### Interpretation
-
-From the pie chart, we can see that around 50% of the crime is commited by male, 45% by female and 5% by non-binary.
-
-## Median Household Income by Year and Zip Code in Los Angeles
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
-LA_zctas <- read.csv("0_Data/Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs
-
-colnames(LA_zctas) <- c("X", "ZCTA5CE20")
-LA_zctas$ZCTA5CE20 <- as.character(LA_zctas$ZCTA5CE20)
-
-# joining the zcta shape file with LA zcta reference file
-zcta_filtered <- zcta %>% inner_join(LA_zctas, by = "ZCTA5CE20")
-
-# getting LA median income using joined file
-median_household_income <- cali_median_income %>% dplyr::select(ZCTA, S1903_C03_001E, year)
-
-zcta_with_data <- zcta_filtered %>% inner_join(median_household_income, join_by("ZCTA5CE20" == "ZCTA"))
-
-# remove NAs
-zcta_no_NAs <- zcta_with_data %>% filter(!is.na(S1903_C03_001E), S1903_C03_001E != "-")
-
-zcta_no_NAs$S1903_C03_001E <- as.integer(zcta_no_NAs$S1903_C03_001E)
-```
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-# quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
-qntls <- round(quantile(zcta_no_NAs$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0)
-
-# 2x3 plot layout for years 2018-2023, setting margins and outer margins, main title sizes
-par(mfrow = c(2, 3), mar = c(1, 1, 1, 5), oma = c(0, 0, 3, 0), cex.main = 1.5)
-
-# color palette for sequential data
-palette <- brewer.pal(7, "Reds")
-
-yrs <- 2018:2023
-
-for (yr in yrs) {
-  # get just that year's data
-  subset_data <- zcta_no_NAs[zcta_no_NAs$year == yr, ]
-
-  # creating bins based on the qtls
-  bins <- cut(subset_data$S1903_C03_001E, breaks = qntls, labels = FALSE, include.lowest = TRUE)
-
-  # Plot
-  plot(st_geometry(subset_data),
-       col = palette[bins],
-       border = 'grey',
-       axes = FALSE,
-       lwd = 0.5)
-  title(yr, line = -1)
-
-  # adding legend
-  legend("right", inset = c(-.75, 0), legend = qntls, fill = palette, title = "Median Income ($)", bty = "n", cex = 0.8, xpd = TRUE)
-}
-
-```
-
-### Interpretation
-
-The maps demonstrate that median income in certain LA zip codes (particularly those to the west) have a much higher median household income than the rest of the city, but over the last six years, the median income in most other LA zip codes has increased.
-However, there are some that a typical family is at or below the poverty line.
-We intend to investigate if there is a spatial correlation between these poorer zip codes and the incidence of crimes, particularly homicide, and whether this is affected by pandemic conditions.
-
-## Crime Rates Compared to Median Household Income
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-# using latitude and longitude to obtain the zip codes of the crime locations
-# crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
-# id_lat_long_crime_la <- subset(crime_la, select = c("LAT", "LON", "DR_NO", "year"))
-
-# taking a random sample to reduce load on model
-#set.seed(123)
-#sampled_ids <- id_lat_long_crime_la[sample(1:nrow(id_lat_long_crime_la), 500),]
-#rownames(sampled_ids) <- NULL
-#zip_codes <- reverse_geocode(sampled_ids, lat = LAT, long = LON, verbose = TRUE)
-#write.csv(zip_codes, "sampled_crimes.csv")
-zip_codes <- read.csv("0_Data/Cleaned_Data/sampled_crimes.csv")
-
-zip_codes$ZCTA5CE20 <- stringr::str_extract(zip_codes$address, "\\d{5}")
-
-crime_income_sample <- zip_codes %>% inner_join(unique(subset(zcta_no_NAs, select = c("ZCTA5CE20", "S1903_C03_001E", "year"))), by = c("ZCTA5CE20", "year"))
-
-# this takes a while, so loading in this csv instead
-
-crime_income_sample <- zip_codes %>% inner_join(unique(subset(zcta_no_NAs, select = c("ZCTA5CE20", "S1903_C03_001E", "year"))), by = c("ZCTA5CE20", "year"))
-
-crime_income_per_zip <- crime_income_sample %>% 
-  group_by(ZCTA5CE20) %>% 
-  summarize(number_of_crimes = n(),
-            cali_median_income = mean(S1903_C03_001E)) %>%
-  ungroup()
-
-plot(x = crime_income_per_zip$cali_median_income, 
-     y = crime_income_per_zip$number_of_crimes, 
-     type = "p",
-     ylab = "Number of Crimes",  
-     xlab = "Median Income ($)",  
-     bty = "l")
-```
-
-### Interpretation
-
-Based on this plot, we do not observe a clear trend between the number of crimes committed and the median income in that zip code.
-However, this scatterplot only accounts for a small portion of the dataset and averages across pre- and post-pandemic conditions.
-Future visualizations and analyses could consider pandemic vs non-pandemic years as variables.
-
-# Exploratory Data Analysis With ggplot2
-
-## Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-# fix mistake
-evictions_la$Date.Received <- gsub("2205", "2025", evictions_la$Date.Received)
-
-# convert to date
-evictions_la$month_year <- format(as.Date(evictions_la$Date.Received, format="%m-%d-%Y"), "%Y-%m")
-
-# get unique months
-unique_months <- data.frame(month_year = sort(unique(evictions_la$month_year)), index = 1:length(unique(evictions_la$month_year)))
-
-evictions_la_indexed <- evictions_la %>% left_join(unique_months, by = "month_year")
-
-# create color scheme
-crime_la$month_year <- format(crime_la$DateParsed, "%Y-%m")
-crime_la_monthly <- crime_la %>% 
-  group_by(month_year) %>% 
-  summarize(num_crimes = n())
-
-evictions_la_indexed_crimes <- evictions_la_indexed %>% left_join(crime_la_monthly, by = "month_year")
-
-avg_crimes <- mean(evictions_la_indexed_crimes$num_crimes)
-
-evictions_la_indexed_crimes$high_low_crime <- as.factor(ifelse(evictions_la_indexed_crimes$num_crimes > avg_crimes, "Above Average", "Below Average"))
-
-ggplot(data = evictions_la_indexed_crimes, aes(x = index)) + 
-  geom_histogram(aes(fill = high_low_crime), bins = nrow(unique_months)) +
-  scale_x_continuous(breaks=unique_months$index,
-        labels=unique_months$month_year) + 
-  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", fill = "Crime Level") +
-  ggthemes::scale_fill_fivethirtyeight() + 
-  ggthemes::theme_stata(scheme = "s1mono") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
-        axis.title.x = element_text(vjust = -.75),
-        plot.title = element_text(size=12))
-```
-
-### Interpretation
-
-This histogram plots the frequency of evictions in Los Angeles from February 2023 to February 2025, where the City of Los Angeles Housing Department has published case data.
-By summarizing both the crime and the evictions datasets by month, we see that the number of evictions was highest in early-mid 2023, and since, there have typically been slightly lower frequencies of evictions.
-Additionally, we split the number of crimes per month based on the average number of crimes of 14,439 to denote above and below average number of crimes per month.
-There is a clear split between March and April of 2024, as all months before that had above average crime, and all after had below average.
-Interestingly, the months where there is below average crime, seem to be the ones with slightly lower numbers of evictions.
-
-## Annual Number of Crimes by Los Angeles Police Department Bureau
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-crime_la$year <- lubridate::year(crime_la$DateParsed)
-crime_by_area <- crime_la %>% group_by(AREA.NAME, year) %>% summarize(numberOfCrimes = n()) %>% ungroup()
-
-bureaus <- read.csv("0_Data/Raw_Data/Division_Bureau_Ref.csv")
-
-bureaus$Division <- gsub("77th", "77th Street", bureaus$Division)
-bureaus$Division <- gsub("North", "N", bureaus$Division)
-bureaus$Division <- gsub("Neast", "Northeast", bureaus$Division)
-
-crime_by_area <- crime_by_area %>% left_join(bureaus, join_by("AREA.NAME" == "Division"))
-
-ggplot(crime_by_area, aes(y = numberOfCrimes, group = Bureau, x = Bureau)) + 
-        geom_boxplot(aes(fill = Bureau)) +
-  labs(y = "Annual Number of Crimes")
-```
-
-### Interpretation
-
-This boxplot shows the number of crimes in the four different bureaus of Los Angeles annually.
-We categorized the areas into their respective Central, South, Valley, and West Bureaus.
-The South Bureau has the highest median number of crimes, as well as the largest variability, while the Central, Valley, and West bureaus have relatively similar distributions with lower crime counts.The outliers suggests some years had significantly lower crime counts than others.
-This visualization helps us understand how crime is distributed across different regions, with the South experiencing the highest fluctuation and overall crime levels.
-
-## Correlations Between the Number of Crimes and Different Law Enforcement and Socioeconomic Factors
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-library(lubridate)
-# creating dataset for correlations
-
-# since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions. 
-LA_county_evictions <- read.csv("0_Data/Raw_Data/LA_county_evictions.csv")
-
-# covid data
-covid_14day$date <- ymd(covid_14day$ep_date)
-covid_14day$month <- month(covid_14day$date)
-covid_14day$year <- year(covid_14day$date)
-covid_14day$date <- day(covid_14day$date)
-# since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
-covid_14day$lagged <- covid_14day$date + 7
-
-covid_la_lagged <- covid_14day %>% filter(lagged %in% c(21, 35)) %>% arrange(month, year)
-
-covid_la_monthly <- covid_la_lagged %>% group_by(month, year) %>% summarize(cases = sum(cases_14day)) %>% ungroup() %>% arrange(year, month)
-
-covid_la_monthly$index <- 1:nrow(covid_la_monthly)
-
-# next, unemployment rate
-unemployment <- read.csv("0_Data/Raw_Data/la_county_unemployment_rate.csv")
-unemployment$month <- as.numeric(gsub("M", "", unemployment$Period))
-
-# lapd budget
-lapd_budget <- read.csv("0_Data/Raw_Data/lapd_budget.csv")
-
-crime_la_monthly$date <- ymd(paste0(crime_la_monthly$month_year, "-01"))
-crime_la_monthly$month <- month(crime_la_monthly$date)
-crime_la_monthly$year <- year(crime_la_monthly$date)
-
-correlations_df <- crime_la_monthly %>% inner_join(unemployment, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, Value, num_crimes) %>% rename("unemployment_rate" = "Value")
-
-correlations_df2 <- correlations_df %>% inner_join(covid_la_monthly, by = c("month", "year"))
-
-correlations_df3 <- correlations_df2 %>% inner_join(LA_county_evictions, join_by("year" == "Year", "month" == "Month")) %>% rename("num_evictions" = "Count")
-
-correlations_df4 <- correlations_df3 %>% inner_join(lapd_budget, by = "year")
-
-correlations <- correlations_df4 %>% dplyr::select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)
-
-Cor_mat <- round(cor(correlations), 2)
-
-long_Cor <- reshape2::melt(Cor_mat)
-
-long_Cor <- long_Cor %>%
-    mutate(Var1 = recode(Var1, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget ($ billions)"),
-           Var2 = recode(Var2, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget \n ($ billions)"))
-
-ggplot(data = long_Cor, aes(x = Var1, y = Var2, fill = value)) +
-  geom_tile(color = "white") +
-  scale_fill_viridis_c(direction = -1,limit = c(-1, 1)) +
-  theme(axis.title.x =  element_blank(),
-        axis.title.y = element_blank()) +
-  geom_text(aes(label = value), color = "black", size = 4) + 
-   theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        plot.title = element_text(size=12)) +
-  labs(fill = "Correlation")
-```
-
-### Interpretation
-
-This heat map shows the correlations between different variables that could affect the number of crimes committed.
-We see that the most significant correlations for the number of crimes are the unemployment rate, year, and number of evictions.
-Years is positively correlated with number of crimes, meaning that as time passed, the number of crimes increased.
-Similarly, as the number of evictions increases, so does the number of crimes.
-Interestingly, as the unemployment rate increases, the number of crimes decreases, which is the opposite effect of what was expected (one would anticipate that higher unemployment would result in more crime).
-The LAPD budget interestingly had a medium-strength positive correlation with number of crimes, suggesting we should further examine the increase in the LAPD budget compared to the increase in crimes.
-
-## Homicide Victims by Age and Gender in Los Angeles 2010-Present
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% dplyr::select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex))
-
-levels(homicide_by_gender$gender) <- c("Female", "Male", "Other")
-ggplot(homicide_by_gender, aes(x = gender, y = Vict.Age, color = gender)) +
-  geom_jitter(alpha = 0.4, width = 0.2) +
-  labs(x = "Gender",
-       y = "Age") +
-  theme_minimal() +
-  scale_color_manual(c("M", "F", "N"), values = c("pink2", "blue", "purple")) + 
-  theme(legend.position = 'none')
-```
-
-### Interpretation
-
-This jitter plot shows us both the frequencies and the demographics of homicide victims in Los Angeles from 2010 to present.
-Based on the density of the points, we see that most homicide victims in the last 15 years have been men between the ages of around 15 to 65.
-Additionally, we see that there are not many victims who are toddlers and younger children, but there is a good number of babies who have been killed.
-This is an intricacy in our data that might be interesting to look into further, particularly into the circumstances of those deaths and potential correlations between reports of child abuse and/or domestic violence and infanticide.
-
-## Frequency of Crimes in Los Angeles Throughout the Day
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-crime_la$time <- sprintf("%04d",crime_la$TIME.OCC)
-crime_la$Hour <- as.numeric(substr(as.character(crime_la$time), start = 1, stop = 2))
-
-ggplot(crime_la, aes(x = Hour, y = 0, fill = after_stat(density))) +
-  ggridges::geom_density_ridges_gradient(scale = 1, alpha = 0.5) +
-  scale_fill_gradient(low = "lightblue", high = "darkblue") +
-  labs(x = "Hour (Military Time)",
-       y = "Density") +
-  theme_minimal() +
-  theme(legend.position = 'none')
-```
-
-### Interpretation
-
-This density plot shows the frequency of crimes depending on the hour (in military time) that they occurred.
-Based on the plot, we see that fewer crimes occur early in the morning (with the lowest at 5 AM).
-The data peaks at noon, meaning the most crimes occur in that hour of the day.
-Crimes are more frequent after that, until the early morning.
-The second most common crime occurrence is 6 PM, or 18 hours.
-
-## Time series analyzing homicide rate trends over time by socioeconomic quartiles
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-# Data Preparation
-homicides <- crime_la %>% inner_join(crimes_with_zips, by = c("DR_NO", "LAT", "LON","X", "year")) %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% 
-  mutate(date = as.Date(DateParsed),
-         zip = as.character(ZCTA5CE20)) %>% dplyr::select(-X)
-
-homicides <- homicides %>% inner_join(cali_median_income, join_by("zip" == "ZCTA", "year" == "year"))
-homicides$S1903_C03_001E <- as.numeric(homicides$S1903_C03_001E)
-
-# Form income quartiles
-homicides <- homicides %>% mutate(income_quartile = ntile(S1903_C03_001E, 4)) 
-homicide_trends <- homicides %>% group_by(year, income_quartile) %>% summarize(homicide_count = n(), 
-          ref = mean(S1903_C03_001E),
-          .groups = "drop")
-
-homicide_trends$income_quartile <- as.factor(homicide_trends$income_quartile)
-# removing NAs
-homicide_trends <- homicide_trends %>% filter(!is.na(homicide_count) & !is.na(income_quartile))
-
-# creating labels
-levels(homicide_trends$income_quartile) <- c("Low Income", "Lower-Middle Income", "Higher-Middle Income", "High Income")
-# Plot homicide trends over time by income quartile
-ggplot(homicide_trends, aes(x = year, y = homicide_count)) +
-  geom_area(aes(fill = as.factor(income_quartile)), alpha = 0.6, position = "identity") +
-  labs(
-    x = "Year",
-    y = "Number of Homicides",
-    fill = "Income Quartile",
-    title = "Homicide Rates in Los Angeles By Median Income Level"
-  ) +
-  theme_minimal() +
-  scale_fill_viridis_d(option = "plasma", direction = -1)
-```
-
-### Interpretation
-
-Homicide rates were low pre-pandemic and had a sharp increase since the pandemic, except in areas of low income, where homicide rates were high before the pandemic.
-In lower income areas, the rates of homicide have decreased significantly since the end of the pandemic.
-On the other hand, areas of higher-middle and high income have seen large increases in homicide in the six years shown.
-
-[BASED ON THESE PLOTS, WE WILL DO XYZ TO EXAMINE ABC]
-
-# Statistical Analyses
-
-To examine the relationship between the number of homicides and various socioeconomic and enforcement factors, we will use a variety of statistical analyses to see if certain factors are more influential in causing increases or decreases in homicide rates.
-
-For the purposes of this analysis, homicide rate is quantified as the annual number of homicides in a particular entity, whether that is ZIP Code Tabulation Area (ZCTA) or Census Tract.
-Additionally, we will examine the influence of general geographic location through the latitude and longitude where the homicide was committed.
-
-## Regression to Examine Weekly Crime Rates
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-library(broom)
-
-weekly_counts$CrimesLag <- c(NA, weekly_counts$Crimes[-nrow(weekly_counts)])
-
-ar1_trend_regression <- lm(Crimes ~ Index + CrimesLag, data = weekly_counts, na.action = na.omit)
-
-ar1_trend_regression %>%
-tidy() %>%
-  mutate(
-    p.value = scales::pvalue(p.value)
-  ) %>%
-  knitr::kable(
-    caption = "Linear Regression Model Results Estimating Crimes Per Week",
-    col.names = c("Predictor", "B", "SE", "t", "p"),
-    digits = c(0, 2, 3, 2, 3),
-    align = c("l", "r", "r", "r", "r")
-  )
-```
-
-### Discussion
-
-The lagged crime coefficient (0.81949) suggests that a high crime week is often followed by another high crime week.
-Since the trend term is not significant, there is no clear drift in weekly crime over time.
-
-## Regression Between Median Income and Crimes per Zip Code - DROP THIS
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-lm_income_crimes <- lm(number_of_crimes ~ cali_median_income, data = crime_income_per_zip)
-
-# summary(lm_income_crimes)
-
-lm_income_crimes %>%
-tidy() %>%
-  mutate(
-    p.value = scales::pvalue(p.value)
-  ) %>%
-  knitr::kable(
-    caption = "Linear Regression Model Results Estimating Crimes Based on Median Income",
-    col.names = c("Predictor", "B", "SE", "t", "p"),
-    digits = c(0, 2, 3, 2, 3),
-    align = c("l", "r", "r", "r", "r")
-  )
-```
-
-### Discussion
-
-Based on the p-value, we observe that median income is not significant at a 0.05 level, meaning it does not significantly decrease the amount of crimes committed.
-The r\^2 value additionally tells us that the income only explains a small portion of the variation in the number of crimes.
-As stated with the scatterplot, a larger sample and considering other factors might yield different results.
-
-## Negative Binomial for Census Block
-
-```{r, warning=FALSE, message=FALSE, echo=FALSE}
-library(kableExtra)
-crimes_by_cb <- read.csv("0_Data/Cleaned_Data/crimes_by_census_block.csv")
-colnames(crimes_by_cb) <- c(colnames(crimes_by_cb)[-c(1, length(colnames(crimes_by_cb)))],"LON", "LAT")
-
-crimes_by_cb$LON <- as.numeric(gsub('c(', "", crimes_by_cb$LON, fixed = TRUE))
-crimes_by_cb$LAT <- as.numeric(gsub(')', "", crimes_by_cb$LAT, fixed = TRUE))
-
-crimes_by_cb$DateParsed <- as.POSIXct(crimes_by_cb$Date.Rptd, 
-                                  format = "%m/%d/%Y %I:%M:%S %p", 
-                                  tz = "America/Los_Angeles")
-
-crimes_by_cb$YearWeek <- format(crimes_by_cb$DateParsed, "%Y-%U")
-
-crimes_by_cb$year <- as.integer(stringr::str_sub(crimes_by_cb$YearWeek, 1, 4))
-
-homicides_by_cb <- crimes_by_cb %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% group_by(BLKGRPCE, year) %>% summarize(HomicideCount = n()) %>% ungroup()
-
-# shapiro.test(homicides_by_cb$HomicideCount) # cannot use parametric tests. Strong evidence against normality.
-
-mean_homicides <- mean(homicides_by_cb$HomicideCount)
-var_homicides <- var(homicides_by_cb$HomicideCount)
-# since variance > mean, use negative binomial 
-library(MASS)
-model <- glm.nb(HomicideCount ~ as.factor(BLKGRPCE), data = homicides_by_cb)
-
-summary(model)
-```
-### Discussion
-
-## Model Progression
-### Initial Spatial Analysis of Crime in Los Angeles
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
-homicides_latlon_rounded <- homicides %>% mutate(LAT_rounded = round(LAT, 1),
-                                                 LON_rounded = round(LON, 1))
-homicides_lat_long <- homicides_latlon_rounded %>% group_by(LAT_rounded, LON_rounded) %>% summarize(HomicideCount = n()) %>% ungroup()
-
-linear_model1 <- lm(HomicideCount ~ LAT_rounded + LON_rounded, data = homicides_lat_long)
-
-summary(linear_model1) # longitude rounded is significant, LAT_rounded is not
-```
-
-### By Year
-```{r, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
-homicides_lat_long <- homicides_latlon_rounded %>% group_by(LAT_rounded, LON_rounded) %>% summarize(HomicideCount = n()) %>% ungroup()
-
-linear_model2 <- lm(HomicideCount ~ LAT_rounded + LON_rounded + year, data = homicides_by_year_lat_long)
-
-summary(linear_model2) # longitude rounded is significant, LAT_rounded is not. Neither is year
-```
-
-# References {.unnumbered}
-
-::: {#refs}
-:::
-
-# Data Sources {.unnumbered}
-
-## Primary Datasets {.unnumbered}
-
-Crime Data From 2010-2019: <https://data.lacity.org/Public-Safety/Crime-Data-from-2010-to-2019/63jg-8b9z/about_data>
-
-Crime Data From 2020-Present: <https://catalog.data.gov/dataset/crime-data-from-2020-to-present>
-
-## Secondary Datasets {.unnumbered}
-
-Los Angeles County Eviction Data: <https://calmatters.org/housing/homelessness/2023/11/california-evictions-post-pandemic/>
-
-City of Los Angeles Eviction Data: <https://housing.lacity.gov/residents/renters/eviction-notices-filed>
-
-Median Income Data: <https://data.census.gov/table/ACSST1Y2023.S1903?q=income>
-
-# Appendices {.appendix}
-
-#### Data Wrangling {.unnumbered}
-
-##### Reading in Raw Datasets
-
-```{r Wrangling Main, eval=FALSE}
-library(sf) # for map
-library(tidyverse) # using only for joining datasets to build map visual
-library(RColorBrewer) # for map color scheme
-library(tidygeocoder) # to get zip codes of addresses
-
-# Primary datasets: Crime Data 2020-Present and 2010-2019
-crime_la_2010_2019 <- read.csv("0_Data/Raw_Data/Crime_Data_from_2010_to_2019.csv")
-
-crime_la_2020_present <- read.csv("0_Data/Raw_Data/Crime_Data_from_2020_to_Present.csv")
-
-# Combine the two datasets
-crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
-```
-
-```{r Combining Median Income, eval=FALSE}
-# Secondary datasets
-california_median_income_2018_2023 <- list()
-yrs <- 2018:2023
-for (i in 1:length(yrs)) {
-  obj <- read.csv(paste0("0_Data/Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
-  obj <- obj[-1,]
-  obj$year <- yrs[[i]]
-  california_median_income_2018_2023[[i]] <- obj
-}
-
-cali_median_income <- do.call(rbind, california_median_income_2018_2023)
-
-# adding a variable that's just the ZCTA zip code
-cali_median_income$ZCTA <- stringr::str_sub(cali_median_income$NAME, -5)
-
-write.csv(cali_median_income, "cali_median_income.csv")
-```
-
-##### Zip Codes for Crimes
-
-```{r Get Zip Codes, eval=FALSE}
-library(sf)
-library(dplyr)
-
-crime_la_2010_2019 <- read.csv("0_Data/Raw_Data/Crime_Data_from_2010_to_2019.csv")
-
-crime_la_2020_present <- read.csv("0_Data/Raw_Data/Crime_Data_from_2020_to_Present.csv")
-
-# Combine the two datasets
-crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
-
-crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
-                                  format = "%m/%d/%Y %I:%M:%S %p", 
-                                  tz = "America/Los_Angeles")
-
-crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
-
-crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
-
-coords <- crime_la %>% dplyr::select(LAT, LON, DR_NO, year)
-
-zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
-
-  points <- st_as_sf(coords, coords = c("LON", "LAT"), crs = 4326)  # WGS84
-
-  zcta <- st_transform(zcta, 4326)  # Convert ZCTA to WGS84 if needed
-  cat("Starting Zip Code Retrieval \n")
-
-  result <- st_join(points, zcta)
-
-  cat("Finished Zip Code Retrieval \n")
-
-  write.csv(result, "0_Data/Raw_Data/all_zips.csv")
-
-crime_zctas <- read.csv("0_Data/Raw_Data/all_zips.csv")
-colnames(crime_zctas) <- c(colnames(crime_zctas)[-c(1, length(colnames(crime_zctas)))],"LON", "LAT")
-
-crime_zctas$LON <- as.numeric(gsub('c(', "", crime_zctas$LON, fixed = TRUE))
-crime_zctas$LAT <- as.numeric(gsub(')', "", crime_zctas$LAT, fixed = TRUE))
-
-# Repeat same process for Census Blocks
-```
-
-```{r Get Zip Codes 2, eval=FALSE}
-
-duplicates <- as.integer(as.character(data.frame(table(crime_zctas$universal_idx)[table(crime_zctas$universal_idx) > 1]) %>% pull(Var1)))
-
-# there are approximately 200 rows that are duplicated (100 incidents) as the area lies on the intersection between two zctas. 
-duplicate_data <- crime_zctas %>% filter(universal_idx %in% duplicates)
-
-rev_geo_duplicates <- reverse_geocode(duplicates2, lat = LAT, long = LON)
-# the above query identifies the address of these coordinates as KFC, 8644, Balboa Boulevard, Northridge South Neighborhood Council District, Los Angeles, Los Angeles County, California, 91325, United States.
-# therefore, we will only keep the 91325 zip code rows.
-
-crime_zctas_no_duplicates <- crime_zctas %>% mutate(duplicate = ifelse(universal_idx %in% duplicate_data$universal_idx & ZCTA5CE20 != 91325, "Drop", "Keep")) %>% filter(duplicate != "Drop") %>% dplyr::select(-duplicate)
-
-# it's empty now 
-# duplicates <- table(crime_zctas_no_duplicates$universal_idx)[table(crime_zctas_no_duplicates$universal_idx) > 1]
-
-# next, check for NAs
-sum(is.na(crime_zctas_no_duplicates$ZCTA5CE20)) # there are 3724 NAs. Let's find them a zip code
-
-# some also have lat long = 0. We'll find those separately
-no_zipcode_not0 <- crime_zctas_no_duplicates %>% filter(is.na(ZCTA5CE20) & LON != 0)
-
-no_zipcode_not0_zips <- reverse_geocode(no_zipcode_not0, lat = LAT, long = LON)
-
-no_zipcode_not0_zips$ZCTA5CE20 <- stringr::str_extract(no_zipcode_not0_zips$address, "\\d{5}")
-
-no_zipcode_not0_zips <- no_zipcode_not0_zips %>% dplyr::select(-address)
-
-crime_zctas2 <- crime_zctas_no_duplicates %>% filter(!is.na(ZCTA5CE20))
-
-crime_zctas3 <- rbind(crime_zctas2, no_zipcode_not0_zips)
-
-crime_zctas4 <- crime_zctas3 %>% filter(!is.na(ZCTA5CE20))
-
-write.csv(crime_zctas4, "crimes_with_zips.csv")
-
-no_zipcode_0s <- crime_zctas_no_duplicates %>% filter(is.na(ZCTA5CE20) & LON == 0) %>% pull(DR_NO)
-
-main_dataset_0s <- crime_la %>% filter(DR_NO %in% no_zipcode_0s)
-
-summary(main_dataset_0s)
-
-summary(crime_la)
-```
-
-#### Figure 1 - Examining Crime Density from 2010–2019 and 2020-Present {.unnumbered}
-
-```{r Figure 1 Code, eval=FALSE}
 
@@
 
+plot(lapd_shape$geometry, lwd=.5, add=TRUE)
+
+plot(cali$geometry, add = TRUE, lwd=.5)
+
 smoothScatter(
@@
 
+plot(lapd_shape$geometry, lwd=.5, add=TRUE)
+
+plot(cali$geometry, add = TRUE, lwd=.5)
+
 par(mfrow = c(1, 1))
@@
 
-#### Figure 2 - Examining Number Of Crimes Over Time {.unnumbered}
-
-```{r Figure 2 Code, eval=FALSE}
+Figure 1 shows that both time periods show major clusters of reported crimes in central LA.
+The 2020–present data appears slightly more dispersed to the north, but overall, the density of crimes has remained the same over time.
+The West LA and Foothill LAPD divisions seem to have the lowest density of crimes.
+West LA Division has crimes mostly located in the south of its area.
+Beverly Hills, which is not one of the LAPD divisions and has its own police force, explains the whiter spot in central LA.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap = "Weekly Crime Counts in Los Angeles from 2010 to present", fig.align = "center"}
 crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
@@
                                   tz = "America/Los_Angeles")
+
+crime_la$month <- lubridate::month(crime_la$DateParsed)
 
@@
   xlab = "Weekly Index",
-  ylab = "Number of Crimes",
-  main = "Weekly Crime Counts in Los Angeles (2010–Present)"
-)
+  ylab = "Number of Crimes")
 
@@
 
-#### Figure 3 - Crime rate frequency by day of the week {.unnumbered}
-
-```{r Figure 3 Code, eval=FALSE}
+Figure 2 analyzes the number of crimes over time using time series.
+The time series shows stable weekly crime counts with occasional large spikes.
+These surges indicate certain weeks where crime reporting jumped significantly, there is a recent interesting surge that seems to mean-revert in 2024.
+
+```{r, echo=FALSE, fig.cap = "Los Angeles Crimes by Day of the Week", warning=FALSE, message=FALSE, fig.align = "center"}
 # Convert Date.Rptd to Date format (ignore time part)
@@
 day_order <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
-crime_counts <- crime_counts[day_order]
+crime_counts <- data.frame(crime_counts[day_order])
+colnames(crime_counts) <- c("Day", "CrimeCount")
+
+crime_counts$thousands <- crime_counts$CrimeCount/1000
 
 # Plot bar chart with horizontal labels
-barplot(crime_counts, col = "lightblue",
-        xlab = "Day of the Week", ylab = "Number of Crimes",
-        names.arg = names(crime_counts), las = 1, cex.names = 0.8, cex.axis = 0.6)
-```
-
-#### Figure 4 - Gender Distribution for crimes over time {.unnumbered}
-
-```{r Figure 4 Code, eval=FALSE}
+ggplot(crime_counts, aes(x = Day, y = thousands)) +
+  geom_col(fill = "blue4") +
+  labs(x = "Day of the Week",
+       y = "Number of Crimes (thousands)")
+```
+
+In Figure 3, the number of crimes are rather evenly distributed, with the weekends having slightly lower crime compared to week days, which may be a result from more people around during weekends which indirectly supervises people's behavior.
+Most crimes occur on Mondays, and the fewest crimes occur on Sunday.
+The plot analyzes the crime rate frequency frequency by day of the week using a bar graph.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Percent of crime victims by gender, 2010 to present", fig.align = "center"}
 # Remove missing values and empty strings in Vict.Sex column
@@
                  round(gender_counts / sum(gender_counts) * 100, 1), "%"), 
-  col = color[names(gender_counts)],
-  main = "Distribution of Crime Victims by Gender"
+  col = color[names(gender_counts)]
 )
@@
 
-#### Figure 5 - Median Household Income by Zip Code in Los Angeles by Year {.unnumbered}
-
-```{r Figure 5 Wrangling, eval=FALSE}
+From the pie chart shown in Figure 4, we can see that around 50% of crime victims are male, 45% are female and 5% are non-binary.
+This shows that the distribution of crime victims by gender is relatively even, though men do account for a plurality of the victims.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
 zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
@@
 # getting LA median income using joined file
-median_household_income <- cali_median_income %>% dplyr::select(ZCTA5CE20, S1903_C03_001E, year)
+median_household_income <- cali_median_income %>% dplyr::select(ZCTA, S1903_C03_001E, year)
 
@@
 
-```{r Figure 5 Code, eval=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Median household income by year and by ZIP Code Tabulation Area (ZCTA) in Los Angeles", fig.align = "center"}
 # quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
@@
 
-# main title for all plots
-mtext("Median Household Income by Year and Zip Code in LA", outer = TRUE, cex = 1, font = 2)
-
-```
-
-#### Figure 6 - Crime Rates Compared to Median Household Income {.unnumbered}
-
-```{r Figure 6 Code, eval=FALSE}
-# using latitude and longitude to obtain the zip codes of the crime locations
-# crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
-# id_lat_long_crime_la <- subset(crime_la, select = c("LAT", "LON", "DR_NO", "year"))
-
-# taking a random sample to reduce load on model
-#set.seed(123)
-#sampled_ids <- id_lat_long_crime_la[sample(1:nrow(id_lat_long_crime_la), 500),]
-#rownames(sampled_ids) <- NULL
-#zip_codes <- reverse_geocode(sampled_ids, lat = LAT, long = LON, verbose = TRUE)
-#write.csv(zip_codes, "sampled_crimes.csv")
-zip_codes <- read.csv("0_Data/sampled_crimes.csv")
-
-zip_codes$ZCTA5CE20 <- stringr::str_extract(zip_codes$address, "\\d{5}")
-
-crime_income_sample <- zip_codes %>% inner_join(unique(subset(zcta_no_NAs, select = c("ZCTA5CE20", "S1903_C03_001E", "year"))), by = c("ZCTA5CE20", "year"))
-
-# this takes a while, so loading in this csv instead
-
-crime_income_sample <- zip_codes %>% inner_join(unique(subset(zcta_no_NAs, select = c("ZCTA5CE20", "S1903_C03_001E", "year"))), by = c("ZCTA5CE20", "year"))
-
-crime_income_per_zip <- crime_income_sample %>% 
-  group_by(ZCTA5CE20) %>% 
-  summarize(number_of_crimes = n(),
-            cali_median_income = mean(S1903_C03_001E)) %>%
+```
+
+The maps in Figure 5 demonstrate that median income in certain LA zip codes (particularly those to the west) have a much higher median household income than the rest of the city, but over the last six years, the median income in most other LA zip codes has increased.
+However, there are some that a typical family is at or below the poverty line.
+We intend to investigate if there is a spatial correlation between these poorer zip codes and the incidence of crimes, particularly homicide, and whether this is affected by pandemic conditions.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Annual Median Income and Homicide Count by ZCTA in Los Angeles with top 1% income and top 1% homicide rates labelled", fig.align = "center"}
+cali_median_income$ZCTA5CE20 <- as.integer(cali_median_income$ZCTA)
+crime_income <- crimes_with_zips %>% inner_join(cali_median_income, by = c("year", "ZCTA5CE20"))
+
+crime_type <- crime_la %>% dplyr::select(year, Crm.Cd.Desc, DR_NO)
+
+crime_income_type <- crime_income %>% inner_join(crime_type, by = c("year", "DR_NO"))
+
+homicide_income_by_year <- crime_income_type %>% 
+  filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>%
+  group_by(year, ZCTA5CE20) %>%
+  summarize(HomicideCount = n(),
+            cali_median_income = median(as.numeric(S1903_C03_001E))) %>%
   ungroup()
 
-plot(x = crime_income_per_zip$cali_median_income, 
-     y = crime_income_per_zip$number_of_crimes, 
-     type = "p",
-     ylab = "Number of Crimes",  
-     xlab = "Median Income ($)",  
-     main = "Crime vs. Median Income by ZIP Code",
-     bty = "l")
-```
-
-####Figure 7 - Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025 {.unnumbered}
-
-```{r, eval=FALSE}
+homicide_income_by_year <- homicide_income_by_year %>% filter(!is.na(HomicideCount) & !is.na(cali_median_income))
+
+homicide_income_by_year$label <- ifelse(
+  homicide_income_by_year$cali_median_income > quantile(homicide_income_by_year$cali_median_income, 0.99) | 
+  homicide_income_by_year$HomicideCount > quantile(homicide_income_by_year$HomicideCount, 0.99), 
+  homicide_income_by_year$ZCTA5CE20, 
+  "")
+
+ggplot(data = homicide_income_by_year, aes(x = cali_median_income, 
+                                        y = HomicideCount)) +
+  geom_point(aes(color = as.factor(year))) +
+  ggthemes::scale_color_stata() +
+  ggrepel::geom_text_repel(aes(label = label), size = 2.5, nudge_x = 2, nudge_y = 1) + 
+  labs(x = "Median Annual Income ($)",
+       y = "Annual Homicide Count",
+       color = "Year")
+```
+
+Based on Figure 6, we observe a weak inverse relationship between median annual income and homicide count.
+Specifically, we see that higher homicide rates occur in ZCTA areas with lower median annual income, though most observations are still near or at zero, meaning few homicides typically occur.
+However, there are no observations with a median annual income of over \$100,000 where the homicide rate is greater than 10, whereas there are several of those observations for ZCTA/year combinations under \$100,000.
+The zip code which stands out most on this figure is 90003, which has 3 of the 4 highest homicide counts in the data.
+This suggests that there might be a relationship between median income and homicide count, though the relationship might be non-linear or weak, suggesting there are other factors at play.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025"}
 # fix mistake
@@
 
-#### Figure 8 - Annual Number of Crimes by Los Angeles Police Department Bureau {.unnumbered}
-
-```{r, eval=FALSE}
+Figure 7 plots the frequency of evictions in Los Angeles from February 2023 to February 2025, where the City of Los Angeles Housing Department has published case data.
+By summarizing both the crime and the evictions datasets by month, we see that the number of evictions was highest in early-mid 2023, and since, there have typically been slightly lower frequencies of evictions.
+Additionally, we split the number of crimes per month based on the average number of crimes of 14,439 to denote above and below average number of crimes per month.
+There is a clear split between March and April of 2024, as all months before that had above average crime, and all after had below average.
+Interestingly, the months where there is below average crime, seem to be the ones with slightly lower numbers of evictions.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Annual Number of Crimes by Los Angeles Police Department Bureau"}
 crime_la$year <- lubridate::year(crime_la$DateParsed)
@@
 
-#### Figure 9 - Correlations Between the Number of Crimes and Different Law Enforcement and Socioeconomic Factors {.unnumbered}
-
-```{r, eval=FALSE}
+Figure 8 shows the number of crimes in the four different bureaus of Los Angeles annually.
+We categorized the areas into their respective Central, South, Valley, and West Bureaus.
+The South Bureau has the highest median number of crimes, as well as the largest variability, while the Central, Valley, and West bureaus have relatively similar distributions with lower crime counts.The outliers suggests some years had significantly lower crime counts than others.
+This visualization helps us understand how crime is distributed across different regions, with the South experiencing the highest fluctuation and overall crime levels.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Strength of Correlations Between the Number of Crimes and Different Law Enforcement and Socioeconomic Factors"}
 library(lubridate)
@@
 
-#### Figure 10 - Homicide Victims by Age and Gender in Los Angeles 2010-Present {.unnumbered}
-
-```{r, eval=FALSE}
-homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex))
+Figure 9 shows the correlations between different variables that could affect the number of crimes committed.
+We see that the most significant correlations for the number of crimes are the unemployment rate, year, and number of evictions.
+Years is positively correlated with number of crimes, meaning that as time passed, the number of crimes increased.
+Similarly, as the number of evictions increases, so does the number of crimes.
+Interestingly, as the unemployment rate increases, the number of crimes decreases, which is the opposite effect of what was expected (one would anticipate that higher unemployment would result in more crime).
+The LAPD budget interestingly had a medium-strength positive correlation with number of crimes, suggesting we should further examine the increase in the LAPD budget compared to the increase in crimes.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Homicide Victims by Age and Gender in Los Angeles 2010-Present"}
+homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% dplyr::select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex))
 
@@
 
-#### Figure 11 - Frequency of Crimes in Los Angeles Throughout the Day {.unnumbered}
-
-```{r, eval=FALSE}
+The jitter plot in Figure 10 shows us both the frequencies and the demographics of homicide victims in Los Angeles from 2010 to present.
+Based on the density of the points, we see that most homicide victims in the last 15 years have been men between the ages of around 15 to 65.
+Additionally, we see that there are not many victims who are toddlers and younger children, but there is a good number of babies who have been killed.
+This is an intricacy in our data that might be interesting to look into further, particularly into the circumstances of those deaths and potential correlations between reports of child abuse and/or domestic violence and infanticide.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Frequency of Crimes in Los Angeles Throughout the Day"}
 crime_la$time <- sprintf("%04d",crime_la$TIME.OCC)
@@
 
-#### Figure 12 - Time series analyzing homicide rate trends over time by socioeconomic quartiles {.unnumbered}
-
-```{r, eval=FALSE}
+The density plot in Figure 11 shows the frequency of crimes depending on the hour (in military time) that they occurred.
+Based on the plot, we see that fewer crimes occur early in the morning (with the lowest at 5 AM).
+The data peaks at noon, meaning the most crimes occur in that hour of the day.
+Crimes are more frequent after that, until the early morning.
+The second most common crime occurrence is 6 PM, or 18 hours.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Time series analyzing homicide rate trends over time by median income quartiles"}
 # Data Preparation
@@
     y = "Number of Homicides",
-    fill = "Income Quartile",
-    title = "Homicide Rates in Los Angeles By Median Income Level"
+    fill = "Income Quartile"
   ) +
@@
 
-#### Crime \~ Week Regression Code {.unnumbered}
-
-```{r Regression 1, eval=FALSE}
+Figure 12 shows that homicide rates were low pre-pandemic and had a sharp increase since the pandemic, except in areas of low income, where homicide rates were high before the pandemic.
+In lower income areas, the rates of homicide have decreased significantly since the end of the pandemic.
+On the other hand, areas of higher-middle and high income have seen large increases in homicide in the six years shown.
+
+# Statistical Analyses
+
+To examine the relationship between the number of homicides and various socioeconomic and enforcement factors, we will use a variety of statistical analyses to see if certain factors are more influential in causing increases or decreases in homicide rates.
+
+For the purposes of this analysis, homicide rate is quantified as the annual number of homicides in a particular entity, whether that is ZIP Code Tabulation Area (ZCTA) or Census Tract.
+Additionally, we will examine the influence of general geographic location through the latitude and longitude where the homicide was committed.
+
+## Regression to Examine Weekly Crime Rates
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
 library(broom)
@@
 
-ar1_trend_regression %>%
-tidy() %>%
-  mutate(
-    p.value = scales::pvalue(p.value)
-  ) %>%
-  knitr::kable(
-    caption = "Linear Regression Model Results Estimating Crimes Per Week",
-    col.names = c("Predictor", "B", "SE", "t", "p"),
-    digits = c(0, 2, 3, 2, 3),
-    align = c("l", "r", "r", "r", "r")
+sjPlot::tab_model(ar1_trend_regression, dv.labels = "Table 1: Linear Regression Model Results Estimating Crimes Per Week", pred.labels = c("Intercept", 'Median Income'))
+```
+
+The lagged crime coefficient (0.81949) suggests that a high crime week is often followed by another high crime week.
+Since the trend term is not significant, there is no clear drift in weekly crime over time.
+
+## Regression Between Median Income and Crimes per Zip Code
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+lm_income_crimes <- lm(HomicideCount ~ cali_median_income, data = homicide_income_by_year)
+
+# summary(lm_income_crimes)
+
+sjPlot::tab_model(lm_income_crimes, dv.labels = "Table 2: Linear Regression Model Results Estimating Crimes Based on Median Income", pred.labels = c("Intercept", 'Median Income'))
+```
+
+Based on the p-value, we observe that median income is not significant at a 0.05 level, meaning it does not significantly decrease the amount of crimes committed.
+The r\^2 value additionally tells us that the income only explains a small portion of the variation in the number of crimes.
+As stated with the scatterplot, a larger sample and considering other factors might yield different results.
+
+## Negative Binomial for Census Block
+
+```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.align = "center"}
+library(kableExtra)
+crimes_by_cb <- read.csv("0_Data/Cleaned_Data/crimes_by_census_block.csv")
+colnames(crimes_by_cb) <- c(colnames(crimes_by_cb)[-c(1, length(colnames(crimes_by_cb)))],"LON", "LAT")
+
+crimes_by_cb$LON <- as.numeric(gsub('c(', "", crimes_by_cb$LON, fixed = TRUE))
+crimes_by_cb$LAT <- as.numeric(gsub(')', "", crimes_by_cb$LAT, fixed = TRUE))
+
+crimes_by_cb$DateParsed <- as.POSIXct(crimes_by_cb$Date.Rptd, 
+                                  format = "%m/%d/%Y %I:%M:%S %p", 
+                                  tz = "America/Los_Angeles")
+
+crimes_by_cb$YearWeek <- format(crimes_by_cb$DateParsed, "%Y-%U")
+
+crimes_by_cb$year <- as.integer(stringr::str_sub(crimes_by_cb$YearWeek, 1, 4))
+
+homicides_by_cb <- crimes_by_cb %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% group_by(BLKGRPCE, year) %>% summarize(HomicideCount = n()) %>% ungroup()
+
+# shapiro.test(homicides_by_cb$HomicideCount) # cannot use parametric tests. Strong evidence against normality.
+
+mean_homicides <- mean(homicides_by_cb$HomicideCount)
+var_homicides <- var(homicides_by_cb$HomicideCount)
+# since variance > mean, use negative binomial 
+library(MASS)
+model <- glm.nb(HomicideCount ~ as.factor(BLKGRPCE), data = homicides_by_cb)
+
+pretty_labels <- c("(Intercept)", paste("Census Block", 2:6))
+sjPlot::tab_model(model, pred.labels = pretty_labels, dv.labels = "Table 3: Annual Homicide Count by Census Block", show.est = TRUE)
+```
+
+### Discussion
+
+The results of this model show a significant relationship between Census Block and the homicide rate, except for Census Block 2.
+Census Block 2 experiences a higher incidence of homicides, but it is not significantly different.
+The other Census Blocks show a statistically significant decrease in homicide rates.
+However, the R2 might suggest overfitting.
+
+## Linear Regression Models
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+homicides_latlon_rounded <- homicides %>%
+  mutate(LAT_rounded = round(LAT, 2),
+         LON_rounded = round(LON, 2))
+
+homicides_lat_long_yr <- homicides_latlon_rounded %>% 
+  group_by(year, LAT_rounded, LON_rounded) %>%
+  summarize(HomicideCount = n()) %>% ungroup()
+
+linear_model1 <- lm(HomicideCount ~ LAT_rounded + LON_rounded, data = homicides_lat_long_yr)
+```
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+
+linear_model2 <- lm(HomicideCount ~ year, data = homicides_lat_long_yr)
+
+linear_model3 <- lm(HomicideCount ~ LAT_rounded + LON_rounded + year, data = homicides_lat_long_yr)
+
+linear_model4 <- lm(HomicideCount ~ LAT_rounded * LON_rounded * year, data = homicides_lat_long_yr)
+
+sjPlot::tab_model(linear_model1, linear_model2, linear_model3, linear_model4, pred.labels = c("(Intercept)", "Latitude", "Longitude", "Year", "Latitude:Longitude", "Latitude:Year", "Longitude:Year"), dv.labels = c("Model 1", "Model 2", "Model 3", "Model 4"), show.ci = FALSE, title =
+                    "Table 4: Geospatial Linear Regression Models Predicting Homicide Count")
+
+```
+
+Table 4 shows the results of four different linear regression models that were run using homicide counts by year and by latitude and longitude.
+Longitude and latitude were rounded to two decimal places to summarize the counts but maintain a good level of granularity in the coordinates data.
+Model 1 looked only at the geospatial aspect of the data and found that both latitude and longitude were significant predictors of homicide rate.
+As latitude increases, homicide rate decreases, meaning that homicides are more common towards the south and as you move north, homicide counts drop.
+As longitude increases, the number of homicides increases as well, meaning that homicides are more common to the East.
+
+The second model finds that year is significant as a predictor of homicide rate, and a model looking at both spatial and temporal characteristics also finds year to be significant.
+
+The fourth model, which also considered the interactions between these three features, did not find any significance.
+
+# References {.unnumbered}
+
+::: {#refs}
+:::
+
+# Data Sources {.unnumbered}
+
+## Primary Datasets {.unnumbered}
+
+Crime Data From 2010-2019: <https://data.lacity.org/Public-Safety/Crime-Data-from-2010-to-2019/63jg-8b9z/about_data>
+
+Crime Data From 2020-Present: <https://catalog.data.gov/dataset/crime-data-from-2020-to-present>
+
+## Secondary Datasets {.unnumbered}
+
+Los Angeles County Eviction Data: <https://calmatters.org/housing/homelessness/2023/11/california-evictions-post-pandemic/>
+
+City of Los Angeles Eviction Data: <https://housing.lacity.gov/residents/renters/eviction-notices-filed>
+
+Median Income Data: <https://data.census.gov/table/ACSST1Y2023.S1903?q=income>
+
+# Appendices {.appendix}
+
+#### Data Wrangling {.unnumbered}
+
+##### Reading in Raw Datasets
+
+```{r DATA_Wrangling, output=FALSE, eval=FALSE}
+library(sf) # for map
+library(tidyverse) # using only for joining datasets to build map visual
+library(RColorBrewer) # for map color scheme
+library(tidygeocoder) # to get zip codes of addresses
+library(dplyr)
+
+# primary datasets
+# Primary datasets: Crime Data 2020-Present and 2010-2019
+crime_la_2010_2019 <- read.csv("0_Data/Raw_Data/Crime_Data_from_2010_to_2019.csv")
+
+crime_la_2020_present <- read.csv("0_Data/Raw_Data/Crime_Data_from_2020_to_Present.csv")
+
+crime_la <- read.csv("0_Data/Cleaned_Data/all_crimes_2010_present.csv")
+
+# primary dataset zip code and geometry data
+crimes_with_zips <- read.csv("0_Data/Cleaned_Data/crimes_with_zips.csv") # just inner_join with the crime_la dataset
+
+# Secondary dataset - Median Income by ZCTA
+cali_median_income <- read.csv("0_Data/Cleaned_Data/cali_median_income.csv")
+
+# adding a variable that's just the ZCTA zip code
+cali_median_income$ZCTA <- stringr::str_sub(cali_median_income$NAME, -5)
+
+# Secondary dataset - COVID cases in Los Angeles County
+covid_14day <- read.csv("0_Data/Raw_Data/LA_County_Covid19_CSA_14day_case_death_table.csv")
+
+evictions_la <- read.csv("0_Data/Raw_Data/LA_evictions.csv")
+
+crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
+                                  format = "%m/%d/%Y %I:%M:%S %p", 
+                                  tz = "America/Los_Angeles")
+
+crime_la$month <- lubridate::month(crime_la$DateParsed)
+
+crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
+
+crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
+```
+
+```{r Combining Median Income, eval=FALSE}
+# Secondary datasets
+california_median_income_2018_2023 <- list()
+yrs <- 2018:2023
+for (i in 1:length(yrs)) {
+  obj <- read.csv(paste0("0_Data/Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
+  obj <- obj[-1,]
+  obj$year <- yrs[[i]]
+  california_median_income_2018_2023[[i]] <- obj
+}
+
+cali_median_income <- do.call(rbind, california_median_income_2018_2023)
+
+# adding a variable that's just the ZCTA zip code
+cali_median_income$ZCTA <- stringr::str_sub(cali_median_income$NAME, -5)
+
+write.csv(cali_median_income, "cali_median_income.csv")
+```
+
+##### Zip Codes for Crimes
+
+```{r Get Zip Codes, eval=FALSE}
+library(sf)
+library(dplyr)
+
+crime_la_2010_2019 <- read.csv("0_Data/Raw_Data/Crime_Data_from_2010_to_2019.csv")
+
+crime_la_2020_present <- read.csv("0_Data/Raw_Data/Crime_Data_from_2020_to_Present.csv")
+
+# Combine the two datasets
+crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
+
+crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
+                                  format = "%m/%d/%Y %I:%M:%S %p", 
+                                  tz = "America/Los_Angeles")
+
+crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
+
+crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
+
+coords <- crime_la %>% dplyr::select(LAT, LON, DR_NO, year)
+
+zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
+
+  points <- st_as_sf(coords, coords = c("LON", "LAT"), crs = 4326)  # WGS84
+
+  zcta <- st_transform(zcta, 4326)  # Convert ZCTA to WGS84 if needed
+  cat("Starting Zip Code Retrieval \n")
+
+  result <- st_join(points, zcta)
+
+  cat("Finished Zip Code Retrieval \n")
+
+  write.csv(result, "0_Data/Raw_Data/all_zips.csv")
+
+crime_zctas <- read.csv("0_Data/Raw_Data/all_zips.csv")
+colnames(crime_zctas) <- c(colnames(crime_zctas)[-c(1, length(colnames(crime_zctas)))],"LON", "LAT")
+
+crime_zctas$LON <- as.numeric(gsub('c(', "", crime_zctas$LON, fixed = TRUE))
+crime_zctas$LAT <- as.numeric(gsub(')', "", crime_zctas$LAT, fixed = TRUE))
+
+# Repeat same process for Census Blocks
+```
+
+```{r Get Zip Codes 2, eval=FALSE}
+
+duplicates <- as.integer(as.character(data.frame(table(crime_zctas$universal_idx)[table(crime_zctas$universal_idx) > 1]) %>% pull(Var1)))
+
+# there are approximately 200 rows that are duplicated (100 incidents) as the area lies on the intersection between two zctas. 
+duplicate_data <- crime_zctas %>% filter(universal_idx %in% duplicates)
+
+rev_geo_duplicates <- reverse_geocode(duplicates2, lat = LAT, long = LON)
+# the above query identifies the address of these coordinates as KFC, 8644, Balboa Boulevard, Northridge South Neighborhood Council District, Los Angeles, Los Angeles County, California, 91325, United States.
+# therefore, we will only keep the 91325 zip code rows.
+
+crime_zctas_no_duplicates <- crime_zctas %>% mutate(duplicate = ifelse(universal_idx %in% duplicate_data$universal_idx & ZCTA5CE20 != 91325, "Drop", "Keep")) %>% filter(duplicate != "Drop") %>% dplyr::select(-duplicate)
+
+# it's empty now 
+# duplicates <- table(crime_zctas_no_duplicates$universal_idx)[table(crime_zctas_no_duplicates$universal_idx) > 1]
+
+# next, check for NAs
+sum(is.na(crime_zctas_no_duplicates$ZCTA5CE20)) # there are 3724 NAs. Let's find them a zip code
+
+# some also have lat long = 0. We'll find those separately
+no_zipcode_not0 <- crime_zctas_no_duplicates %>% filter(is.na(ZCTA5CE20) & LON != 0)
+
+no_zipcode_not0_zips <- reverse_geocode(no_zipcode_not0, lat = LAT, long = LON)
+
+no_zipcode_not0_zips$ZCTA5CE20 <- stringr::str_extract(no_zipcode_not0_zips$address, "\\d{5}")
+
+no_zipcode_not0_zips <- no_zipcode_not0_zips %>% dplyr::select(-address)
+
+crime_zctas2 <- crime_zctas_no_duplicates %>% filter(!is.na(ZCTA5CE20))
+
+crime_zctas3 <- rbind(crime_zctas2, no_zipcode_not0_zips)
+
+crime_zctas4 <- crime_zctas3 %>% filter(!is.na(ZCTA5CE20))
+
+write.csv(crime_zctas4, "crimes_with_zips.csv")
+
+no_zipcode_0s <- crime_zctas_no_duplicates %>% filter(is.na(ZCTA5CE20) & LON == 0) %>% pull(DR_NO)
+
+main_dataset_0s <- crime_la %>% filter(DR_NO %in% no_zipcode_0s)
+
+summary(main_dataset_0s)
+
+summary(crime_la)
+```
+
+```{r eval=FALSE, fig.cap="Crime densities in Los Angeles by LAPD Division, message=FALSE, warning=FALSE, 2010.2019 and 2020.Present", fig.align="center"}
+lapd_shape <- read_sf(dsn = "0_Data/Raw_Data/LAPD_Divisions", layer = "LAPD_Divisions")
+
+lapd_shape <- st_transform(lapd_shape, crs = 4326)
+
+cali <- read_sf(dsn = "0_Data/Raw_Data/ca_counties", layer = "ca_counties")
+
+cali <- st_transform(cali, crs = 4326)
+
+par(mfrow = c(1, 2), mar = c(4, 4, 2, 1))
+
+smoothScatter(
+  x    = crime_la_2010_2019$LON,
+  y    = crime_la_2010_2019$LAT,
+  xlim = c(-119, -118),
+  ylim = c(33.5, 34.5),
+  nbin = 300,
+  xlab = "Longitude",
+  ylab = "Latitude",
+  main = "Crime Density (2010–2019)",
+  colramp = colorRampPalette(c("white", "skyblue", "blue", "darkblue")),
+  cex.main=0.8
 )
-```
-
-#### Crime \~ Income Regression Code {.unnumbered}
-
-```{r Regression 2, eval=FALSE}
-lm_income_crimes <- lm(number_of_crimes ~ cali_median_income, data = crime_income_per_zip)
-
-summary(lm_income_crimes)
-
-lm_income_crimes %>%
-tidy() %>%
-  mutate(
-    p.value = scales::pvalue(p.value)
-  ) %>%
-  knitr::kable(
-    caption = "Linear Regression Model Results Estimating Crimes Based on Median Income",
-    col.names = c("Predictor", "B", "SE", "t", "p"),
-    digits = c(0, 2, 3, 2, 3),
-    align = c("l", "r", "r", "r", "r")
+
+plot(lapd_shape$geometry, lwd=.5, add=TRUE)
+
+plot(cali$geometry, add = TRUE, lwd=.5)
+
+smoothScatter(
+  x    = crime_la_2020_present$LON,
+  y    = crime_la_2020_present$LAT,
+  xlim = c(-119, -118),
+  ylim = c(33.5, 34.5),
+  nbin = 300,
+  xlab = "Longitude",
+  ylab = "Latitude",
+  main = "Crime Density (2020–Present)",
+  colramp = colorRampPalette(c("white", "pink", "red", "darkred")),
+  cex.main=0.8
 )
-```
+
+plot(lapd_shape$geometry, lwd=.5, add=TRUE)
+
+plot(cali$geometry, add = TRUE, lwd=.5)
+
+par(mfrow = c(1, 1))
+
+```
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.cap = "Weekly Crime Counts in Los Angeles from 2010 to present", fig.align = "center"}
+crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
+                                  format = "%m/%d/%Y %I:%M:%S %p", 
+                                  tz = "America/Los_Angeles")
+
+crime_la$month <- lubridate::month(crime_la$DateParsed)
+
+crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
+
+weekly_counts <- aggregate(DR_NO ~ YearWeek, data = crime_la, FUN = length)
+names(weekly_counts)[2] <- "Crimes"
+
+weekly_counts <- weekly_counts[order(weekly_counts$YearWeek), ]
+
+weekly_counts$Index <- seq_len(nrow(weekly_counts))
+
+plot(
+  weekly_counts$Index,
+  weekly_counts$Crimes,
+  type = "l",
+  col  = "blue",
+  lwd  = 2,
+  xlab = "Weekly Index",
+  ylab = "Number of Crimes")
+
+abline(h = pretty(weekly_counts$Crimes), v = pretty(weekly_counts$Index), 
+       col = "gray", lty = "dotted")
+```
+
+```{r, eval=FALSE, fig.cap = "Los Angeles Crimes by Day of the Week", warning=FALSE, message=FALSE, fig.align = "center"}
+# Convert Date.Rptd to Date format (ignore time part)
+#crime_la$Date.Rptd
+crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")
+
+# Extract the day of the week
+crime_la$DayOfWeek <- weekdays(crime_la$Date)
+# crime_la$DayOfWeek
+
+# Count occurrences of crimes per day
+crime_counts <- table(crime_la$DayOfWeek)
+# crime_counts
+
+# Order the days correctly
+day_order <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
+crime_counts <- data.frame(crime_counts[day_order])
+colnames(crime_counts) <- c("Day", "CrimeCount")
+
+crime_counts$thousands <- crime_counts$CrimeCount/1000
+
+# Plot bar chart with horizontal labels
+ggplot(crime_counts, aes(x = Day, y = thousands)) +
+  geom_col(fill = "blue4") +
+  labs(x = "Day of the Week",
+       y = "Number of Crimes (thousands)")
+```
+```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.cap="Percent of crime victims by gender, 2010 to present", fig.align = "center"}
+# Remove missing values and empty strings in Vict.Sex column
+crime_la <- subset(crime_la, Vict.Sex != "" & !is.na(Vict.Sex))
+
+# Replace all non-"F" and non-"M" values with "N"
+crime_la$Vict.Sex <- ifelse(crime_la$Vict.Sex %in% c("F", "M"), 
+                                  crime_la$Vict.Sex, "N")
+
+# Count occurrences of each gender
+gender_counts <- table(crime_la$Vict.Sex)
+
+# Create a pie chart using base R
+color <- c("F" = "pink", "M" = "lightblue", "N" = "purple")
+pie(
+  gender_counts, 
+  labels = paste(names(gender_counts), 
+                 round(gender_counts / sum(gender_counts) * 100, 1), "%"), 
+  col = color[names(gender_counts)]
+)
+```
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
+LA_zctas <- read.csv("0_Data/Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs
+
+colnames(LA_zctas) <- c("X", "ZCTA5CE20")
+LA_zctas$ZCTA5CE20 <- as.character(LA_zctas$ZCTA5CE20)
+
+# joining the zcta shape file with LA zcta reference file
+zcta_filtered <- zcta %>% inner_join(LA_zctas, by = "ZCTA5CE20")
+
+# getting LA median income using joined file
+median_household_income <- cali_median_income %>% dplyr::select(ZCTA, S1903_C03_001E, year)
+
+zcta_with_data <- zcta_filtered %>% inner_join(median_household_income, join_by("ZCTA5CE20" == "ZCTA"))
+
+# remove NAs
+zcta_no_NAs <- zcta_with_data %>% filter(!is.na(S1903_C03_001E), S1903_C03_001E != "-")
+
+zcta_no_NAs$S1903_C03_001E <- as.integer(zcta_no_NAs$S1903_C03_001E)
+```
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.cap="Median household income by year and by ZIP Code Tabulation Area (ZCTA) in Los Angeles", fig.align = "center"}
+# quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
+qntls <- round(quantile(zcta_no_NAs$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0)
+
+# 2x3 plot layout for years 2018-2023, setting margins and outer margins, main title sizes
+par(mfrow = c(2, 3), mar = c(1, 1, 1, 5), oma = c(0, 0, 3, 0), cex.main = 1.5)
+
+# color palette for sequential data
+palette <- brewer.pal(7, "Reds")
+
+yrs <- 2018:2023
+
+for (yr in yrs) {
+  # get just that year's data
+  subset_data <- zcta_no_NAs[zcta_no_NAs$year == yr, ]
+
+  # creating bins based on the qtls
+  bins <- cut(subset_data$S1903_C03_001E, breaks = qntls, labels = FALSE, include.lowest = TRUE)
+
+  # Plot
+  plot(st_geometry(subset_data),
+       col = palette[bins],
+       border = 'grey',
+       axes = FALSE,
+       lwd = 0.5)
+  title(yr, line = -1)
+
+  # adding legend
+  legend("right", inset = c(-.75, 0), legend = qntls, fill = palette, title = "Median Income ($)", bty = "n", cex = 0.8, xpd = TRUE)
+}
+
+```
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.cap="Annual Median Income and Homicide Count by ZCTA in Los Angeles with top 1% income and top 1% homicide rates labelled", fig.align = "center"}
+cali_median_income$ZCTA5CE20 <- as.integer(cali_median_income$ZCTA)
+crime_income <- crimes_with_zips %>% inner_join(cali_median_income, by = c("year", "ZCTA5CE20"))
+
+crime_type <- crime_la %>% dplyr::select(year, Crm.Cd.Desc, DR_NO)
+
+crime_income_type <- crime_income %>% inner_join(crime_type, by = c("year", "DR_NO"))
+
+homicide_income_by_year <- crime_income_type %>% 
+  filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>%
+  group_by(year, ZCTA5CE20) %>%
+  summarize(HomicideCount = n(),
+            cali_median_income = median(as.numeric(S1903_C03_001E))) %>%
+  ungroup()
+
+homicide_income_by_year <- homicide_income_by_year %>% filter(!is.na(HomicideCount) & !is.na(cali_median_income))
+
+homicide_income_by_year$label <- ifelse(
+  homicide_income_by_year$cali_median_income > quantile(homicide_income_by_year$cali_median_income, 0.99) | 
+  homicide_income_by_year$HomicideCount > quantile(homicide_income_by_year$HomicideCount, 0.99), 
+  homicide_income_by_year$ZCTA5CE20, 
+  "")
+
+ggplot(data = homicide_income_by_year, aes(x = cali_median_income, 
+                                        y = HomicideCount)) +
+  geom_point(aes(color = as.factor(year))) +
+  ggthemes::scale_color_stata() +
+  ggrepel::geom_text_repel(aes(label = label), size = 2.5, nudge_x = 2, nudge_y = 1) + 
+  labs(x = "Median Annual Income ($)",
+       y = "Annual Homicide Count",
+       color = "Year")
+```
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025"}
+# fix mistake
+evictions_la$Date.Received <- gsub("2205", "2025", evictions_la$Date.Received)
+
+# convert to date
+evictions_la$month_year <- format(as.Date(evictions_la$Date.Received, format="%m-%d-%Y"), "%Y-%m")
+
+# get unique months
+unique_months <- data.frame(month_year = sort(unique(evictions_la$month_year)), index = 1:length(unique(evictions_la$month_year)))
+
+evictions_la_indexed <- evictions_la %>% left_join(unique_months, by = "month_year")
+
+# create color scheme
+crime_la$month_year <- format(crime_la$DateParsed, "%Y-%m")
+crime_la_monthly <- crime_la %>% 
+  group_by(month_year) %>% 
+  summarize(num_crimes = n())
+
+evictions_la_indexed_crimes <- evictions_la_indexed %>% left_join(crime_la_monthly, by = "month_year")
+
+avg_crimes <- mean(evictions_la_indexed_crimes$num_crimes)
+
+evictions_la_indexed_crimes$high_low_crime <- as.factor(ifelse(evictions_la_indexed_crimes$num_crimes > avg_crimes, "Above Average", "Below Average"))
+
+ggplot(data = evictions_la_indexed_crimes, aes(x = index)) + 
+  geom_histogram(aes(fill = high_low_crime), bins = nrow(unique_months)) +
+  scale_x_continuous(breaks=unique_months$index,
+        labels=unique_months$month_year) + 
+  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", fill = "Crime Level") +
+  ggthemes::scale_fill_fivethirtyeight() + 
+  ggthemes::theme_stata(scheme = "s1mono") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
+        axis.title.x = element_text(vjust = -.75),
+        plot.title = element_text(size=12))
+```
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Annual Number of Crimes by Los Angeles Police Department Bureau"}
+crime_la$year <- lubridate::year(crime_la$DateParsed)
+crime_by_area <- crime_la %>% group_by(AREA.NAME, year) %>% summarize(numberOfCrimes = n()) %>% ungroup()
+
+bureaus <- read.csv("0_Data/Raw_Data/Division_Bureau_Ref.csv")
+
+bureaus$Division <- gsub("77th", "77th Street", bureaus$Division)
+bureaus$Division <- gsub("North", "N", bureaus$Division)
+bureaus$Division <- gsub("Neast", "Northeast", bureaus$Division)
+
+crime_by_area <- crime_by_area %>% left_join(bureaus, join_by("AREA.NAME" == "Division"))
+
+ggplot(crime_by_area, aes(y = numberOfCrimes, group = Bureau, x = Bureau)) + 
+        geom_boxplot(aes(fill = Bureau)) +
+  labs(y = "Annual Number of Crimes")
+```
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Strength of Correlations Between the Number of Crimes and Different Law Enforcement and Socioeconomic Factors"}
+library(lubridate)
+# creating dataset for correlations
+
+# since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions. 
+LA_county_evictions <- read.csv("0_Data/Raw_Data/LA_county_evictions.csv")
+
+# covid data
+covid_14day$date <- ymd(covid_14day$ep_date)
+covid_14day$month <- month(covid_14day$date)
+covid_14day$year <- year(covid_14day$date)
+covid_14day$date <- day(covid_14day$date)
+# since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
+covid_14day$lagged <- covid_14day$date + 7
+
+covid_la_lagged <- covid_14day %>% filter(lagged %in% c(21, 35)) %>% arrange(month, year)
+
+covid_la_monthly <- covid_la_lagged %>% group_by(month, year) %>% summarize(cases = sum(cases_14day)) %>% ungroup() %>% arrange(year, month)
+
+covid_la_monthly$index <- 1:nrow(covid_la_monthly)
+
+# next, unemployment rate
+unemployment <- read.csv("0_Data/Raw_Data/la_county_unemployment_rate.csv")
+unemployment$month <- as.numeric(gsub("M", "", unemployment$Period))
+
+# lapd budget
+lapd_budget <- read.csv("0_Data/Raw_Data/lapd_budget.csv")
+
+crime_la_monthly$date <- ymd(paste0(crime_la_monthly$month_year, "-01"))
+crime_la_monthly$month <- month(crime_la_monthly$date)
+crime_la_monthly$year <- year(crime_la_monthly$date)
+
+correlations_df <- crime_la_monthly %>% inner_join(unemployment, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, Value, num_crimes) %>% rename("unemployment_rate" = "Value")
+
+correlations_df2 <- correlations_df %>% inner_join(covid_la_monthly, by = c("month", "year"))
+
+correlations_df3 <- correlations_df2 %>% inner_join(LA_county_evictions, join_by("year" == "Year", "month" == "Month")) %>% rename("num_evictions" = "Count")
+
+correlations_df4 <- correlations_df3 %>% inner_join(lapd_budget, by = "year")
+
+correlations <- correlations_df4 %>% dplyr::select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)
+
+Cor_mat <- round(cor(correlations), 2)
+
+long_Cor <- reshape2::melt(Cor_mat)
+
+long_Cor <- long_Cor %>%
+    mutate(Var1 = recode(Var1, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget ($ billions)"),
+           Var2 = recode(Var2, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget \n ($ billions)"))
+
+ggplot(data = long_Cor, aes(x = Var1, y = Var2, fill = value)) +
+  geom_tile(color = "white") +
+  scale_fill_viridis_c(direction = -1,limit = c(-1, 1)) +
+  theme(axis.title.x =  element_blank(),
+        axis.title.y = element_blank()) +
+  geom_text(aes(label = value), color = "black", size = 4) + 
+   theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.title = element_text(size=12)) +
+  labs(fill = "Correlation")
+```
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Homicide Victims by Age and Gender in Los Angeles 2010-Present"}
+homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% dplyr::select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex))
+
+levels(homicide_by_gender$gender) <- c("Female", "Male", "Other")
+ggplot(homicide_by_gender, aes(x = gender, y = Vict.Age, color = gender)) +
+  geom_jitter(alpha = 0.4, width = 0.2) +
+  labs(x = "Gender",
+       y = "Age") +
+  theme_minimal() +
+  scale_color_manual(c("M", "F", "N"), values = c("pink2", "blue", "purple")) + 
+  theme(legend.position = 'none')
+```
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Frequency of Crimes in Los Angeles Throughout the Day"}
+crime_la$time <- sprintf("%04d",crime_la$TIME.OCC)
+crime_la$Hour <- as.numeric(substr(as.character(crime_la$time), start = 1, stop = 2))
+
+ggplot(crime_la, aes(x = Hour, y = 0, fill = after_stat(density))) +
+  ggridges::geom_density_ridges_gradient(scale = 1, alpha = 0.5) +
+  scale_fill_gradient(low = "lightblue", high = "darkblue") +
+  labs(x = "Hour (Military Time)",
+       y = "Density") +
+  theme_minimal() +
+  theme(legend.position = 'none')
+```
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Time series analyzing homicide rate trends over time by median income quartiles"}
+# Data Preparation
+homicides <- crime_la %>% inner_join(crimes_with_zips, by = c("DR_NO", "LAT", "LON","X", "year")) %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% 
+  mutate(date = as.Date(DateParsed),
+         zip = as.character(ZCTA5CE20)) %>% dplyr::select(-X)
+
+homicides <- homicides %>% inner_join(cali_median_income, join_by("zip" == "ZCTA", "year" == "year"))
+homicides$S1903_C03_001E <- as.numeric(homicides$S1903_C03_001E)
+
+# Form income quartiles
+homicides <- homicides %>% mutate(income_quartile = ntile(S1903_C03_001E, 4)) 
+homicide_trends <- homicides %>% group_by(year, income_quartile) %>% summarize(homicide_count = n(), 
+          ref = mean(S1903_C03_001E),
+          .groups = "drop")
+
+homicide_trends$income_quartile <- as.factor(homicide_trends$income_quartile)
+# removing NAs
+homicide_trends <- homicide_trends %>% filter(!is.na(homicide_count) & !is.na(income_quartile))
+
+# creating labels
+levels(homicide_trends$income_quartile) <- c("Low Income", "Lower-Middle Income", "Higher-Middle Income", "High Income")
+# Plot homicide trends over time by income quartile
+ggplot(homicide_trends, aes(x = year, y = homicide_count)) +
+  geom_area(aes(fill = as.factor(income_quartile)), alpha = 0.6, position = "identity") +
+  labs(
+    x = "Year",
+    y = "Number of Homicides",
+    fill = "Income Quartile"
+  ) +
+  theme_minimal() +
+  scale_fill_viridis_d(option = "plasma", direction = -1)
+```
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+library(broom)
+
+weekly_counts$CrimesLag <- c(NA, weekly_counts$Crimes[-nrow(weekly_counts)])
+
+ar1_trend_regression <- lm(Crimes ~ Index + CrimesLag, data = weekly_counts, na.action = na.omit)
+
+sjPlot::tab_model(ar1_trend_regression, dv.labels = "Table 1: Linear Regression Model Results Estimating Crimes Per Week", pred.labels = c("Intercept", 'Median Income'))
+```
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+lm_income_crimes <- lm(HomicideCount ~ cali_median_income, data = homicide_income_by_year)
+
+# summary(lm_income_crimes)
+
+sjPlot::tab_model(lm_income_crimes, dv.labels = "Table 2: Linear Regression Model Results Estimating Crimes Based on Median Income", pred.labels = c("Intercept", 'Median Income'))
+```
+
+```{r, warning=FALSE, message=FALSE, eval=FALSE, fig.align = "center"}
+library(kableExtra)
+crimes_by_cb <- read.csv("0_Data/Cleaned_Data/crimes_by_census_block.csv")
+colnames(crimes_by_cb) <- c(colnames(crimes_by_cb)[-c(1, length(colnames(crimes_by_cb)))],"LON", "LAT")
+
+crimes_by_cb$LON <- as.numeric(gsub('c(', "", crimes_by_cb$LON, fixed = TRUE))
+crimes_by_cb$LAT <- as.numeric(gsub(')', "", crimes_by_cb$LAT, fixed = TRUE))
+
+crimes_by_cb$DateParsed <- as.POSIXct(crimes_by_cb$Date.Rptd, 
+                                  format = "%m/%d/%Y %I:%M:%S %p", 
+                                  tz = "America/Los_Angeles")
+
+crimes_by_cb$YearWeek <- format(crimes_by_cb$DateParsed, "%Y-%U")
+
+crimes_by_cb$year <- as.integer(stringr::str_sub(crimes_by_cb$YearWeek, 1, 4))
+
+homicides_by_cb <- crimes_by_cb %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% group_by(BLKGRPCE, year) %>% summarize(HomicideCount = n()) %>% ungroup()
+
+# shapiro.test(homicides_by_cb$HomicideCount) # cannot use parametric tests. Strong evidence against normality.
+
+mean_homicides <- mean(homicides_by_cb$HomicideCount)
+var_homicides <- var(homicides_by_cb$HomicideCount)
+# since variance > mean, use negative binomial 
+library(MASS)
+model <- glm.nb(HomicideCount ~ as.factor(BLKGRPCE), data = homicides_by_cb)
+
+pretty_labels <- c("(Intercept)", paste("Census Block", 2:6))
+sjPlot::tab_model(model, pred.labels = pretty_labels, dv.labels = "Table 3: Annual Homicide Count by Census Block")
+```
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+homicides_latlon_rounded <- homicides %>%
+  mutate(LAT_rounded = round(LAT, 2),
+         LON_rounded = round(LON, 2))
+
+homicides_lat_long_yr <- homicides_latlon_rounded %>% 
+  group_by(year, LAT_rounded, LON_rounded) %>%
+  summarize(HomicideCount = n()) %>% ungroup()
+
+linear_model1 <- lm(HomicideCount ~ LAT_rounded + LON_rounded, data = homicides_lat_long_yr)
+```
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+
+linear_model2 <- lm(HomicideCount ~ year, data = homicides_lat_long_yr)
+
+linear_model3 <- lm(HomicideCount ~ LAT_rounded + LON_rounded + year, data = homicides_lat_long_yr)
+
+linear_model4 <- lm(HomicideCount ~ LAT_rounded * LON_rounded * year, data = homicides_lat_long_yr)
+
+sjPlot::tab_model(linear_model1, linear_model2, linear_model3, linear_model4, pred.labels = c("(Intercept)", "Latitude", "Longitude", "Year", "Latitude:Longitude", "Latitude:Year", "Longitude:Year"), dv.labels = c("Model 1", "Model 2", "Model 3", "Model 4"), show.ci = FALSE, title =
+                    "Table 4: Geospatial Linear Regression Models Predicting Homicide Count")
+
+```

### 2025-04-10 Draft4.qmd
Draft 4

--- 2025-03-08 Draft3.qmd
@@
 
-```{r DATA, output=FALSE, echo=FALSE}
+```{r DATA, output=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
 library(sf) # for map
@@
 library(dplyr)
-
-# primary datasets
-# Primary datasets: Crime Data 2020-Present and 2010-2019
-crime_la_2010_2019 <- read.csv("0_Data/Raw_Data/Crime_Data_from_2010_to_2019.csv")
-
-crime_la_2020_present <- read.csv("0_Data/Raw_Data/Crime_Data_from_2020_to_Present.csv")
-
-crime_la <- read.csv("0_Data/Cleaned_Data/all_crimes_2010_present.csv")
-
-# primary dataset zip code and geometry data
-crimes_with_zips <- read.csv("0_Data/Cleaned_Data/crimes_with_zips.csv") # just inner_join with the crime_la dataset
-
-# Secondary dataset - Median Income by ZCTA
-cali_median_income <- read.csv("0_Data/Cleaned_Data/cali_median_income.csv")
-
-# adding a variable that's just the ZCTA zip code
-cali_median_income$ZCTA <- stringr::str_sub(cali_median_income$NAME, -5)
-
-# Secondary dataset - COVID cases in Los Angeles County
-covid_14day <- read.csv("0_Data/Raw_Data/LA_County_Covid19_CSA_14day_case_death_table.csv")
-
-evictions_la <- read.csv("0_Data/Raw_Data/LA_evictions.csv")
-
-crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
-                                  format = "%m/%d/%Y %I:%M:%S %p", 
-                                  tz = "America/Los_Angeles")
-
-crime_la$month <- lubridate::month(crime_la$DateParsed)
-
-crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
-
-crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
+library(RSQLite)
+
+con <- dbConnect(SQLite(), "0_Data/crime_in_la.db")
 ```
@@
 
-```{r echo=FALSE, fig.cap="Crime densities in Los Angeles by LAPD Division, message=FALSE, warning=FALSE, 2010.2019 and 2020.Present", fig.align="center"}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Crime densities in Los Angeles by LAPD Division, 2010-2019 and 2020-Present"}
+query <- dbSendQuery(con, "Select * From crime_la Where year < 2020")
+crime_la_2010_2019 <- dbFetch(query,n=-1)
+dbClearResult(query)
+
+query <- dbSendQuery(con, "Select * From crime_la Where year > 2019")
+crime_la_2020_present <- dbFetch(query,n=-1)
+dbClearResult(query)
+
 lapd_shape <- read_sf(dsn = "0_Data/Raw_Data/LAPD_Divisions", layer = "LAPD_Divisions")
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap = "Weekly Crime Counts in Los Angeles from 2010 to present", fig.align = "center"}
-crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
-                                  format = "%m/%d/%Y %I:%M:%S %p", 
-                                  tz = "America/Los_Angeles")
-
-crime_la$month <- lubridate::month(crime_la$DateParsed)
-
-crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
-
-weekly_counts <- aggregate(DR_NO ~ YearWeek, data = crime_la, FUN = length)
-names(weekly_counts)[2] <- "Crimes"
-
-weekly_counts <- weekly_counts[order(weekly_counts$YearWeek), ]
+query <- dbSendQuery(con, "Select Count(*) as Crimes, YearWeek From crime_la Group By YearWeek Order By YearWeek")
+weekly_counts <- dbFetch(query,n=-1)
+dbClearResult(query)
 
 weekly_counts$Index <- seq_len(nrow(weekly_counts))
-
+# Find the Index for March 2020
+march_2020_index <- which(weekly_counts$YearWeek == "2020-09")  # Adjust '202009' to match your data's YearWeek format for March 2020
+
+# Plot
 plot(
@@
        col = "gray", lty = "dotted")
+
+# Add vertical line for March 2020
+abline(v = weekly_counts$Index[march_2020_index], col = "red", lwd = 2)
+
+# Add a label for March 2020
+text(
+  x = weekly_counts$Index[march_2020_index] - 80, 
+  y = 10, 
+  labels = "March 2020", 
+  pos = 3, 
+  col = "red"
+)
 ```
@@
 Figure 2 analyzes the number of crimes over time using time series.
-The time series shows stable weekly crime counts with occasional large spikes.
+The time series shows stable weekly crime counts with occasional large spikes, likely due to shorter weeks.
 These surges indicate certain weeks where crime reporting jumped significantly, there is a recent interesting surge that seems to mean-revert in 2024.
+The red line, which denotes March 2020, shows a decline in crimes, likely during the shutdown and quarantine, but it returned to normal levels quickly.
 
@@
 #crime_la$Date.Rptd
+crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
+# Extract the day of the week
 crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")
 
-# Extract the day of the week
 crime_la$DayOfWeek <- weekdays(crime_la$Date)
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Percent of crime victims by gender, 2010 to present", fig.align = "center"}
-# Remove missing values and empty strings in Vict.Sex column
-crime_la <- subset(crime_la, Vict.Sex != "" & !is.na(Vict.Sex))
-
-# Replace all non-"F" and non-"M" values with "N"
-crime_la$Vict.Sex <- ifelse(crime_la$Vict.Sex %in% c("F", "M"), 
-                                  crime_la$Vict.Sex, "N")
-
-# Count occurrences of each gender
-gender_counts <- table(crime_la$Vict.Sex)
+query <- dbSendQuery(con, "SELECT COUNT(*) As count, Modified_Vict_Sex As `Vict.Sex` FROM (
+    SELECT *,  
+        CASE 
+            WHEN `Vict.Sex` NOT IN ('M', 'F') THEN 'N'
+            ELSE `Vict.Sex`
+        END AS `Modified_Vict_Sex`
+    FROM crime_la c
+    WHERE `Vict.Sex` != '' AND `Vict.Sex` IS NOT NULL) s
+    GROUP BY Modified_Vict_Sex
+")
+gender_counts <- dbFetch(query,n=-1)
+dbClearResult(query)
 
@@
 pie(
-  gender_counts, 
-  labels = paste(names(gender_counts), 
-                 round(gender_counts / sum(gender_counts) * 100, 1), "%"), 
-  col = color[names(gender_counts)]
+  gender_counts$count, 
+  labels = paste(gender_counts$Vict.Sex, 
+                 round(gender_counts$count / sum(gender_counts$count) * 100, 1), "%"), 
+  col = color[gender_counts$Vict.Sex]
 )
@@
 zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
-LA_zctas <- read.csv("0_Data/Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs
+query <- dbSendQuery(con, "SELECT * FROM zctas_in_la")
+LA_zctas <- dbFetch(query,n=-1)
+dbClearResult(query)
 
@@
 # getting LA median income using joined file
-median_household_income <- cali_median_income %>% dplyr::select(ZCTA, S1903_C03_001E, year)
+query <- dbSendQuery(con, "SELECT ZCTA, S1903_C03_001E, year FROM cali_median_income")
+median_household_income <- dbFetch(query,n=-1)
+dbClearResult(query)
 
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Annual Median Income and Homicide Count by ZCTA in Los Angeles with top 1% income and top 1% homicide rates labelled", fig.align = "center"}
-cali_median_income$ZCTA5CE20 <- as.integer(cali_median_income$ZCTA)
-crime_income <- crimes_with_zips %>% inner_join(cali_median_income, by = c("year", "ZCTA5CE20"))
+zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
+query <- dbSendQuery(con, "SELECT i.*, c.*
+FROM cali_median_income i
+INNER JOIN crimes_with_zips c
+ON i.year = c.year AND CAST(i.ZCTA AS int) = c.ZCTA5CE20")
+crime_income <- dbFetch(query,n=-1)
+dbClearResult(query)
 
@@
 
-crime_income_type <- crime_income %>% inner_join(crime_type, by = c("year", "DR_NO"))
+crime_income_type <- crime_income[,-c(249, 247)] %>% inner_join(crime_type, join_by("year" == "year", "DR_NO" == "DR_NO"))
 
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025"}
+
+query <- dbSendQuery(con, "SELECT * FROM evictions_city_la")
+evictions_la <- dbFetch(query,n=-1)
+dbClearResult(query)
 # fix mistake
@@
 # create color scheme
-crime_la$month_year <- format(crime_la$DateParsed, "%Y-%m")
+crime_la$month_year <- format(crime_la$Date, "%Y-%m")
 crime_la_monthly <- crime_la %>% 
@@
 
-bureaus <- read.csv("0_Data/Raw_Data/Division_Bureau_Ref.csv")
+query <- dbSendQuery(con, "SELECT * FROM lapd_bureaus")
+bureaus <- dbFetch(query,n=-1)
+dbClearResult(query)
 
@@
 
+query <- dbSendQuery(con, "SELECT COUNT(*) FROM crime_la GROUP BY `AREA.NAME`, year")
+test <- dbFetch(query,)
 crime_by_area <- crime_by_area %>% left_join(bureaus, join_by("AREA.NAME" == "Division"))
@@
 # since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions.
-LA_county_evictions <- read.csv("0_Data/Raw_Data/LA_county_evictions.csv")
-
 # covid data
-covid_14day$date <- ymd(covid_14day$ep_date)
-covid_14day$month <- month(covid_14day$date)
-covid_14day$year <- year(covid_14day$date)
-covid_14day$date <- day(covid_14day$date)
+query <- dbSendQuery(con, "
+  SELECT 
+    CAST(strftime('%Y', ep_date) AS INTEGER) AS year,
+    CAST(strftime('%m', ep_date) AS INTEGER) AS month,
+   CAST(strftime('%d', ep_date) AS INTEGER) AS date, 
+   cases_14day
+  FROM covid_cases
+")
+covid_14day <- dbFetch(query,n=-1)
+dbClearResult(query)
 # since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
@@
 
-# next, unemployment rate
-unemployment <- read.csv("0_Data/Raw_Data/la_county_unemployment_rate.csv")
-unemployment$month <- as.numeric(gsub("M", "", unemployment$Period))
-
-# lapd budget
-lapd_budget <- read.csv("0_Data/Raw_Data/lapd_budget.csv")
+# next, unemployment rate, lapd budget, and evictions
+query <- dbSendQuery(con, "
+  SELECT 
+    u.month, 
+    u.Year, 
+    u.Value AS unemployment_rate, 
+    l.police_budget, 
+    e.Count AS num_evictions
+  FROM 
+    unemployment_la AS u
+  INNER JOIN 
+    lapd_budget AS l ON u.Year = l.year
+  INNER JOIN 
+    evictions_county_la AS e ON u.Year = e.Year AND u.month = e.Month
+")
+unemployment_lapd_budget_evictions <- dbFetch(query,n=-1)
+dbClearResult(query)
 
@@
 
-correlations_df <- crime_la_monthly %>% inner_join(unemployment, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, Value, num_crimes) %>% rename("unemployment_rate" = "Value")
+correlations_df <- crime_la_monthly %>% inner_join(unemployment_lapd_budget_evictions, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, num_evictions, unemployment_rate, num_crimes, police_budget)
 
@@
 
-correlations_df3 <- correlations_df2 %>% inner_join(LA_county_evictions, join_by("year" == "Year", "month" == "Month")) %>% rename("num_evictions" = "Count")
-
-correlations_df4 <- correlations_df3 %>% inner_join(lapd_budget, by = "year")
-
-correlations <- correlations_df4 %>% dplyr::select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)
+correlations <- correlations_df2 %>% dplyr::select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)
 
@@
 
-long_Cor <- reshape2::melt(Cor_mat)
-
-long_Cor <- long_Cor %>%
+# get only the upper triangle of variables 
+# (because a correlation heat map shows each correlation value twice)
+get_upper_tri <- function(cormat){
+  # lower.tri -> Returns a matrix of logicals, so set the values that are true to NA
+  cormat[lower.tri(cormat)]<- NA # set the correlation matrix lower triangle to NA
+  return(cormat) 
+}
+
+# reorder them by the strength of the correlation
+reorder_cormat <- function(cormat){
+  # Use correlation between variables as distance
+  dd <- as.dist((1-cormat)/2)
+  hc <- hclust(dd) # use hierarchical clustering to reorder
+  cormat <-cormat[hc$order, hc$order]
+}
+
+# use the functions above to remove the unnecessary parts
+Cor_mat<-reorder_cormat(Cor_mat)
+
+Cor_mat_upper<-get_upper_tri(Cor_mat)
+
+# using melt instead of pivot_longer because pivot_longer did not order correctly
+Cor_mat_upper<-reshape2::melt(Cor_mat_upper, na.rm=TRUE)
+
+# remove the central values 
+# which show the correlation for each variable with itself
+Cor_mat_upper<- Cor_mat_upper %>% filter(value != 1.00)
+
+long_Cor <- Cor_mat_upper %>%
     mutate(Var1 = recode(Var1, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget ($ billions)"),
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Homicide Victims by Age and Gender in Los Angeles 2010-Present"}
-homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% dplyr::select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex))
+homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% dplyr::select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex, levels  = c("F", "M", "X"), labels = c("Female", "Male", "Other"))) %>% drop_na()
 
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Time series analyzing homicide rate trends over time by median income quartiles"}
-# Data Preparation
-homicides <- crime_la %>% inner_join(crimes_with_zips, by = c("DR_NO", "LAT", "LON","X", "year")) %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% 
-  mutate(date = as.Date(DateParsed),
-         zip = as.character(ZCTA5CE20)) %>% dplyr::select(-X)
-
-homicides <- homicides %>% inner_join(cali_median_income, join_by("zip" == "ZCTA", "year" == "year"))
-homicides$S1903_C03_001E <- as.numeric(homicides$S1903_C03_001E)
+query <- dbSendQuery(con, "
+  SELECT i.S1903_C03_001E, x.year, i.ZCTA, x.LAT, x.LON
+  FROM (
+    SELECT c.LAT, c.LON, CAST(z.ZCTA5CE20 AS INTEGER) AS ZCTA5CE20, CAST(c.year AS INTEGER) AS year
+    FROM crime_la AS c
+    INNER JOIN crimes_with_zips AS z 
+      ON c.DR_NO = z.DR_NO 
+      AND c.LAT = z.LAT 
+      AND c.LON = z.LON 
+      AND c.X = z.X 
+      AND c.year = z.year
+    WHERE c.\"Crm.Cd.Desc\" = 'CRIMINAL HOMICIDE'
+  ) AS x INNER JOIN (SELECT m.S1903_C03_001E, CAST(m.year AS INTEGER) AS year, CAST(m.ZCTA AS INTEGER) AS ZCTA FROM cali_median_income m) AS i 
+    ON x.ZCTA5CE20 = i.ZCTA AND x.year = i.year
+")
+
+homicides <- dbFetch(query,n=-1)
+dbClearResult(query)
 
@@
 
+\begin{equation}
+\text{Crimes}_i = \beta_0 + \beta_1 \text{Index}_i + \beta_2 \text{CrimesLag}_i + \epsilon_i
+\end{equation}
+
 ```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
@@
 ## Regression Between Median Income and Crimes per Zip Code
+
+\begin{equation}
+\text{HomicideCount}_i = \beta_0 + \beta_1 \text{cali\_median\_income}_i + \epsilon_i
+\end{equation}
 
@@
 
+\begin{equation}
+\text{HomicideCount}_i = \beta_0 + \sum_{j=1}^{k} \beta_j \text{Factor}_j + \epsilon_i
+\end{equation}
+
 ```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.align = "center"}
 library(kableExtra)
-crimes_by_cb <- read.csv("0_Data/Cleaned_Data/crimes_by_census_block.csv")
-colnames(crimes_by_cb) <- c(colnames(crimes_by_cb)[-c(1, length(colnames(crimes_by_cb)))],"LON", "LAT")
-
-crimes_by_cb$LON <- as.numeric(gsub('c(', "", crimes_by_cb$LON, fixed = TRUE))
-crimes_by_cb$LAT <- as.numeric(gsub(')', "", crimes_by_cb$LAT, fixed = TRUE))
-
-crimes_by_cb$DateParsed <- as.POSIXct(crimes_by_cb$Date.Rptd, 
-                                  format = "%m/%d/%Y %I:%M:%S %p", 
-                                  tz = "America/Los_Angeles")
-
-crimes_by_cb$YearWeek <- format(crimes_by_cb$DateParsed, "%Y-%U")
-
-crimes_by_cb$year <- as.integer(stringr::str_sub(crimes_by_cb$YearWeek, 1, 4))
-
-homicides_by_cb <- crimes_by_cb %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% group_by(BLKGRPCE, year) %>% summarize(HomicideCount = n()) %>% ungroup()
+query <- dbSendQuery(con, "SELECT * FROM crimes_by_census_block")
+crimes_by_cb <- dbFetch(query,n=-1)
+dbClearResult(query)
+
+homicides_by_cb <- crimes_by_cb %>% filter(str_detect(Crm.Cd.Desc, "CRIMINAL HOMICIDE")) %>% group_by(BLKGRPCE, year) %>% summarize(HomicideCount = n()) %>% ungroup()
 
@@
 
+Model 1 is defined as: 
+\begin{equation}
+\text{HomicideCount}_i = \beta_0 + \beta_1 \text{LAT\_rounded}_i + \beta_2 \text{LON\_rounded}_i + \epsilon_i
+\end{equation}
+
+Model 2 is defined as: 
+\begin{equation}
+\text{HomicideCount}_i = \beta_0 + \beta_1 \text{year}_i + \epsilon_i
+\end{equation}
+
+Model 3 is defined as:
+\begin{equation}
+\text{HomicideCount}_i = \beta_0 + \beta_1 \text{LAT\_rounded}_i + \beta_2 \text{LON\_rounded}_i + \beta_3 \text{year}_i + \epsilon_i
+\end{equation}
+
+Model 4 is defined as:
+\begin{align}
+\text{HomicideCount}_i &= \beta_0 + \beta_1 \text{LAT\_rounded}_i + \beta_2 \text{LON\_rounded}_i + \beta_3 \text{year}_i \nonumber \\
+&\quad + \beta_4 (\text{LAT\_rounded}_i \times \text{LON\_rounded}_i) + \beta_5 (\text{LAT\_rounded}_i \times \text{year}_i) \nonumber \\
+&\quad + \beta_6 (\text{LON\_rounded}_i \times \text{year}_i) + \beta_7 (\text{LAT\_rounded}_i \times \text{LON\_rounded}_i \times \text{year}_i) + \epsilon_i
+\end{align}
+
 ```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
@@
 
-#### Data Wrangling {.unnumbered}
-
-##### Reading in Raw Datasets
-
-```{r DATA_Wrangling, output=FALSE, eval=FALSE}
+```{r}
+knitr::opts_chunk$set(eval = FALSE)
+```
+
+```{r, output=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
 library(sf) # for map
@@
 library(dplyr)
-
-# primary datasets
-# Primary datasets: Crime Data 2020-Present and 2010-2019
-crime_la_2010_2019 <- read.csv("0_Data/Raw_Data/Crime_Data_from_2010_to_2019.csv")
-
-crime_la_2020_present <- read.csv("0_Data/Raw_Data/Crime_Data_from_2020_to_Present.csv")
-
-crime_la <- read.csv("0_Data/Cleaned_Data/all_crimes_2010_present.csv")
-
-# primary dataset zip code and geometry data
-crimes_with_zips <- read.csv("0_Data/Cleaned_Data/crimes_with_zips.csv") # just inner_join with the crime_la dataset
-
-# Secondary dataset - Median Income by ZCTA
-cali_median_income <- read.csv("0_Data/Cleaned_Data/cali_median_income.csv")
-
-# adding a variable that's just the ZCTA zip code
-cali_median_income$ZCTA <- stringr::str_sub(cali_median_income$NAME, -5)
-
-# Secondary dataset - COVID cases in Los Angeles County
-covid_14day <- read.csv("0_Data/Raw_Data/LA_County_Covid19_CSA_14day_case_death_table.csv")
-
-evictions_la <- read.csv("0_Data/Raw_Data/LA_evictions.csv")
-
-crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
-                                  format = "%m/%d/%Y %I:%M:%S %p", 
-                                  tz = "America/Los_Angeles")
-
-crime_la$month <- lubridate::month(crime_la$DateParsed)
-
-crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
-
-crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
-```
-
-```{r Combining Median Income, eval=FALSE}
-# Secondary datasets
-california_median_income_2018_2023 <- list()
-yrs <- 2018:2023
-for (i in 1:length(yrs)) {
-  obj <- read.csv(paste0("0_Data/Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
-  obj <- obj[-1,]
-  obj$year <- yrs[[i]]
-  california_median_income_2018_2023[[i]] <- obj
-}
-
-cali_median_income <- do.call(rbind, california_median_income_2018_2023)
-
-# adding a variable that's just the ZCTA zip code
-cali_median_income$ZCTA <- stringr::str_sub(cali_median_income$NAME, -5)
-
-write.csv(cali_median_income, "cali_median_income.csv")
-```
-
-##### Zip Codes for Crimes
-
-```{r Get Zip Codes, eval=FALSE}
-library(sf)
-library(dplyr)
-
-crime_la_2010_2019 <- read.csv("0_Data/Raw_Data/Crime_Data_from_2010_to_2019.csv")
-
-crime_la_2020_present <- read.csv("0_Data/Raw_Data/Crime_Data_from_2020_to_Present.csv")
-
-# Combine the two datasets
-crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
-
-crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
-                                  format = "%m/%d/%Y %I:%M:%S %p", 
-                                  tz = "America/Los_Angeles")
-
-crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
-
-crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
-
-coords <- crime_la %>% dplyr::select(LAT, LON, DR_NO, year)
-
-zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
-
-  points <- st_as_sf(coords, coords = c("LON", "LAT"), crs = 4326)  # WGS84
-
-  zcta <- st_transform(zcta, 4326)  # Convert ZCTA to WGS84 if needed
-  cat("Starting Zip Code Retrieval \n")
-
-  result <- st_join(points, zcta)
-
-  cat("Finished Zip Code Retrieval \n")
-
-  write.csv(result, "0_Data/Raw_Data/all_zips.csv")
-
-crime_zctas <- read.csv("0_Data/Raw_Data/all_zips.csv")
-colnames(crime_zctas) <- c(colnames(crime_zctas)[-c(1, length(colnames(crime_zctas)))],"LON", "LAT")
-
-crime_zctas$LON <- as.numeric(gsub('c(', "", crime_zctas$LON, fixed = TRUE))
-crime_zctas$LAT <- as.numeric(gsub(')', "", crime_zctas$LAT, fixed = TRUE))
-
-# Repeat same process for Census Blocks
-```
-
-```{r Get Zip Codes 2, eval=FALSE}
-
-duplicates <- as.integer(as.character(data.frame(table(crime_zctas$universal_idx)[table(crime_zctas$universal_idx) > 1]) %>% pull(Var1)))
-
-# there are approximately 200 rows that are duplicated (100 incidents) as the area lies on the intersection between two zctas. 
-duplicate_data <- crime_zctas %>% filter(universal_idx %in% duplicates)
-
-rev_geo_duplicates <- reverse_geocode(duplicates2, lat = LAT, long = LON)
-# the above query identifies the address of these coordinates as KFC, 8644, Balboa Boulevard, Northridge South Neighborhood Council District, Los Angeles, Los Angeles County, California, 91325, United States.
-# therefore, we will only keep the 91325 zip code rows.
-
-crime_zctas_no_duplicates <- crime_zctas %>% mutate(duplicate = ifelse(universal_idx %in% duplicate_data$universal_idx & ZCTA5CE20 != 91325, "Drop", "Keep")) %>% filter(duplicate != "Drop") %>% dplyr::select(-duplicate)
-
-# it's empty now 
-# duplicates <- table(crime_zctas_no_duplicates$universal_idx)[table(crime_zctas_no_duplicates$universal_idx) > 1]
-
-# next, check for NAs
-sum(is.na(crime_zctas_no_duplicates$ZCTA5CE20)) # there are 3724 NAs. Let's find them a zip code
-
-# some also have lat long = 0. We'll find those separately
-no_zipcode_not0 <- crime_zctas_no_duplicates %>% filter(is.na(ZCTA5CE20) & LON != 0)
-
-no_zipcode_not0_zips <- reverse_geocode(no_zipcode_not0, lat = LAT, long = LON)
-
-no_zipcode_not0_zips$ZCTA5CE20 <- stringr::str_extract(no_zipcode_not0_zips$address, "\\d{5}")
-
-no_zipcode_not0_zips <- no_zipcode_not0_zips %>% dplyr::select(-address)
-
-crime_zctas2 <- crime_zctas_no_duplicates %>% filter(!is.na(ZCTA5CE20))
-
-crime_zctas3 <- rbind(crime_zctas2, no_zipcode_not0_zips)
-
-crime_zctas4 <- crime_zctas3 %>% filter(!is.na(ZCTA5CE20))
-
-write.csv(crime_zctas4, "crimes_with_zips.csv")
-
-no_zipcode_0s <- crime_zctas_no_duplicates %>% filter(is.na(ZCTA5CE20) & LON == 0) %>% pull(DR_NO)
-
-main_dataset_0s <- crime_la %>% filter(DR_NO %in% no_zipcode_0s)
-
-summary(main_dataset_0s)
-
-summary(crime_la)
-```
-
-```{r eval=FALSE, fig.cap="Crime densities in Los Angeles by LAPD Division, message=FALSE, warning=FALSE, 2010.2019 and 2020.Present", fig.align="center"}
+library(RSQLite)
+
+con <- dbConnect(SQLite(), "0_Data/crime_in_la.db")
+```
+
+# Exploratory Data Analysis
+
+```{r echo=FALSE, fig.cap="Crime densities in Los Angeles by LAPD Division, message=FALSE, warning=FALSE, 2010.2019 and 2020.Present", fig.align="center"}
+query <- dbSendQuery(con, "Select * From crime_la Where year < 2020")
+crime_la_2010_2019 <- dbFetch(query,n=-1)
+dbClearResult(query)
+
+query <- dbSendQuery(con, "Select * From crime_la Where year > 2019")
+crime_la_2020_present <- dbFetch(query,n=-1)
+dbClearResult(query)
+
 lapd_shape <- read_sf(dsn = "0_Data/Raw_Data/LAPD_Divisions", layer = "LAPD_Divisions")
@@
 
-```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.cap = "Weekly Crime Counts in Los Angeles from 2010 to present", fig.align = "center"}
-crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
-                                  format = "%m/%d/%Y %I:%M:%S %p", 
-                                  tz = "America/Los_Angeles")
-
-crime_la$month <- lubridate::month(crime_la$DateParsed)
-
-crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")
-
-weekly_counts <- aggregate(DR_NO ~ YearWeek, data = crime_la, FUN = length)
-names(weekly_counts)[2] <- "Crimes"
-
-weekly_counts <- weekly_counts[order(weekly_counts$YearWeek), ]
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap = "Weekly Crime Counts in Los Angeles from 2010 to present", fig.align = "center"}
+query <- dbSendQuery(con, "Select Count(*) as Crimes, YearWeek From crime_la Group By YearWeek Order By YearWeek")
+weekly_counts <- dbFetch(query,n=-1)
+dbClearResult(query)
 
 weekly_counts$Index <- seq_len(nrow(weekly_counts))
-
+# Find the Index for March 2020
+march_2020_index <- which(weekly_counts$YearWeek == "2020-09")  # Adjust '202009' to match your data's YearWeek format for March 2020
+
+# Plot
 plot(
@@
        col = "gray", lty = "dotted")
-```
-
-```{r, eval=FALSE, fig.cap = "Los Angeles Crimes by Day of the Week", warning=FALSE, message=FALSE, fig.align = "center"}
+
+# Add vertical line for March 2020
+abline(v = weekly_counts$Index[march_2020_index], col = "red", lwd = 2)
+
+# Add a label for March 2020
+text(
+  x = weekly_counts$Index[march_2020_index] - 80, 
+  y = 10, 
+  labels = "March 2020", 
+  pos = 3, 
+  col = "red"
+)
+```
+
+```{r, echo=FALSE, fig.cap = "Los Angeles Crimes by Day of the Week", warning=FALSE, message=FALSE, fig.align = "center"}
 # Convert Date.Rptd to Date format (ignore time part)
 #crime_la$Date.Rptd
+crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
+# Extract the day of the week
 crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")
 
-# Extract the day of the week
 crime_la$DayOfWeek <- weekdays(crime_la$Date)
@@
 ```
-```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.cap="Percent of crime victims by gender, 2010 to present", fig.align = "center"}
-# Remove missing values and empty strings in Vict.Sex column
-crime_la <- subset(crime_la, Vict.Sex != "" & !is.na(Vict.Sex))
-
-# Replace all non-"F" and non-"M" values with "N"
-crime_la$Vict.Sex <- ifelse(crime_la$Vict.Sex %in% c("F", "M"), 
-                                  crime_la$Vict.Sex, "N")
-
-# Count occurrences of each gender
-gender_counts <- table(crime_la$Vict.Sex)
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Percent of crime victims by gender, 2010 to present", fig.align = "center"}
+query <- dbSendQuery(con, "SELECT COUNT(*) As count, Modified_Vict_Sex As `Vict.Sex` FROM (
+    SELECT *,  
+        CASE 
+            WHEN `Vict.Sex` NOT IN ('M', 'F') THEN 'N'
+            ELSE `Vict.Sex`
+        END AS `Modified_Vict_Sex`
+    FROM crime_la c
+    WHERE `Vict.Sex` != '' AND `Vict.Sex` IS NOT NULL) s
+    GROUP BY Modified_Vict_Sex
+")
+gender_counts <- dbFetch(query,n=-1)
+dbClearResult(query)
 
@@
 
-```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
 zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
-LA_zctas <- read.csv("0_Data/Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs
+query <- dbSendQuery(con, "SELECT * FROM zctas_in_la")
+LA_zctas <- dbFetch(query,n=-1)
+dbClearResult(query)
 
@@
 
-```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.cap="Median household income by year and by ZIP Code Tabulation Area (ZCTA) in Los Angeles", fig.align = "center"}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Median household income by year and by ZIP Code Tabulation Area (ZCTA) in Los Angeles", fig.align = "center"}
 # quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
@@
 
-```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.cap="Annual Median Income and Homicide Count by ZCTA in Los Angeles with top 1% income and top 1% homicide rates labelled", fig.align = "center"}
-cali_median_income$ZCTA5CE20 <- as.integer(cali_median_income$ZCTA)
-crime_income <- crimes_with_zips %>% inner_join(cali_median_income, by = c("year", "ZCTA5CE20"))
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Annual Median Income and Homicide Count by ZCTA in Los Angeles with top 1% income and top 1% homicide rates labelled", fig.align = "center"}
+zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
+query <- dbSendQuery(con, "SELECT i.*, c.*
+FROM cali_median_income i
+INNER JOIN crimes_with_zips c
+ON i.year = c.year AND CAST(i.ZCTA AS int) = c.ZCTA5CE20")
+crime_income <- dbFetch(query,n=-1)
+dbClearResult(query)
 
@@
 
-```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025"}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025"}
+
+query <- dbSendQuery(con, "SELECT * FROM evictions_city_la")
+evictions_la <- dbFetch(query,n=-1)
+dbClearResult(query)
 # fix mistake
@@
 
-```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Annual Number of Crimes by Los Angeles Police Department Bureau"}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Annual Number of Crimes by Los Angeles Police Department Bureau"}
 crime_la$year <- lubridate::year(crime_la$DateParsed)
@@
 
-```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Strength of Correlations Between the Number of Crimes and Different Law Enforcement and Socioeconomic Factors"}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Strength of Correlations Between the Number of Crimes and Different Law Enforcement and Socioeconomic Factors"}
 library(lubridate)
@@
 # since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions. 
-LA_county_evictions <- read.csv("0_Data/Raw_Data/LA_county_evictions.csv")
+query <- dbSendQuery(con, "SELECT * FROM evictions_county_la")
+LA_county_evictions <- dbFetch(query,n=-1)
+dbClearResult(query)
 
 # covid data
-covid_14day$date <- ymd(covid_14day$ep_date)
-covid_14day$month <- month(covid_14day$date)
-covid_14day$year <- year(covid_14day$date)
-covid_14day$date <- day(covid_14day$date)
+query <- dbSendQuery(con, "
+  SELECT 
+    CAST(strftime('%Y', ep_date) AS INTEGER) AS year,
+    CAST(strftime('%m', ep_date) AS INTEGER) AS month,
+   CAST(strftime('%d', ep_date) AS INTEGER) AS date, 
+   cases_14day
+  FROM covid_cases
+")
+covid_14day <- dbFetch(query,n=-1)
+dbClearResult(query)
 # since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
@@
 # next, unemployment rate
-unemployment <- read.csv("0_Data/Raw_Data/la_county_unemployment_rate.csv")
-unemployment$month <- as.numeric(gsub("M", "", unemployment$Period))
-
-# lapd budget
-lapd_budget <- read.csv("0_Data/Raw_Data/lapd_budget.csv")
+query <- dbSendQuery(con, "
+  SELECT 
+    u.month, 
+    u.Year, 
+    u.Value, 
+    l.police_budget
+  FROM unemployment_la AS u
+  INNER JOIN lapd_budget AS l ON u.Year = l.year
+")
+unemployment_lapd_budget <- dbFetch(query,n=-1)
+dbClearResult(query)
 
@@
 
-correlations_df <- crime_la_monthly %>% inner_join(unemployment, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, Value, num_crimes) %>% rename("unemployment_rate" = "Value")
+correlations_df <- crime_la_monthly %>% inner_join(unemployment_lapd_budget, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, Value, num_crimes, police_budget) %>% rename("unemployment_rate" = "Value")
 
@@
 
-correlations_df4 <- correlations_df3 %>% inner_join(lapd_budget, by = "year")
-
-correlations <- correlations_df4 %>% dplyr::select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)
+correlations <- correlations_df3 %>% dplyr::select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)
 
@@
 
-```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Homicide Victims by Age and Gender in Los Angeles 2010-Present"}
-homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% dplyr::select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex))
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Homicide Victims by Age and Gender in Los Angeles 2010-Present"}
+homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% dplyr::select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex, levels  = c("F", "M", "X"), labels = c("Female", "Male", "Other"))) %>% drop_na()
 
@@
 
-```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Frequency of Crimes in Los Angeles Throughout the Day"}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Frequency of Crimes in Los Angeles Throughout the Day"}
 crime_la$time <- sprintf("%04d",crime_la$TIME.OCC)
@@
 
-```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Time series analyzing homicide rate trends over time by median income quartiles"}
-# Data Preparation
-homicides <- crime_la %>% inner_join(crimes_with_zips, by = c("DR_NO", "LAT", "LON","X", "year")) %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% 
-  mutate(date = as.Date(DateParsed),
-         zip = as.character(ZCTA5CE20)) %>% dplyr::select(-X)
-
-homicides <- homicides %>% inner_join(cali_median_income, join_by("zip" == "ZCTA", "year" == "year"))
-homicides$S1903_C03_001E <- as.numeric(homicides$S1903_C03_001E)
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Time series analyzing homicide rate trends over time by median income quartiles"}
+query <- dbSendQuery(con, "
+  SELECT i.S1903_C03_001E, x.year, i.ZCTA, x.LAT, x.LON
+  FROM (
+    SELECT c.LAT, c.LON, CAST(z.ZCTA5CE20 AS INTEGER) AS ZCTA5CE20, CAST(c.year AS INTEGER) AS year
+    FROM crime_la AS c
+    INNER JOIN crimes_with_zips AS z 
+      ON c.DR_NO = z.DR_NO 
+      AND c.LAT = z.LAT 
+      AND c.LON = z.LON 
+      AND c.X = z.X 
+      AND c.year = z.year
+    WHERE c.\"Crm.Cd.Desc\" = 'CRIMINAL HOMICIDE'
+  ) AS x INNER JOIN (SELECT m.S1903_C03_001E, CAST(m.year AS INTEGER) AS year, CAST(m.ZCTA AS INTEGER) AS ZCTA FROM cali_median_income m) AS i 
+    ON x.ZCTA5CE20 = i.ZCTA AND x.year = i.year
+")
+
+homicides <- dbFetch(query,n=-1)
+dbClearResult(query)
 
@@
 
-```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
 library(broom)
@@
 
-```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
 lm_income_crimes <- lm(HomicideCount ~ cali_median_income, data = homicide_income_by_year)
@@
 
-```{r, warning=FALSE, message=FALSE, eval=FALSE, fig.align = "center"}
+```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.align = "center"}
 library(kableExtra)
-crimes_by_cb <- read.csv("0_Data/Cleaned_Data/crimes_by_census_block.csv")
-colnames(crimes_by_cb) <- c(colnames(crimes_by_cb)[-c(1, length(colnames(crimes_by_cb)))],"LON", "LAT")
-
-crimes_by_cb$LON <- as.numeric(gsub('c(', "", crimes_by_cb$LON, fixed = TRUE))
-crimes_by_cb$LAT <- as.numeric(gsub(')', "", crimes_by_cb$LAT, fixed = TRUE))
-
-crimes_by_cb$DateParsed <- as.POSIXct(crimes_by_cb$Date.Rptd, 
-                                  format = "%m/%d/%Y %I:%M:%S %p", 
-                                  tz = "America/Los_Angeles")
-
-crimes_by_cb$YearWeek <- format(crimes_by_cb$DateParsed, "%Y-%U")
-
-crimes_by_cb$year <- as.integer(stringr::str_sub(crimes_by_cb$YearWeek, 1, 4))
-
-homicides_by_cb <- crimes_by_cb %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% group_by(BLKGRPCE, year) %>% summarize(HomicideCount = n()) %>% ungroup()
+query <- dbSendQuery(con, "SELECT * FROM crimes_by_census_block")
+crimes_by_cb <- dbFetch(query,n=-1)
+dbClearResult(query)
+
+homicides_by_cb <- crimes_by_cb %>% filter(str_detect(Crm.Cd.Desc, "CRIMINAL HOMICIDE")) %>% group_by(BLKGRPCE, year) %>% summarize(HomicideCount = n()) %>% ungroup()
 
@@
 pretty_labels <- c("(Intercept)", paste("Census Block", 2:6))
-sjPlot::tab_model(model, pred.labels = pretty_labels, dv.labels = "Table 3: Annual Homicide Count by Census Block")
-```
-
-```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+sjPlot::tab_model(model, pred.labels = pretty_labels, dv.labels = "Table 3: Annual Homicide Count by Census Block", show.est = TRUE)
+```
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
 homicides_latlon_rounded <- homicides %>%
@@
 
-```{r, eval=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
 

### 2025-04-15 Slides.qmd
Slides files

---
title: "Revealjs presentation framework"
subtitle: "Module 11"
format:
  revealjs:
# embed-resources: true
    slide-number: true
    chalkboard:
      buttons: false
    preview-links: auto
    css: styles.css
    footer: Group 5 - STAT 405/605 - Bianca Schutz, Kevin Hu, Kaitlyn Kirt, Marco Stine - Department of Statistics, Rice University - Source [quarto.org/revealjs](https://quarto.org/docs/presentations/revealjs/)
server: shiny
---

# Section  {background="#6B6B6B"}

## Source
This presentation is a subset of the demo created by quarto, plus other material.

The complete source of the demo can be found following [this link](https://quarto.org/docs/presentations/revealjs/).

## Hello, There

This presentation will show you examples of what you can do with Quarto and [Reveal.js](https://revealjs.com), including:

-   Presenting code and LaTeX equations
-   Including computations in slide output
-   Image, video, and iframe backgrounds
-   Fancy transitions and animations
-   Printing to PDF

...and much more

## Pretty Code {auto-animate="true"}

-   Over 20 syntax highlighting themes available
-   Default theme optimized for accessibility

``` r
# Define a server for the Shiny app
function(input, output) {

  # Fill in the spot we created for a plot
  output$phonePlot <- renderPlot({
    # Render a barplot
  })
}
```

::: footer
Learn more: [Syntax Highlighting](https://quarto.org/docs/output-formats/html-code.html#highlighting)
:::

## Code Animations {auto-animate="true"}

-   Over 20 syntax highlighting themes available
-   Default theme optimized for accessibility

``` r
# Define a server for the Shiny app
function(input, output) {

  # Fill in the spot we created for a plot
  output$phonePlot <- renderPlot({
    # Render a barplot
    barplot(WorldPhones[,input$region]*1000, 
            main=input$region,
            ylab="Number of Telephones",
            xlab="Year")
  })
}
```

::: footer
Learn more: [Code Animations](https://quarto.org/docs/presentations/revealjs/advanced.html#code-animations)
:::

## Line Highlighting

-   Highlight specific lines for emphasis
-   Incrementally highlight additional lines

``` {.python code-line-numbers="4-5|7|10"}
import numpy as np
import matplotlib.pyplot as plt

r = np.arange(0, 2, 0.01)
theta = 2 * np.pi * r
fig, ax = plt.subplots(subplot_kw={'projection': 'polar'})
ax.plot(theta, r)
ax.set_rticks([0.5, 1, 1.5, 2])
ax.grid(True)
plt.show()
```

::: footer
Learn more: [Line Highlighting](https://quarto.org/docs/presentations/revealjs/#line-highlighting)
:::

## Executable Code

```{r}
#| echo: true
#| fig-width: 10
#| fig-height: 4.5
library(ggplot2)
ggplot(mtcars, aes(hp, mpg, color = am)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "loess")
```

::: footer
Learn more: [Executable Code](https://quarto.org/docs/presentations/revealjs/#executable-code)
:::

## LaTeX Equations

[MathJax](https://www.mathjax.org/) rendering of equations to HTML

::: columns
::: {.column width="40%"}
``` tex
\begin{gather*}
a_1=b_1+c_1\\
a_2=b_2+c_2-d_2+e_2
\end{gather*}

\begin{align}
a_{11}& =b_{11}&
  a_{12}& =b_{12}\\
a_{21}& =b_{21}&
  a_{22}& =b_{22}+c_{22}
\end{align}
```
:::

::: {.column width="60%"}
```{=tex}
\begin{gather*}
a_1=b_1+c_1\\
a_2=b_2+c_2-d_2+e_2
\end{gather*}
```
```{=tex}
\begin{align}
a_{11}& =b_{11}&
  a_{12}& =b_{12}\\
a_{21}& =b_{21}&
  a_{22}& =b_{22}+c_{22}
\end{align}
```
:::
:::

::: footer
Learn more: [LaTeX Equations](https://quarto.org/docs/authoring/markdown-basics.html#equations)
:::

## Column Layout {.smaller}

Arrange content into columns of varying widths:

::: columns
::: {.column width="35%"}
#### Motor Trend Car Road Tests

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles.
:::

::: {.column width="3%"}
:::

::: {.column width="62%"}
```{r}
knitr::kable(head(mtcars)[,c("mpg",	"cyl", "disp", "hp", "wt")])
```
:::
:::

::: footer
Learn more: [Multiple Columns](https://quarto.org/docs/presentations/revealjs/#multiple-columns)
:::

## Incremental Lists

Lists can optionally be displayed incrementally:

::: incremental
-   First item
-   Second item
-   Third item
:::

. . .

<br/> Insert pauses to make other types of content display incrementally.

::: footer
Learn more: [Incremental Lists](https://quarto.org/docs/presentations/revealjs/#incremental-lists)
:::

## Fragments

Incremental text display and animation with fragments:

<br/>

::: {.fragment .fade-in}
Fade in
:::

::: {.fragment .fade-up}
Slide up while fading in
:::

::: {.fragment .fade-left}
Slide left while fading in
:::

::: {.fragment .fade-in-then-semi-out}
Fade in then semi out
:::

. . .

::: {.fragment .strike}
Strike
:::

::: {.fragment .highlight-red}
Highlight red
:::

::: footer
Learn more: [Fragments](https://quarto.org/docs/presentations/revealjs/advanced.html#fragments)
:::

## Slide Backgrounds {background="#43464B"}

Set the `background` attribute on a slide to change the background color (all CSS color formats are supported).

Different background transitions are available via the `background-transition` option.

::: footer
Learn more: [Slide Backgrounds](https://quarto.org/docs/presentations/revealjs/#color-backgrounds)
:::

## Media Backgrounds {background="#43464B" background-image="images/milky-way.jpeg"}

You can also use the following as a slide background:

-   An image: `background-image`

-   A video: `background-video`

-   An iframe: `background-iframe`

::: footer
Learn more: [Media Backgrounds](https://quarto.org/docs/presentations/revealjs/#image-backgrounds)
:::

## Auto-Animate {auto-animate="true" auto-animate-easing="ease-in-out"}

Automatically animate matching elements across slides with Auto-Animate.

::: r-hstack
::: {data-id="box1" auto-animate-delay="0" style="background: #2780e3; width: 200px; height: 150px; margin: 10px;"}
:::

::: {data-id="box2" auto-animate-delay="0.1" style="background: #3fb618; width: 200px; height: 150px; margin: 10px;"}
:::

::: {data-id="box3" auto-animate-delay="0.2" style="background: #e83e8c; width: 200px; height: 150px; margin: 10px;"}
:::
:::

::: footer
Learn more: [Auto-Animate](https://quarto.org/docs/presentations/revealjs/advanced.html#auto-animate)
:::

## Auto-Animate {auto-animate="true" auto-animate-easing="ease-in-out"}

Automatically animate matching elements across slides with Auto-Animate.

::: r-stack
::: {data-id="box1" style="background: #2780e3; width: 350px; height: 350px; border-radius: 200px;"}
:::

::: {data-id="box2" style="background: #3fb618; width: 250px; height: 250px; border-radius: 200px;"}
:::

::: {data-id="box3" style="background: #e83e8c; width: 150px; height: 150px; border-radius: 200px;"}
:::
:::

::: footer
Learn more: [Auto-Animate](https://quarto.org/docs/presentations/revealjs/advanced.html#auto-animate)
:::

## Slide Transitions {.smaller}

The next few slides will transition using the `slide` transition

| Transition | Description                                                            |
|------------|------------------------------------------------------------------------|
| `none`     | No transition (default, switch instantly)                              |
| `fade`     | Cross fade                                                             |
| `slide`    | Slide horizontally                                                     |
| `convex`   | Slide at a convex angle                                                |
| `concave`  | Slide at a concave angle                                               |
| `zoom`     | Scale the incoming slide so it grows in from the center of the screen. |

::: footer
Learn more: [Slide Transitions](https://quarto.org/docs/presentations/revealjs/advanced.html#slide-transitions)
:::

## Tabsets {.smaller .scrollable transition="slide"}

::: panel-tabset
### Plot

```{r}
library(ggplot2)
ggplot(mtcars, aes(hp, mpg, color = am)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "loess")
```

### Data

```{r}
knitr::kable(mtcars)
```
:::

::: footer
Learn more: [Tabsets](https://quarto.org/docs/presentations/revealjs/#tabsets)
:::

## Interactive Slides {.smaller transition="slide"}

Include Jupyter widgets and htmlwidgets in your presentations

```{r}
#| echo: false
#| fig-height: 5
library(leaflet)
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=174.768, lat=-36.852, popup="The birthplace of R")
```

::: footer
Learn more: [Jupyter widgets](https://quarto.org/docs/interactive/widgets/jupyter.html), [htmlwidgets](https://quarto.org/docs/interactive/widgets/htmlwidgets.html)
:::

## Preview Links

Navigate to hyperlinks without disrupting the flow of your presentation.

Use the `preview-links` option to open links in an iframe on top of your slides. Try clicking the link below for a demonstration:

::: {style="text-align: center; margin-top: 1em"}
[quarto.org/revealjs](https://quarto.org/docs/presentations/revealjs/){preview-link="true" style="text-align: center"}
:::

::: footer
Learn more: [Preview Links](https://quarto.org/docs/presentations/revealjs/presenting.html#preview-links)
:::

# Shiny examples

## Shiny example 1

::: columns
::: {.column width="25%"}
::: {style="font-size: 40%"}
```{r}
sliderInput("bins", "Number of bins:", 
            min = 1, max = 50, value = 30)
```
:::
:::

::: {.column width="75%"}
```{r}
plotOutput("distPlot", height = "500px")
```
:::
:::

```{r}
#| context: server
output$distPlot <- renderPlot({
   x <- faithful[, 2]  # Old Faithful Geyser data
   bins <- seq(min(x), max(x), length.out = input$bins + 1)
   hist(x, breaks = bins, col = 'darkgray', border = 'white',
        xlab = 'Waiting time to next eruption (in mins)',
        main = 'Histogram of waiting times')
}, res = 120)
```

## Shiny example 2

::: {style="font-size: 40%"}

```{r}
#| panel: sidebar
vars <- setdiff(names(iris), "Species")
selectInput('xcol', 'X Variable', vars)
selectInput('ycol', 'Y Variable', vars, selected = vars[[2]])
numericInput('clusters', 'Cluster count', 3, min = 1, max = 9)
```
:::

```{r}
#| panel: fill
plotOutput('plot1')
```

```{r}
#| context: server
selectedData <- reactive({
    iris[, c(input$xcol, input$ycol)]
  })
clusters <- reactive({
  kmeans(selectedData(), input$clusters)
})
output$plot1 <- renderPlot({
  palette(c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3",
    "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))
  par(mar = c(5.1, 4.1, 0, 1))
  plot(selectedData(),
       col = clusters()$cluster,
       pch = 20, cex = 3)
  points(clusters()$centers, pch = 4, cex = 4, lwd = 4)
})
```

## Shiny example 3

```{r}
#| context: setup
library(ggplot2)
dataset <- diamonds
```

```{r}
#| panel: fill
plotOutput('plot')
```

::: {style="font-size: 40%"}

::: {layout-ncol="3"}
```{r}
sliderInput('sampleSize', 'Sample Size', 
            min=1, max=nrow(dataset),
            value=min(1000, nrow(dataset)), 
            step=500, round=0)
br()
checkboxInput('jitter', 'Jitter')
checkboxInput('smooth', 'Smooth')
```

```{r}
selectInput('x', 'X', names(dataset))
selectInput('y', 'Y', names(dataset), names(dataset)[[2]])
selectInput('color', 'Color', c('None', names(dataset)))
```

```{r}
selectInput('facet_row', 'Facet Row',
            c(None='.', names(diamonds[sapply(diamonds, is.factor)])))
selectInput('facet_col', 'Facet Column',
            c(None='.', names(diamonds[sapply(diamonds, is.factor)])))
```
:::
:::

```{r}
#| context: server
dataset <- reactive({
  diamonds[sample(nrow(diamonds), input$sampleSize),]
})

output$plot <- renderPlot({

  p <- ggplot(dataset(), aes_string(x=input$x, y=input$y)) + geom_point()

  if (input$color != 'None')
    p <- p + aes_string(color=input$color)

  facets <- paste(input$facet_row, '~', input$facet_col)
  if (facets != '. ~ .')
    p <- p + facet_grid(facets)

  if (input$jitter)
    p <- p + geom_jitter()
  if (input$smooth)
    p <- p + geom_smooth()

  print(p)

}, res = 120)
```

### 2025-04-15 Slides/Slides.qmd
added slides files

---
title: "Crime in Los Angeles: Homicide Increases in the COVID-19 Era"
subtitle: "Bianca Schutz, Kevin Hu, Kaitlyn Kirt, Marco Stine"
format:
  revealjs:
# embed-resources: true
    slide-number: true
    chalkboard:
      buttons: false
    preview-links: auto
    css: styles.css
    footer: Group 5 - STAT 405/605 - Bianca Schutz, Kevin Hu, Kaitlyn Kirt, Marco Stine - Department of Statistics, Rice University - Source [quarto.org/revealjs](https://quarto.org/docs/presentations/revealjs/)
server: shiny
---

# Introduction {background="#6B6B6B"}

## Why research crime in Los Angeles?

::: incremental
-   Homicide rates in Los Angeles increased substantially during pandemic
-   One of the largest increases in the country, alongside Houston and Philadelphia
-   Caused by pre-pandemic and pandemic-induced economic instability and inequalities
-   Los Angeles makes their crime data publicly available (2010-Present)
:::

::: footer
Learn more: [Thurgood Marshall Institute Report](https://www.naacpldf.org/the-truth-behind-crime-statistics/)
:::

## Our hypothesis

We predicted that:

::: incremental
-   Socioeconomic conditions have improved in Los Angeles since the pandemic

-   As these socioeconomic measures have bounced back, the rate of homicide has returned to a more "normal" rate

-   A similar trend might be seen with overall crime rates in LA

-   We also examined if increased police funding is correlated with a reduction in the number of homicides.
:::

## Primary Data

Our primary datasets are the "Crime Data from 2010 to 2019" and "Crime Data from 2020 to Present" National Incident-Based Reporting System (NIBRS) datasets, available on [Data.gov](https://catalog.data.gov/).

```{r, echo=FALSE}
library(ggplot2)
library(RSQLite)
library(dplyr)
con <- dbConnect(RSQLite::SQLite(), normalizePath("la_stars.db"))

crime_la_annually <- dbGetQuery(con, "SELECT * FROM crime_la_annually") %>% group_by(year) %>% summarize(obs = first(total_crimes))

dbDisconnect(con)

ggplot(crime_la_annually, aes(x = year, y = obs/1000)) +
  geom_col(fill = "#19297C") +
  ggthemes::theme_fivethirtyeight() +
  labs(title = "Number of Annual Observations in Primary Dataset",
       y = "Thousands of Observations (Crimes)",
       x = "Year") +
  theme(axis.title = element_text())
```

## Secondary

Our supplementary datasets were:

-   [Los Angeles County Eviction Data](https://calmatters.org/housing/homelessness/2023/11/california-evictions-post-pandemic/)

-   [City of Los Angeles Eviction Data](https://housing.lacity.gov/residents/renters/eviction-notices-filed)

-   [Los Angeles County COVID Data](http://dashboard.publichealth.lacounty.gov/covid19_surveillance_dashboard/)

-   [Census Bureau Median Income Data](https://data.census.gov/table/ACSST1Y2023.S1903?q=income)

-   [Los Angeles Police Department Budget Data](https://openbudget.lacity.org/#!/year/2023/operating/0/department_name/Police/0/program_name?vis=lineChart)

-   [Los Angeles Area Unemployment Rate Data](https://www.bls.gov/regions/west/ca_losangeles_msa.htm#eag)

# Data Exploration {background="#6B6B6B"}

## Homicides Over Time by Income Quartile

```{r}
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

query <- dbSendQuery(con, "
  SELECT i.S1903_C03_001E, x.year, i.ZCTA, x.LAT, x.LON
  FROM (
    SELECT c.LAT, c.LON, CAST(z.ZCTA5CE20 AS INTEGER) AS ZCTA5CE20, CAST(c.year AS INTEGER) AS year
    FROM crime_la AS c
    INNER JOIN crimes_with_zips AS z 
      ON c.DR_NO = z.DR_NO 
      AND c.LAT = z.LAT 
      AND c.LON = z.LON 
      AND c.X = z.X 
      AND c.year = z.year
    WHERE c.\"Crm.Cd.Desc\" = 'CRIMINAL HOMICIDE'
  ) AS x INNER JOIN (SELECT m.S1903_C03_001E, CAST(m.year AS INTEGER) AS year, CAST(m.ZCTA AS INTEGER) AS ZCTA FROM cali_median_income m) AS i 
    ON x.ZCTA5CE20 = i.ZCTA AND x.year = i.year
")

homicides <- dbFetch(query,n=-1)
dbClearResult(query)
dbDisconnect(con)

# Form income quartiles
homicides <- homicides %>% mutate(income_quartile = ntile(S1903_C03_001E, 4)) 
homicide_trends <- homicides %>% group_by(year, income_quartile) %>% summarize(homicide_count = n(), 
          ref = mean(S1903_C03_001E),
          .groups = "drop")

homicide_trends$income_quartile <- as.factor(homicide_trends$income_quartile)
# removing NAs
homicide_trends <- homicide_trends %>% filter(!is.na(homicide_count) & !is.na(income_quartile))

# creating labels
levels(homicide_trends$income_quartile) <- c("Low Income", "Lower-Middle Income", "Higher-Middle Income", "High Income")
# Plot homicide trends over time by income quartile
ggplot(homicide_trends, aes(x = year, y = homicide_count)) +
  geom_area(aes(fill = as.factor(income_quartile)), alpha = 0.6, position = "identity") +
  labs(
    x = "Year",
    y = "Number of Homicides",
    fill = "Income Quartile"
  ) +
  theme_minimal() +
  scale_fill_viridis_d(option = "plasma", direction = -1)
```

## Homicides and Median Income By Zip Code
```{r, echo=FALSE}
library(RSQLite)
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

query <- dbSendQuery(con, "SELECT i.*, c.*
FROM cali_median_income i
INNER JOIN crimes_with_zips c
ON i.year = c.year AND CAST(i.ZCTA AS int) = c.ZCTA5CE20")
crime_income <- dbFetch(query,n=-1)
dbClearResult(query)

query <- dbSendQuery(con, "SELECT `Crm.Cd.Desc`, year, DR_NO FROM crime_la")
crime_la <- dbFetch(query,n=-1)
dbClearResult(query)

dbDisconnect(con)

crime_type <- crime_la %>% dplyr::select(year, Crm.Cd.Desc, DR_NO)

crime_income_type <- crime_income[,-c(249, 247)] %>% inner_join(crime_type, join_by("year" == "year", "DR_NO" == "DR_NO"))

homicide_income_by_year <- crime_income_type %>% 
  filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>%
  group_by(year, ZCTA5CE20) %>%
  summarize(HomicideCount = n(),
            cali_median_income = median(as.numeric(S1903_C03_001E))) %>%
  ungroup()

homicide_income_by_year <- homicide_income_by_year %>% filter(!is.na(HomicideCount) & !is.na(cali_median_income))

homicide_income_by_year$label <- ifelse(
  homicide_income_by_year$cali_median_income > quantile(homicide_income_by_year$cali_median_income, 0.99) | 
  homicide_income_by_year$HomicideCount > quantile(homicide_income_by_year$HomicideCount, 0.99), 
  homicide_income_by_year$ZCTA5CE20, 
  "")

ggplot(data = homicide_income_by_year, aes(x = cali_median_income, 
                                        y = HomicideCount)) +
  geom_point() +
  ggrepel::geom_text_repel(aes(label = label), size = 2.5, nudge_x = 2, nudge_y = 1) + 
  labs(x = "Median Annual Income ($)",
       y = "Annual Homicide Count",
       color = "Year") +
  ggthemes::theme_fivethirtyeight() + 
  theme(axis.title = element_text())
```

## Geographical Trends in Crime Rates By LAPD Bureau
```{r}
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

query <- dbSendQuery(con, "SELECT `AREA.NAME`, year, COUNT(*) AS numberOfCrimes FROM crime_la GROUP BY `AREA.NAME`, year")
crime_by_area <- dbFetch(query,n=-1)
dbClearResult(query)

query <- dbSendQuery(con, "SELECT * FROM lapd_bureaus")
bureaus <- dbFetch(query,n=-1)
dbClearResult(query)
dbDisconnect(con)

bureaus$Division <- gsub("77th", "77th Street", bureaus$Division)
bureaus$Division <- gsub("North", "N", bureaus$Division)
bureaus$Division <- gsub("Neast", "Northeast", bureaus$Division)

crime_by_area <- crime_by_area %>% left_join(bureaus, join_by("AREA.NAME" == "Division"))

ggplot(crime_by_area, aes(y = numberOfCrimes, group = Bureau, x = Bureau)) + 
        geom_boxplot(aes(fill = Bureau)) +
  labs(y = "Annual Number of Crimes")
```

## Evictions and Crime Levels

```{r}
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

query <- dbSendQuery(con, "SELECT * FROM evictions_city_la")
evictions_la <- dbFetch(query,n=-1)
dbClearResult(query)

query <- dbSendQuery(con, "SELECT `Crm.Cd.Desc`, year, month, DR_NO FROM crime_la")
crime_la <- dbFetch(query,n=-1)
dbClearResult(query)

dbDisconnect(con)
# fix mistake
evictions_la$Date.Received <- gsub("2205", "2025", evictions_la$Date.Received)

# convert to date
evictions_la$month_year <- format(as.Date(evictions_la$Date.Received, format="%m-%d-%Y"), "%Y-%m")

# get unique months
unique_months <- data.frame(month_year = sort(unique(evictions_la$month_year)), index = 1:length(unique(evictions_la$month_year)))

evictions_la_indexed <- evictions_la %>% left_join(unique_months, by = "month_year")

# create color scheme
crime_la$month_year <- paste0(crime_la$year, "-", sprintf("%02d", crime_la$month))

crime_la_monthly <- crime_la %>% 
  group_by(month_year) %>% 
  summarize(num_crimes = n())

evictions_la_indexed_crimes <- evictions_la_indexed %>% left_join(crime_la_monthly, by = "month_year")

avg_crimes <- mean(evictions_la_indexed_crimes$num_crimes)

evictions_la_indexed_crimes$high_low_crime <- as.factor(ifelse(evictions_la_indexed_crimes$num_crimes > avg_crimes, "Above Average", "Below Average"))

ggplot(data = evictions_la_indexed_crimes, aes(x = index)) + 
  geom_histogram(aes(fill = high_low_crime), color = "white", bins = nrow(unique_months)) +
  scale_x_continuous(breaks=unique_months$index,
        labels=unique_months$month_year) + 
  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", fill = "Crime Level") +
  ggthemes::scale_fill_fivethirtyeight() + 
  ggthemes::theme_fivethirtyeight() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.title.y = element_text(),
        axis.title.x = element_text(vjust = -.75),
        plot.title = element_text(size=12))
```

## Geographical Distribution of Homicides
:::::: columns
:::: {.column width="35%"}
::: {style="font-size: 50%"}
```{r}
sliderInput("selected_year", "Select Year:",
                  min = 2010, max = 2024,
                  value = 2010, sep = "", step= 1)
```
:::
::::

::: {.column width="65%"}
```{r}
library(leaflet)
leafletOutput("crime_map", height = 600)
```
:::
::::::

```{r}
#| context: server
library(leaflet)
library(RSQLite)
library(shiny)
library(leaflet)
library(DBI)
library(RSQLite)
library(dplyr)

crime_data <- reactive({
    req(input$selected_year)
    con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))
    crime_la_coords <- dbGetQuery(con, paste("SELECT LAT, LON, year FROM crime_la WHERE `Crm.Cd.Desc` = 'CRIMINAL HOMICIDE' AND year = ", input$selected_year))
    dbDisconnect(con)
    return(crime_la_coords)
  })

output$crime_map <- renderLeaflet({ 
  filtered <- crime_data()
  leaflet() %>%
  addTiles(
    urlTemplate = "https://{s}.basemaps.cartocdn.com/light_all/{z}/{x}/{y}{r}.png",
    attribution = '&copy; <a href="https://www.openstreetmap.org/copyright">OpenStreetMap</a> contributors &copy; <a href="https://carto.com/">CARTO</a>'
  ) %>%
  setView(lat = 34.003, lng = -118.3, zoom = 10) %>% 
  addCircleMarkers(data=filtered, lng=~LON, lat=~LAT)
  })
```

# Correlations and Modeling {background="#6B6B6B"}

## Correlation Heat Map
```{r}
library(lubridate)
# creating dataset for correlations
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))
# since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions.
# covid data
query <- dbSendQuery(con, "
  SELECT 
    CAST(strftime('%Y', ep_date) AS INTEGER) AS year,
    CAST(strftime('%m', ep_date) AS INTEGER) AS month,
   CAST(strftime('%d', ep_date) AS INTEGER) AS date, 
   cases_14day
  FROM covid_cases
")
covid_14day <- dbFetch(query,n=-1)
dbClearResult(query)
# since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
covid_14day$lagged <- covid_14day$date + 7

covid_la_lagged <- covid_14day %>% filter(lagged %in% c(21, 35)) %>% arrange(month, year)

covid_la_monthly <- covid_la_lagged %>% group_by(month, year) %>% summarize(cases = sum(cases_14day)) %>% ungroup() %>% arrange(year, month)

covid_la_monthly$index <- 1:nrow(covid_la_monthly)

# next, unemployment rate, lapd budget, and evictions
query <- dbSendQuery(con, "
  SELECT 
    u.month, 
    u.Year, 
    u.Value AS unemployment_rate, 
    l.police_budget, 
    e.Count AS num_evictions
  FROM 
    unemployment_la AS u
  INNER JOIN 
    lapd_budget AS l ON u.Year = l.year
  INNER JOIN 
    evictions_county_la AS e ON u.Year = e.Year AND u.month = e.Month
")
unemployment_lapd_budget_evictions <- dbFetch(query,n=-1)
dbClearResult(query)

dbDisconnect(con)

crime_la_monthly$date <- ymd(paste0(crime_la_monthly$month_year, "-01"))
crime_la_monthly$month <- month(crime_la_monthly$date)
crime_la_monthly$year <- year(crime_la_monthly$date)

correlations_df <- crime_la_monthly %>% inner_join(unemployment_lapd_budget_evictions, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, num_evictions, unemployment_rate, num_crimes, police_budget)

correlations_df2 <- correlations_df %>% inner_join(covid_la_monthly, by = c("month", "year"))

correlations <- correlations_df2 %>% dplyr::select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)

Cor_mat <- round(cor(correlations), 2)

# get only the upper triangle of variables 
# (because a correlation heat map shows each correlation value twice)
get_upper_tri <- function(cormat){
  # lower.tri -> Returns a matrix of logicals, so set the values that are true to NA
  cormat[lower.tri(cormat)]<- NA # set the correlation matrix lower triangle to NA
  return(cormat) 
}

# reorder them by the strength of the correlation
reorder_cormat <- function(cormat){
  # Use correlation between variables as distance
  dd <- as.dist((1-cormat)/2)
  hc <- hclust(dd) # use hierarchical clustering to reorder
  cormat <-cormat[hc$order, hc$order]
}

# use the functions above to remove the unnecessary parts
Cor_mat<-reorder_cormat(Cor_mat)

Cor_mat_upper<-get_upper_tri(Cor_mat)

# using melt instead of pivot_longer because pivot_longer did not order correctly
Cor_mat_upper<-reshape2::melt(Cor_mat_upper, na.rm=TRUE)

# remove the central values 
# which show the correlation for each variable with itself
Cor_mat_upper<- Cor_mat_upper %>% filter(value != 1.00)

long_Cor <- Cor_mat_upper %>%
    mutate(Var1 = recode(Var1, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget ($ billions)"),
           Var2 = recode(Var2, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget \n ($ billions)"))

ggplot(data = long_Cor, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(direction = -1,limit = c(-1, 1)) +
  theme(axis.title.x =  element_blank(),
        axis.title.y = element_blank()) +
  geom_text(aes(label = value), color = "black", size = 4) + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size=12)) +
  labs(fill = "Correlation")
```

## Modeling Crime Rates Spatially

\begin{align}
\text{HomicideCount}_i = \beta_0 + 
\beta_1 \text{Latitude}_i +
\beta_2 \text{Longitude}_i + \epsilon_i
\end{align}

```{r, warning=FALSE, message=FALSE, fig.align = "center"}
con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))

query <- dbSendQuery(con, "
  SELECT i.S1903_C03_001E, x.year, i.ZCTA, x.LAT, x.LON
  FROM (
    SELECT c.LAT, c.LON, CAST(z.ZCTA5CE20 AS INTEGER) AS ZCTA5CE20, CAST(c.year AS INTEGER) AS year
    FROM crime_la AS c
    INNER JOIN crimes_with_zips AS z 
      ON c.DR_NO = z.DR_NO 
      AND c.LAT = z.LAT 
      AND c.LON = z.LON 
      AND c.X = z.X 
      AND c.year = z.year
    WHERE c.\"Crm.Cd.Desc\" = 'CRIMINAL HOMICIDE'
  ) AS x INNER JOIN (SELECT m.S1903_C03_001E, CAST(m.year AS INTEGER) AS year, CAST(m.ZCTA AS INTEGER) AS ZCTA FROM cali_median_income m) AS i 
    ON x.ZCTA5CE20 = i.ZCTA AND x.year = i.year
")

homicides <- dbFetch(query,n=-1)
dbClearResult(query)
dbDisconnect(con)

homicides_latlon_rounded <- homicides %>%
  mutate(LAT_rounded = round(LAT, 2),
         LON_rounded = round(LON, 2))

homicides_lat_long_yr <- homicides_latlon_rounded %>% 
  group_by(year, LAT_rounded, LON_rounded) %>%
  summarize(HomicideCount = n()) %>% ungroup()

linear_model1 <- lm(HomicideCount ~ LAT_rounded + LON_rounded, data = homicides_lat_long_yr)

d <- coef(summary(linear_model1))[,c(1, 4)]
colnames(d) <- c("Coefficient", "p-value")
rownames(d) <- c("(Intercept", "Latitude", "Longitude")
knitr::kable(d)
```

## Homicides Have Increased in the Pandemic

\begin{equation}
\text{HomicideCount}_i = \beta_0 + \beta_1 \text{year}_i + \epsilon_i
\end{equation}

:::::: columns
:::: {.column width="35%"}
Year is significant at a 0.01 level. Each year, homicide rates increase by 0.05 on average.
::::

::: {.column width="65%"}
```{r}
con <- dbConnect(RSQLite::SQLite(), normalizePath("la_stars.db"))

query <- dbSendQuery(con, "SELECT * FROM crime_la_annually WHERE `Crm.Cd.Desc` == 'CRIMINAL HOMICIDE'")
homicides_annual <- dbFetch(query,n=-1)
dbClearResult(query)
dbDisconnect(con)

ggplot(homicides_annual, aes(x = year, y = variable)) +
  geom_col(fill = "#19297C") +
  ggthemes::theme_fivethirtyeight() +
  theme(axis.title = element_text(), title = element_text(size = 12)) +
  labs(title="Number of Homicides in LA, Annually",
       x = "Year",
       y = "Homicide Count")
```
:::
::::::

# Killer Plot {background="#6B6B6B"}

## Walk of Crime

```{r}
#| context: setup
#| include: false
# load functions
source("dependencies.R")
library(dplyr)

# load data
con <- dbConnect(RSQLite::SQLite(), normalizePath("la_stars.db"))

crime_la_annually <- dbGetQuery(con, "SELECT * FROM crime_la_annually")
covid_la_annually <- dbGetQuery(con, "SELECT * FROM covid_la_annually")
evictions_county_la <- dbGetQuery(con, "SELECT * FROM evictions_county_la")
lapd_budget_annually <- dbGetQuery(con, "SELECT * FROM lapd_budget_annually")
unemployment_la <- dbGetQuery(con, "SELECT * FROM unemployment_la")
correlations_summary <- dbGetQuery(con, "SELECT * FROM correlations_quarterly")
correlations_summary$period <- 1:nrow(correlations_summary)

map_datasets <- list("crime" = crime_la_annually,
                     "covid" = covid_la_annually,
                     "evictions" = evictions_county_la,
                     "budget" = lapd_budget_annually,
                     "unemploy" = unemployment_la)

dbDisconnect(con)
```

```{r}
#| panel: sidebar
  # Dropdown to select dataset
  selectInput("dataset", "Choose Dataset", 
              choices = c("Crime" = "crime", "COVID" = "covid", "Evictions" = "evictions", "LAPD Budget" = 'budget', "Unemployment" = 'unemploy'))

sliderInput("year_range", "Select Time Range:",
              min = 2010, max = 2025, value = c(2015, 2020), sep = "")  # Placeholder, updated dynamically
uiOutput("crime_type_ui")
```

```{r}
#| panel: fill

plotOutput("stars")
```

```{r}
#| context: server

selectedKillerData <- reactive({
  if (input$dataset == 'crime') {
    map_datasets[[input$dataset]] %>% filter(year %in% seq(input$year_range[1], input$year_range[2]),
                                             Crm.Cd.Desc == input$crime_type)
  }

  else {
    map_datasets[[input$dataset]] %>% filter(year %in% seq(input$year_range[1], input$year_range[2]))
  }

  })

observeEvent(input$dataset, {
    # Get year vector based on dataset selected
    year_vals <- switch(input$dataset,
                        "crime" = crime_la_annually$year,
                        "covid" = covid_la_annually$year,
                        "evictions" = evictions_county_la$year,
                        "budget" = lapd_budget_annually$year,
                        "unemploy" = unemployment_la$year)

    # Update slider with dynamic range
    updateSliderInput(session, "year_range",
                      min = min(year_vals),
                      max = max(year_vals),
                      value = c(min(year_vals), max(year_vals)))
  })

output$crime_type_ui <- renderUI({
    if (input$dataset == "crime") {
      selectInput("crime_type", "Select Crime Type",
                  choices = unique(crime_la_annually$Crm.Cd.Desc),
                  selected = unique(crime_la_annually$Crm.Cd.Desc)[1])
    }
  })

output$stars <- renderPlot({
      df <- selectedKillerData()
      print(df)
      rows <- nrow(df)

      col_row <- if (rows <= 13) {
        c(rows, 1)
      } else if (rows == 14) {
        c(7, 2)
      } else if (rows == 15) {
        c(5, 3)
      } else if (rows == 16) {
        c(4, 4)
      } else {
        c(ceiling(sqrt(rows)), ceiling(rows / ceiling(sqrt(rows))))
      }

      stars <- generate_star_centers(col_row[1], col_row[2])
      labels <- levels(standardize_scale(df$variable))
      print(labels)
      df_sorted <- df[order(df$year), ]

      stars$scale <- df_sorted$bins
      stars$year <- df_sorted$year
      print(stars)

      grid.newpage()
      grid.rect(gp = gpar(fill = "gray15", col = NA))  # No border, black fill

      for (star in 1:nrow(stars)) {
        row <- as.list(stars[star,])
        grid.draw(create_star(scale_factor = row$scale, center = c(row$x, row$y)))

        grid.text(label = as.character(row$year),
    x = unit(row$x, "npc"),
    y = unit(row$y - 0.05, "npc"),  # offset below star; adjust as needed
    gp = gpar(col = "white", fontsize = 10))

        grid.rect(x = 0.5, y = 0, width = 1, height = 0.125, just = "bottom", gp = gpar(col = "white", fill = "gray15"))

        grid.text(label = paste(stringr::str_to_title(input$dataset), "Level"),
    x = unit(.1, "npc"),
    y = unit(0.0625, "npc"),  # offset below star; adjust as needed
    gp = gpar(col = "white", fontsize = 14))
      }

      grid.draw(create_star(scale_factor = 1, center = c(.2, .0625)))

      grid.draw(create_star(scale_factor = 2, center = c(.375, .0625)))

      grid.draw(create_star(scale_factor = 3, center = c(.6, .0625)))

      grid.draw(create_star(scale_factor = 4, center = c(.85, .0625)))

      grid.text(label = as.character(labels[2]), x = unit(.275, "npc"),
                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11))
      grid.text(label = as.character(labels[3]), x = unit(.475, "npc"),
                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11))
      grid.text(label = as.character(labels[4]), x = unit(.725, "npc"),
                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11))
      grid.text(label = as.character(labels[5]), x = unit(.925, "npc"),
                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11))
    })
```

### 2025-04-15 Slides/dependencies.R

@@ -1,0 +1,71 @@
+library(grid)
+library(RSQLite)
+library(dplyr)
+library(shiny)
+
+# killer plot pipeline functions
+standardize_scale <- function(column) {
+  breaks <- quantile(column, probs = seq(0, 1, 1/3))
+  cuts <- cut(column, breaks = c(-Inf, breaks, Inf))
+  lbls <- c(
+    paste0("< ", round(breaks[1])),
+    paste0("[", round(breaks[1]), ", ", round(breaks[2]), "]"),
+    paste0("[", round(breaks[2]), ", ", round(breaks[3]), "]"),
+    paste0("[", round(breaks[3]), ", ", round(breaks[4]), "]"),
+    paste0("> ", round(breaks[4]))
+  )
+  return(factor(cuts, levels = levels(cuts), labels = lbls))
+}
+
+generate_star_centers <- function(n_cols, n_rows, radius = 0.05 / 2, padding = c(0.05, .1)) {
+  # X spacing
+  usable_width <- 1 - 2 * padding[1]
+  step_x <- (usable_width - 2 * radius) / (n_cols - 1)
+  spaceX <- step_x - 2 * radius
+  x_start <- padding[1]
+  x <- seq(x_start, by = (2 * radius + spaceX), length.out = n_cols)
+  
+  # Y spacing with top and bottom padding
+  usable_height <- 1 - 2 * padding[2] - .1
+  
+  # Handle the case where n_rows is 1 to avoid NaN (division by zero)
+  if (n_rows > 1) {
+    step_y <- (usable_height - 2 * radius) / (n_rows - 1)
+    spaceY <- step_y - 2 * radius
+  } else {
+    step_y <- usable_height
+    spaceY <- 0
+  }
+  
+  y_start <- ifelse(n_rows > 1, padding[2] + .125, .5)
+  y <- seq(y_start, by = (2 * radius + spaceY), length.out = n_rows)
+  
+  # Create grid of star centers
+  stars <- expand.grid(x = x, y = y)
+  
+  # Track rows and columns to resort by row
+  stars$row <- rep(1:n_rows, times = n_cols)
+  stars$col <- rep(1:n_cols, each = n_rows)
+  stars <- stars[order(stars$col, stars$row), ]
+  
+  return(stars)
+}
+create_star <- function(scale_factor = 1, center = c(0.06, 0.06), fill = '#a44c5c', color = "#d3af37") {
[…]

### 2025-04-16 Draft4_2.qmd
final slides version and PDF

--- 2025-04-10 Draft4.qmd
@@
 In particular, we will look at the LAPD budget in comparison to the homicide rate to determine if increased police funding is effective in reducing the number of homicides.
-
-```{r DATA, output=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo = FALSE}
+knitr::opts_chunk$set(eval = TRUE, echo = FALSE)
+```
+
+```{r DATA, output=FALSE,  message=FALSE, warning=FALSE}
 library(sf) # for map
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Crime densities in Los Angeles by LAPD Division, 2010-2019 and 2020-Present"}
+```{r  message=FALSE, warning=FALSE, fig.align="center", fig.cap="Crime densities in Los Angeles by LAPD Division, 2010-2019 and 2020-Present"}
 query <- dbSendQuery(con, "Select * From crime_la Where year < 2020")
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap = "Weekly Crime Counts in Los Angeles from 2010 to present", fig.align = "center"}
-query <- dbSendQuery(con, "Select Count(*) as Crimes, YearWeek From crime_la Group By YearWeek Order By YearWeek")
-weekly_counts <- dbFetch(query,n=-1)
-dbClearResult(query)
+```{r,  warning=FALSE, message=FALSE, fig.cap = "Weekly Crime Counts in Los Angeles from 2010 to present", fig.align = "center"}
+
+query <- dbSendQuery(
+  con,
+  "
+  SELECT
+    YearWeek,
+    MIN(Date) AS start_of_week,
+    MAX(Date) AS end_of_week,
+    COUNT(*)  AS Crimes
+  FROM crime_la
+  GROUP BY YearWeek
+  ORDER BY YearWeek
+  "
+)
+weekly_counts <- dbFetch(query, n = -1)
+dbClearResult(query)
+
+weekly_counts$start_of_week <- as.Date(weekly_counts$start_of_week)
+weekly_counts$end_of_week   <- as.Date(weekly_counts$end_of_week)
+
+weekly_counts$days_in_week <-
+  as.numeric(difftime(weekly_counts$end_of_week,
+                      weekly_counts$start_of_week,
+                      units = "days")) + 1
+
+weekly_counts <- subset(weekly_counts, days_in_week == 7)
+
+weekly_counts$CrimesScaled <-
+  weekly_counts$Crimes * (7 / weekly_counts$days_in_week)
 
 weekly_counts$Index <- seq_len(nrow(weekly_counts))
-# Find the Index for March 2020
-march_2020_index <- which(weekly_counts$YearWeek == "2020-09")  # Adjust '202009' to match your data's YearWeek format for March 2020
-
-# Plot
+
+march_2020_index <- which(weekly_counts$YearWeek == "2020-09")
+
 plot(
   weekly_counts$Index,
-  weekly_counts$Crimes,
+  weekly_counts$CrimesScaled,
   type  = "l",
@@
   xlab  = "Weekly Index",
-  ylab = "Number of Crimes")
-
-abline(h = pretty(weekly_counts$Crimes), v = pretty(weekly_counts$Index), 
-       col = "gray", lty = "dotted")
-
-# Add vertical line for March 2020
+  ylab  = "Number of Crimes (Scaled to 7 Days)"
+)
+
+abline(
+  h   = pretty(weekly_counts$CrimesScaled),
+  v   = pretty(weekly_counts$Index),
+  col = "gray",
+  lty = "dotted"
+)
+
 abline(v = weekly_counts$Index[march_2020_index], col = "red", lwd = 2)
 
-# Add a label for March 2020
 text(
@@
 )
+
 ```
@@
 
-```{r, echo=FALSE, fig.cap = "Los Angeles Crimes by Day of the Week", warning=FALSE, message=FALSE, fig.align = "center"}
+```{r,  fig.cap = "Los Angeles Crimes by Day of the Week", warning=FALSE, message=FALSE, fig.align = "center"}
 # Convert Date.Rptd to Date format (ignore time part)
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Percent of crime victims by gender, 2010 to present", fig.align = "center"}
+```{r,  warning=FALSE, message=FALSE, fig.cap="Percent of crime victims by gender, 2010 to present", fig.align = "center"}
 query <- dbSendQuery(con, "SELECT COUNT(*) As count, Modified_Vict_Sex As `Vict.Sex` FROM (
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+```{r,  warning=FALSE, message=FALSE, fig.align = "center"}
 zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Median household income by year and by ZIP Code Tabulation Area (ZCTA) in Los Angeles", fig.align = "center"}
+```{r,  warning=FALSE, message=FALSE, fig.cap="Median household income by year and by ZIP Code Tabulation Area (ZCTA) in Los Angeles", fig.align = "center"}
 # quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Annual Median Income and Homicide Count by ZCTA in Los Angeles with top 1% income and top 1% homicide rates labelled", fig.align = "center"}
+```{r,  warning=FALSE, message=FALSE, fig.cap="Annual Median Income and Homicide Count by ZCTA in Los Angeles with top 1% income and top 1% homicide rates labelled", fig.align = "center"}
 zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025"}
+```{r,  warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025"}
 
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Annual Number of Crimes by Los Angeles Police Department Bureau"}
+```{r,  warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Annual Number of Crimes by Los Angeles Police Department Bureau"}
 crime_la$year <- lubridate::year(crime_la$DateParsed)
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Strength of Correlations Between the Number of Crimes and Different Law Enforcement and Socioeconomic Factors"}
+```{r,  warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Strength of Correlations Between the Number of Crimes and Different Law Enforcement and Socioeconomic Factors"}
 library(lubridate)
@@
 
-correlations_df <- crime_la_monthly %>% inner_join(unemployment_lapd_budget_evictions, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, num_evictions, unemployment_rate, num_crimes, police_budget)
+correlations_df <- crime_la_monthly %>% inner_join(unemployment_lapd_budget_evictions, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, num_evictions, unemployment_rate, num_crimes, police_budget, Crm.Cd.Desc)
 
@@
 correlations <- correlations_df2 %>% dplyr::select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)
+
+write.csv(correlations, "corr.csv")
 
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Homicide Victims by Age and Gender in Los Angeles 2010-Present"}
+```{r,  warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Homicide Victims by Age and Gender in Los Angeles 2010-Present"}
 homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% dplyr::select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex, levels  = c("F", "M", "X"), labels = c("Female", "Male", "Other"))) %>% drop_na()
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Frequency of Crimes in Los Angeles Throughout the Day"}
+```{r,  warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Frequency of Crimes in Los Angeles Throughout the Day"}
 crime_la$time <- sprintf("%04d",crime_la$TIME.OCC)
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Time series analyzing homicide rate trends over time by median income quartiles"}
+```{r,  warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Time series analyzing homicide rate trends over time by median income quartiles"}
 query <- dbSendQuery(con, "
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+```{r,  warning=FALSE, message=FALSE, fig.align = "center"}
 library(broom)
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+```{r,  warning=FALSE, message=FALSE, fig.align = "center"}
 lm_income_crimes <- lm(HomicideCount ~ cali_median_income, data = homicide_income_by_year)
@@
 
-```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.align = "center"}
+```{r, warning=FALSE, message=FALSE,  fig.align = "center"}
 library(kableExtra)
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+```{r,  warning=FALSE, message=FALSE, fig.align = "center"}
 homicides_latlon_rounded <- homicides %>%
@@
 
+write.csv(homicides_lat_long_yr, "homicides.csv")
+
 linear_model1 <- lm(HomicideCount ~ LAT_rounded + LON_rounded, data = homicides_lat_long_yr)
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align = "center"}
+```{r,  warning=FALSE, message=FALSE, fig.align = "center"}
 
@@
 
-```{r}
-knitr::opts_chunk$set(eval = FALSE)
-```
-
-```{r, output=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo = FALSE}
+knitr::opts_chunk$set(eval = FALSE, echo = TRUE)
+```
+
+```{r, output=FALSE,  message=FALSE, warning=FALSE}
 library(sf) # for map
@@
 
-# Exploratory Data Analysis
-
-```{r echo=FALSE, fig.cap="Crime densities in Los Angeles by LAPD Division, message=FALSE, warning=FALSE, 2010.2019 and 2020.Present", fig.align="center"}
+```{r  fig.cap="Crime densities in Los Angeles by LAPD Division, message=FALSE, warning=FALSE, 2010.2019 and 2020.Present", fig.align="center"}
 query <- dbSendQuery(con, "Select * From crime_la Where year < 2020")
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap = "Weekly Crime Counts in Los Angeles from 2010 to present", fig.align = "center"}
+```{r,  warning=FALSE, message=FALSE, fig.cap = "Weekly Crime Counts in Los Angeles from 2010 to present", fig.align = "center"}
 query <- dbSendQuery(con, "Select Count(*) as Crimes, YearWeek From crime_la Group By YearWeek Order By YearWeek")

### 2025-04-16 Slides/Slides_2.qmd
final slides version and PDF

--- 2025-04-15 Slides/Slides.qmd
@@
 
+## Weekly Crime Counts in Los Angeles from 2010 to present
+
+```{r}
+con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))
+
+query <- dbSendQuery(
+  con,
+  "
+  SELECT
+    YearWeek,
+    MIN(Date) AS start_of_week,
+    MAX(Date) AS end_of_week,
+    COUNT(*)  AS Crimes
+  FROM crime_la
+  GROUP BY YearWeek
+  ORDER BY YearWeek
+  "
+)
+weekly_counts <- dbFetch(query, n = -1)
+dbClearResult(query)
+dbDisconnect(con)
+
+weekly_counts$start_of_week <- as.Date(weekly_counts$start_of_week)
+weekly_counts$end_of_week   <- as.Date(weekly_counts$end_of_week)
+
+weekly_counts$days_in_week <-
+  as.numeric(difftime(weekly_counts$end_of_week,
+                      weekly_counts$start_of_week,
+                      units = "days")) + 1
+
+weekly_counts <- subset(weekly_counts, days_in_week == 7)
+
+weekly_counts$CrimesScaled <-
+  weekly_counts$Crimes * (7 / weekly_counts$days_in_week)
+
+weekly_counts$Index <- seq_len(nrow(weekly_counts))
+
+march_2020_index <- which(weekly_counts$YearWeek == "2020-09")
+
+plot(
+  weekly_counts$Index,
+  weekly_counts$CrimesScaled,
+  type  = "l",
+  col   = "blue",
+  lwd   = 2,
+  xlab  = "Weekly Index",
+  ylab  = "Number of Crimes (Scaled to 7 Days)"
+)
+
+abline(
+  h   = pretty(weekly_counts$CrimesScaled),
+  v   = pretty(weekly_counts$Index),
+  col = "gray",
+  lty = "dotted"
+)
+
+abline(v = weekly_counts$Index[march_2020_index], col = "red", lwd = 2)
+
+text(
+  x     = weekly_counts$Index[march_2020_index] - 80,
+  y     = 10,
+  labels = "March 2020",
+  pos   = 3,
+  col   = "red"
+)
+
+```
+
 ## Homicides Over Time by Income Quartile
 
-```{r}
+```{r, results='asis'}
+library(plotly)
 con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))
@@
 levels(homicide_trends$income_quartile) <- c("Low Income", "Lower-Middle Income", "Higher-Middle Income", "High Income")
+
+homicide_trends <- homicide_trends %>%
+  mutate(income_quartile = as.factor(income_quartile)) %>%
+  rename("Year" = "year",
+         "Homicide Count" = "homicide_count",
+         "Income Quartile" = "income_quartile"
+         )
 # Plot homicide trends over time by income quartile
-ggplot(homicide_trends, aes(x = year, y = homicide_count)) +
-  geom_area(aes(fill = as.factor(income_quartile)), alpha = 0.6, position = "identity") +
+p <- ggplot(homicide_trends, aes(x = Year, y = `Homicide Count`)) +
+  geom_area(aes(fill = `Income Quartile`), alpha = 0.6, position = "identity") +
   labs(
@@
   scale_fill_viridis_d(option = "plasma", direction = -1)
+
+ggplotly(p)
 ```
@@
 ## Homicides and Median Income By Zip Code
+
 ```{r, echo=FALSE}
@@
 
+## Relationship Between Homicides and Socioeconomic Variables
+```{r}
+#| context: setup
+#| include: false
+library(RSQLite)
+library(ggplot2)
+library(dplyr)
+
+corr <- read.csv("corr.csv")
+
+colnames(corr) <- c("X", "Year", "Month", "Unemployment", "Crimes", "COVID", "Evictions", "LAPD Budget")
+```
+
+```{r}
+#| panel: sidebar
+radioButtons("opts", "View homicides compared to:", 
+            choices = c("COVID", "Evictions", "LAPD Budget", "Unemployment"))
+```
+
+```{r}
+#| panel: fill
+plotOutput("scatter")
+```
+
+```{r}
+#| context: server
+output$scatter <- renderPlot({
+  # Guessing 'value' is the comparison variable — adjust as needed
+  df <- corr[c("Crimes", input$opts, 'Year')]
+  colnames(df) <- c("Crimes", "var", "Year")
+
+  ggplot(df, aes(x = var, y = Crimes)) +
+    geom_point(aes(colour = as.factor(Year))) +
+    scale_color_viridis_d() +
+  ggthemes::theme_fivethirtyeight() +
+  theme(axis.title = element_text()) + 
+    labs(x = input$opts, y = "Crime Count", title = paste("Monthly Crime vs", input$opts),
+         colour = "Year") +
+    geom_smooth(method = "lm")
+}) 
+```
+
 ## Modeling Crime Rates Spatially
@@
 
-:::::: columns
-:::: {.column width="35%"}
-Year is significant at a 0.01 level. Each year, homicide rates increase by 0.05 on average.
-::::
-
-::: {.column width="65%"}
-```{r}
-con <- dbConnect(RSQLite::SQLite(), normalizePath("la_stars.db"))
-
-query <- dbSendQuery(con, "SELECT * FROM crime_la_annually WHERE `Crm.Cd.Desc` == 'CRIMINAL HOMICIDE'")
-homicides_annual <- dbFetch(query,n=-1)
-dbClearResult(query)
-dbDisconnect(con)
-
-ggplot(homicides_annual, aes(x = year, y = variable)) +
-  geom_col(fill = "#19297C") +
-  ggthemes::theme_fivethirtyeight() +
-  theme(axis.title = element_text(), title = element_text(size = 12)) +
-  labs(title="Number of Homicides in LA, Annually",
-       x = "Year",
-       y = "Homicide Count")
-```
-:::
-::::::
+Year is significant at a 0.01 level, but it has caused a very small increase, almost unnoticeable on the graph. Each year, homicide rates increase by 0.05 on average.
 
@@
 library(dplyr)
-
+spatial <- read.csv("homicides.csv")
 # load data
@@
 
-map_datasets <- list("crime" = crime_la_annually,
-                     "covid" = covid_la_annually,
-                     "evictions" = evictions_county_la,
-                     "budget" = lapd_budget_annually,
-                     "unemploy" = unemployment_la)
+map_datasets <- list("Crime Count" = crime_la_annually,
+                     "Covid Cases" = covid_la_annually,
+                     "Evictions" = evictions_county_la,
+                     "LAPD Budget" = lapd_budget_annually,
+                     "Unemployment Rate" = unemployment_la)
 
@@
 #| panel: sidebar
+
+radioButtons("type", "View:", choices = c("Homicides Spatially", "A Variable Over Time"), selected = "Homicides Spatially")
 # Dropdown to select dataset
-  selectInput("dataset", "Choose Dataset", 
-              choices = c("Crime" = "crime", "COVID" = "covid", "Evictions" = "evictions", "LAPD Budget" = 'budget', "Unemployment" = 'unemploy'))
+uiOutput("dataset")
 
-sliderInput("year_range", "Select Time Range:",
-              min = 2010, max = 2025, value = c(2015, 2020), sep = "")  # Placeholder, updated dynamically
+uiOutput("year_selector")
+
 uiOutput("crime_type_ui")
@@
 selectedKillerData <- reactive({
-  if (input$dataset == 'crime') {
-    map_datasets[[input$dataset]] %>% filter(year %in% seq(input$year_range[1], input$year_range[2]),
+  req(input$dataset, input$year_range)
+
+  if (input$dataset == 'Crime Count') {
+    req(input$crime_type)
+    df <- map_datasets[[input$dataset]] %>%
+      filter(year %in% seq(input$year_range[1], input$year_range[2]),
              Crm.Cd.Desc == input$crime_type)
+  } else {
+    df <- map_datasets[[input$dataset]] %>%
+      filter(year %in% seq(input$year_range[1], input$year_range[2]))
   }
-
-  else {
-    map_datasets[[input$dataset]] %>% filter(year %in% seq(input$year_range[1], input$year_range[2]))
-  }
-
+  return(df)
 })
 
-observeEvent(input$dataset, {
-    # Get year vector based on dataset selected
+output$dataset <- renderUI({
+  req(input$type)
+
+  if (input$type == "A Variable Over Time") {
+  selectInput("dataset", "Choose Dataset", 
+            choices = c("Crime Count", 
+                        "Covid Cases", 
+                        "Evictions", 
+                        "LAPD Budget", 
+                        "Unemployment Rate")) 
+}
+})
+
+output$year_selector <- renderUI({
+  req(input$type)
+
+  if (input$type == "Homicides Spatially") {
+    # Show this when Homicides Spatially is selected
+    selectInput("year_single", "Select Year:", choices = 2019:2023, selected = 2023)
+  } else {
+    req(input$dataset)  # Now only required when viewing variables over time
+
+    # Get relevant years
     year_vals <- switch(input$dataset,
-                        "crime" = crime_la_annually$year,
-                        "covid" = covid_la_annually$year,
-                        "evictions" = evictions_county_la$year,
-                        "budget" = lapd_budget_annually$year,
-                        "unemploy" = unemployment_la$year)
+                        "Crime Count" = crime_la_annually$year,
+                        "Covid Cases" = covid_la_annually$year,
+                        "Evictions" = evictions_county_la$year,
+                        "LAPD Budget" = lapd_budget_annually$year,
+                        "Unemployment Rate" = unemployment_la$year)
 
-    # Update slider with dynamic range
-    updateSliderInput(session, "year_range",
-                      min = min(year_vals),
-                      max = max(year_vals),
-                      value = c(min(year_vals), max(year_vals)))
+    year_vals <- sort(unique(year_vals))
+
+    sliderInput("year_range", "Select Time Range:",
+                min = min(year_vals), max = max(year_vals),
+                value = c(min(year_vals), max(year_vals)), step = 1,
+                sep = "")
+  }
 })
@@
 output$crime_type_ui <- renderUI({
-    if (input$dataset == "crime") {
+  req(input$type)
+  req(input$dataset)
+
+  if (input$type == "A Variable Over Time" && input$dataset == "Crime Count") {
     selectInput("crime_type", "Select Crime Type",
@@
 output$stars <- renderPlot({
+  req(input$type)
+
+  if (input$type == "Homicides Spatially") {
+    req(input$year_single)
+    print(as.integer(input$year_single))
+    print(unique(spatial$year))
+    df <- spatial %>% filter(year == as.integer(input$year_single))
+    df$LAT <- round(df$LAT_rounded, 1)
+    df$LON <- round(df$LON_rounded, 1)
+
+    df <- df %>% group_by(LAT, LON) %>% summarize(avgHomicides = mean(HomicideCount)) %>% ungroup()
+
+    df$LATstd <- .175 + ((df$LAT - min(df$LAT))/(max(df$LAT) - min(df$LAT))) * .75
+    df$LONstd <- .175 + ((df$LON - min(df$LON))/(max(df$LON) - min(df$LON))) * .75
+   labels <- levels(standardize_scale(df$avgHomicides))
+   df$scale <- as.numeric(standardize_scale(df$avgHomicides))
+
+   grid.newpage()
+   grid.rect(gp = gpar(fill = "gray15", col = NA))  # No border, black fill
+   for (star in 1:nrow(df)) {
+        row <- as.list(df[star,])
+        grid.draw(create_star(scale_factor = row$scale, center = c(row$LONstd, row$LATstd))) 
+        grid.text(label = paste0(row$LAT, ", ", row$LON),
+    x = unit(row$LONstd, "npc"),
+    y = unit(row$LATstd + 0.05, "npc"),  # offset below star; adjust as needed
+    gp = gpar(col = "white", fontsize = 10))
+        }
+   grid.rect(x = 0.5, y = 0, width = 1, height = 0.125, just = "bottom", gp = gpar(col = "white", fill = "gray15"))
+
+        grid.text(label = "Average Homicides",
+    x = unit(.125, "npc"),
+    y = unit(0.0625, "npc"),  # offset below star; adjust as needed
+    gp = gpar(col = "white", fontsize = 14))
+
+      grid.draw(create_star(scale_factor = 1, center = c(.25, .0625)))
+
+      grid.draw(create_star(scale_factor = 2, center = c(.375, .0625)))
+
+      grid.draw(create_star(scale_factor = 3, center = c(.6, .0625)))
+
+      grid.draw(create_star(scale_factor = 4, center = c(.85, .0625)))
+
+      grid.text(label = as.character(labels[1]), x = unit(.275, "npc"),
+                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11))
+      grid.text(label = as.character(labels[2]), x = unit(.475, "npc"),
+                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11))
+      grid.text(label = as.character(labels[3]), x = unit(.725, "npc"),
+                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11))
+      grid.text(label = as.character(labels[4]), x = unit(.925, "npc"),
+                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11)) 
+} else {
       df <- selectedKillerData()
-      print(df)
+      print(head(df))
       rows <- nrow(df)
-
+      print(rows)
       col_row <- if (rows <= 13) {
@@
       stars <- generate_star_centers(col_row[1], col_row[2])
+      print(nrow(stars))
       labels <- levels(standardize_scale(df$variable))
@@
     gp = gpar(col = "white", fontsize = 10))
-
+      }
         grid.rect(x = 0.5, y = 0, width = 1, height = 0.125, just = "bottom", gp = gpar(col = "white", fill = "gray15"))
@@
     gp = gpar(col = "white", fontsize = 14))
-      }
 
-      grid.draw(create_star(scale_factor = 1, center = c(.2, .0625)))
+      grid.draw(create_star(scale_factor = 1, center = c(.225, .0625)))
 
@@
       grid.text(label = as.character(labels[5]), x = unit(.925, "npc"),
-                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11))
+                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11)) }
     })

### 2025-04-16 Slides/SlidesFinal.qmd
final slides version

--- 2025-04-16 Slides/Slides_2.qmd
@@
     css: styles.css
-    footer: Group 5 - STAT 405/605 - Bianca Schutz, Kevin Hu, Kaitlyn Kirt, Marco Stine - Department of Statistics, Rice University - Source [quarto.org/revealjs](https://quarto.org/docs/presentations/revealjs/)
+    footer: Group 5 - STAT 405 - Bianca Schutz, Kevin Hu, Kaitlyn Kirt, Marco Stine - Crime in Los Angeles
 server: shiny
@@
 ```{r}
+#| fig.align: 'center'
 con <- dbConnect(RSQLite::SQLite(), normalizePath("crime_in_la.db"))
@@
 ```{r, results='asis'}
+#| fig.align: 'center'
 library(plotly)
@@
 ```{r, echo=FALSE}
+#| fig.align: 'center'
 library(RSQLite)
@@
 ```{r}
+#| fig.align: 'center'
 library(lubridate)
@@
 
+Homicide rates are higher in the southeastern parts of Los Angeles.
+
 ```{r, warning=FALSE, message=FALSE, fig.align = "center"}
@@
 
-## Homicides Have Increased in the Pandemic
+## Homicides Increased In Pandemic Years
 
+\small
 \begin{equation}
@@
 
-Year is significant at a 0.01 level, but it has caused a very small increase, almost unnoticeable on the graph. Each year, homicide rates increase by 0.05 on average.
+```{r}
+#| fig.align: 'center'
+#| fig.width: 6
+#| fig.height: 4
+avg <- homicides_lat_long_yr %>% group_by(year) %>% summarize(avgHC = mean(HomicideCount)) %>% ungroup()
+ggplot(avg, aes(x = year, y = avgHC)) +
+  geom_line(color = "#7e03a8") +
+  ggthemes::theme_fivethirtyeight() +
+  theme(axis.title = element_text(),
+        plot.caption = element_text(size = 8)) +
+  labs(y = "Average Homicide Count",
+       x = "Year",
+       title = "Average Homicide Count By Year",
+       caption = "Average was derived from latitude and longitude groups used in the model.")
+```
 

### 2025-05-01 1_DraftVersions/4_FourthDraft/Draft4.qmd
Reorganized repo

--- 2025-04-16 Draft4_2.qmd
@@
 ```{r,  warning=FALSE, message=FALSE, fig.cap = "Weekly Crime Counts in Los Angeles from 2010 to present", fig.align = "center"}
-
-query <- dbSendQuery(
-  con,
-  "
-  SELECT
-    YearWeek,
-    MIN(Date) AS start_of_week,
-    MAX(Date) AS end_of_week,
-    COUNT(*)  AS Crimes
-  FROM crime_la
-  GROUP BY YearWeek
-  ORDER BY YearWeek
-  "
-)
-weekly_counts <- dbFetch(query, n = -1)
-dbClearResult(query)
-
-weekly_counts$start_of_week <- as.Date(weekly_counts$start_of_week)
-weekly_counts$end_of_week   <- as.Date(weekly_counts$end_of_week)
-
-weekly_counts$days_in_week <-
-  as.numeric(difftime(weekly_counts$end_of_week,
-                      weekly_counts$start_of_week,
-                      units = "days")) + 1
-
-weekly_counts <- subset(weekly_counts, days_in_week == 7)
-
-weekly_counts$CrimesScaled <-
-  weekly_counts$Crimes * (7 / weekly_counts$days_in_week)
+query <- dbSendQuery(con, "Select Count(*) as Crimes, YearWeek From crime_la Group By YearWeek Order By YearWeek")
+weekly_counts <- dbFetch(query,n=-1)
+dbClearResult(query)
 
 weekly_counts$Index <- seq_len(nrow(weekly_counts))
-
-march_2020_index <- which(weekly_counts$YearWeek == "2020-09")
-
+# Find the Index for March 2020
+march_2020_index <- which(weekly_counts$YearWeek == "2020-09")  # Adjust '202009' to match your data's YearWeek format for March 2020
+
+# Plot
 plot(
   weekly_counts$Index,
-  weekly_counts$CrimesScaled,
+  weekly_counts$Crimes,
   type = "l",
@@
   xlab = "Weekly Index",
-  ylab  = "Number of Crimes (Scaled to 7 Days)"
-)
-
-abline(
-  h   = pretty(weekly_counts$CrimesScaled),
-  v   = pretty(weekly_counts$Index),
-  col = "gray",
-  lty = "dotted"
-)
-
+  ylab = "Number of Crimes")
+
+abline(h = pretty(weekly_counts$Crimes), v = pretty(weekly_counts$Index), 
+       col = "gray", lty = "dotted")
+
+# Add vertical line for March 2020
 abline(v = weekly_counts$Index[march_2020_index], col = "red", lwd = 2)
 
+# Add a label for March 2020
 text(
@@
 )
-
 ```
@@
 
-correlations_df <- crime_la_monthly %>% inner_join(unemployment_lapd_budget_evictions, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, num_evictions, unemployment_rate, num_crimes, police_budget, Crm.Cd.Desc)
+correlations_df <- crime_la_monthly %>% inner_join(unemployment_lapd_budget_evictions, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, num_evictions, unemployment_rate, num_crimes, police_budget)
 
@@
 correlations <- correlations_df2 %>% dplyr::select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)
-
-write.csv(correlations, "corr.csv")
 
@@
 
-write.csv(homicides_lat_long_yr, "homicides.csv")
-
 linear_model1 <- lm(HomicideCount ~ LAT_rounded + LON_rounded, data = homicides_lat_long_yr)

### 2025-05-01 1_DraftVersions/4_FourthDraft/Draft4_2.qmd
Reorganized repo

--- 2025-04-16 Draft4_2.qmd
@@
 During the COVID-19 pandemic in 2021, homicide rates in Los Angeles increased substantially.
-A report by the Legal Defense Fund's Thurgood Marshall Institute found that this increase was associated with both pre-pandemic and pandemic-induced economic instability and inequalities.[@moore2022]
+A report by the Legal Defense Fund's Thurgood Marshall Institute found that this increase was associated with both pre-pandemic and pandemic-induced economic instability and inequalities[@moore2022].
 
@@
 To group this data geographically, we will organize the crimes by their latitude and longitude into the appropriate Census Blocks and ZIP Code Tabulation Areas (ZCTAs).
-Additionally, we will examine the impact of latitude and longitude themselves, as humans do not always behave by administrative unit boundaries and latitude and longitude by themselves might suggest a different pattern.[@krieger2002]
+Additionally, we will examine the impact of latitude and longitude themselves, as humans do not always behave by administrative unit boundaries and latitude and longitude by themselves might suggest a different pattern[@krieger2002].
 
@@
 
-To maximize the amount of data available for analysis, we used LA County socioeconomic data from 2019-
+To maximize the amount of data available for analysis, we used LA County socioeconomic data for any correlation analysis and focused on years 2020-2023.
 
@@
 In particular, we will look at the LAPD budget in comparison to the homicide rate to determine if increased police funding is effective in reducing the number of homicides.
+
 ```{r, echo = FALSE}
@@
 Figure 2 analyzes the number of crimes over time using time series.
-The time series shows stable weekly crime counts with occasional large spikes, likely due to shorter weeks.
-These surges indicate certain weeks where crime reporting jumped significantly, there is a recent interesting surge that seems to mean-revert in 2024.
+The time series shows stable weekly crime counts with occasional large spikes.
+One large increase occurs around Weekly Index 300, which is around 2016-17.
+This spike was attributed to a resurgence in gang violence and the growth in homelessness.
 The red line, which denotes March 2020, shows a decline in crimes, likely during the shutdown and quarantine, but it returned to normal levels quickly.
+Since then, crimes have decreased recently, but this could be caused by some issues with 2024 data, as the LAPD said they changed their reporting standards: "Records Management System for reporting crimes and arrests" and that they were experiencing issues in 2024.
 
@@
 
-correlations_df <- crime_la_monthly %>% inner_join(unemployment_lapd_budget_evictions, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, num_evictions, unemployment_rate, num_crimes, police_budget, Crm.Cd.Desc)
+correlations_df <- crime_la_monthly %>% inner_join(unemployment_lapd_budget_evictions, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, num_evictions, unemployment_rate, num_crimes, police_budget, num_crimes)
 
@@
 Interestingly, as the unemployment rate increases, the number of crimes decreases, which is the opposite effect of what was expected (one would anticipate that higher unemployment would result in more crime).
-The LAPD budget interestingly had a medium-strength positive correlation with number of crimes, suggesting we should further examine the increase in the LAPD budget compared to the increase in crimes.
+Studies have shown that crime often increases during periods of low unemployment and that employed people commit many crimes[@tarling1982].
+The LAPD budget interestingly had a weak-strength positive correlation with number of crimes.
 
@@
 
-# Statistical Analyses
-
-To examine the relationship between the number of homicides and various socioeconomic and enforcement factors, we will use a variety of statistical analyses to see if certain factors are more influential in causing increases or decreases in homicide rates.
+# Modeling
+
+To examine the relationship between the number of homicides and various socioeconomic and enforcement factors, we will use a variety of models to see if certain factors are more influential in causing increases or decreases in homicide rates.
 
@@
 
-Model 1 is defined as: 
-\begin{equation}
+Model 1 is defined as: \begin{equation}
 \text{HomicideCount}_i = \beta_0 + \beta_1 \text{LAT\_rounded}_i + \beta_2 \text{LON\_rounded}_i + \epsilon_i
@@
 
-Model 2 is defined as: 
-\begin{equation}
+Model 2 is defined as: \begin{equation}
 \text{HomicideCount}_i = \beta_0 + \beta_1 \text{year}_i + \epsilon_i
@@
 
-Model 3 is defined as:
-\begin{equation}
+Model 3 is defined as: \begin{equation}
 \text{HomicideCount}_i = \beta_0 + \beta_1 \text{LAT\_rounded}_i + \beta_2 \text{LON\_rounded}_i + \beta_3 \text{year}_i + \epsilon_i
@@
 
-Model 4 is defined as:
-\begin{align}
+Model 4 is defined as: \begin{align}
 \text{HomicideCount}_i &= \beta_0 + \beta_1 \text{LAT\_rounded}_i + \beta_2 \text{LON\_rounded}_i + \beta_3 \text{year}_i \nonumber \\
@@
 
+# Killer Plot Visualization
+
+The killer plot visualizes crime spatially and over time, stylized based on the Hollywood Walk of Fame since that’s an iconic part of Los Angeles.
+The plot shows the average homicides by latitude and longitude, and you can select a different year.
+You can see that the rates have remained higher in the southeastern areas than the rest of the city, similar to the density map earlier.
+Still, the rates in the northwestern areas are higher in 2023, relative to the rest of the city, compared to previous years where there was a clear difference.
+
+![Understand crimes spatially through the killer plot.](kp1.png)
+
+You can also view different variables over time.
+For example, you can select a range of years to view a particular crime over time, such as homicide.
+You can see that homicide rates have been higher in the last few years.
+You can also look at the decrease in COVID cases since the pandemic, the variability in the number of evictions over time, the increase in the LAPD budget, and the decrease in the unemployment rate since the pandemic.
+
+![Understand crimes and other variables over time in a timeline-style view.](kp2.png)
+
+# Conclusion
+
+Through data visualization and modeling, we have discerned some interesting spatiotemporal patterns in crimes over the last 15 years.
+Figure 2, for example, showed that there have been two spikes in crimes since 2010, one in 2016-17 and one a few weeks after March 2020.
+The increase in 2016-17 is explained by difficulties managing gang activities and a growing homeless population in Los Angeles[@chang2016].
+The second increase can likely be attributed to the pandemic factors described in the Thurgood Marshall Institute report.
+
+Correlations found that evictions were strongly correlated with the number of crimes, suggesting that as the number of evictions increased, crimes also increased.
+This aligns with the finding of the report about homicides, that "social and economic instability and heightened inequalities generated by the COVID-19 pandemic" explains the increase in homicides.
+While evictions in Los Angeles in mid-2020 were lower due to renter protections, by November 2020 the levels once again increased, and reached pre-pandemic levels by 2022[@covid-19].
+However, there was no evidence that COVID-19 case levels had a strong relationship, and unemployment rate had a strong negative relationship with crime rate, suggesting that as unemployment decreases, the number of crimes increases.
+This, while surprising, aligns with past findings in the literature that employed people commit more crimes.
+
+The Thurgood Marshall Institute report also noted that increased police funding was not effective in reducing homicide rates.
+The increase in LAPD funding over the years was a moderately strong positive relationship between time and funding, while crime had a strong positive relationship over time.
+Additionally, a linear model found year to be a significant predictor of homicides, suggesting that over time, the number of homicides has increased.
+This, combined with the weak correlation between crime and LAPD budget, supports this assertion and suggests that increasing police funding might not be an effective strategy.
+
+Finally, we found evidence of spatial trends in crime and homicide rates.
+Table 4 shows that, even when controlling for year, latitude and longitude are significant predictors of homicide rate.
+Namely, homicides are more common in the south and eastern parts of Los Angeles.
+
+In conclusion, we found that crimes have increased through the pandemic years, like the findings in the Thurgood Marshall Institute report.
+We found that in contrast to our initial hypothesis, socioeconomic conditions have not necessarily improved in LA.
+Namely, evictions have returned and surpassed pre-pandemic levels and income inequality appears to be worse than before as more of LA becomes wealthier but certain neighborhoods remain below the poverty line.
+Higher homicide counts occur in areas with lower income, though we've found evidence that this pattern is changing as homicides have decreased in lower income areas and increased in higher income areas in the past few years.
+Ultimately, in order to reduce homicide and crime levels, the City of Los Angeles must do more to combat socioeconomic factors that are associated with increases in crimes, like housing insecurity.
+
 # References {.unnumbered}
@@
 Median Income Data: <https://data.census.gov/table/ACSST1Y2023.S1903?q=income>
-
-# Appendices {.appendix}
-
-```{r, echo = FALSE}
-knitr::opts_chunk$set(eval = FALSE, echo = TRUE)
-```
-
-```{r, output=FALSE,  message=FALSE, warning=FALSE}
-library(sf) # for map
-library(tidyverse) # using only for joining datasets to build map visual
-library(RColorBrewer) # for map color scheme
-library(tidygeocoder) # to get zip codes of addresses
-library(dplyr)
-library(RSQLite)
-
-con <- dbConnect(SQLite(), "0_Data/crime_in_la.db")
-```
-
-```{r  fig.cap="Crime densities in Los Angeles by LAPD Division, message=FALSE, warning=FALSE, 2010.2019 and 2020.Present", fig.align="center"}
-query <- dbSendQuery(con, "Select * From crime_la Where year < 2020")
-crime_la_2010_2019 <- dbFetch(query,n=-1)
-dbClearResult(query)
-
-query <- dbSendQuery(con, "Select * From crime_la Where year > 2019")
-crime_la_2020_present <- dbFetch(query,n=-1)
-dbClearResult(query)
-
-lapd_shape <- read_sf(dsn = "0_Data/Raw_Data/LAPD_Divisions", layer = "LAPD_Divisions")
-
-lapd_shape <- st_transform(lapd_shape, crs = 4326)
-
-cali <- read_sf(dsn = "0_Data/Raw_Data/ca_counties", layer = "ca_counties")
-
-cali <- st_transform(cali, crs = 4326)
-
-par(mfrow = c(1, 2), mar = c(4, 4, 2, 1))
-
-smoothScatter(
-  x    = crime_la_2010_2019$LON,
-  y    = crime_la_2010_2019$LAT,
-  xlim = c(-119, -118),
-  ylim = c(33.5, 34.5),
-  nbin = 300,
-  xlab = "Longitude",
-  ylab = "Latitude",
-  main = "Crime Density (2010–2019)",
-  colramp = colorRampPalette(c("white", "skyblue", "blue", "darkblue")),
-  cex.main=0.8
-)
-
-plot(lapd_shape$geometry, lwd=.5, add=TRUE)
-
-plot(cali$geometry, add = TRUE, lwd=.5)
-
-smoothScatter(
-  x    = crime_la_2020_present$LON,
-  y    = crime_la_2020_present$LAT,
-  xlim = c(-119, -118),
-  ylim = c(33.5, 34.5),
-  nbin = 300,
-  xlab = "Longitude",
-  ylab = "Latitude",
-  main = "Crime Density (2020–Present)",
-  colramp = colorRampPalette(c("white", "pink", "red", "darkred")),
-  cex.main=0.8
-)
-
-plot(lapd_shape$geometry, lwd=.5, add=TRUE)
-
-plot(cali$geometry, add = TRUE, lwd=.5)
-
-par(mfrow = c(1, 1))
-
-```
-
-```{r,  warning=FALSE, message=FALSE, fig.cap = "Weekly Crime Counts in Los Angeles from 2010 to present", fig.align = "center"}
-query <- dbSendQuery(con, "Select Count(*) as Crimes, YearWeek From crime_la Group By YearWeek Order By YearWeek")
-weekly_counts <- dbFetch(query,n=-1)
-dbClearResult(query)
-
-weekly_counts$Index <- seq_len(nrow(weekly_counts))
-# Find the Index for March 2020
-march_2020_index <- which(weekly_counts$YearWeek == "2020-09")  # Adjust '202009' to match your data's YearWeek format for March 2020
-
-# Plot
-plot(
-  weekly_counts$Index,
-  weekly_counts$Crimes,
-  type = "l",
-  col  = "blue",
-  lwd  = 2,
-  xlab = "Weekly Index",
-  ylab = "Number of Crimes")
-
-abline(h = pretty(weekly_counts$Crimes), v = pretty(weekly_counts$Index), 
-       col = "gray", lty = "dotted")
-
-# Add vertical line for March 2020
-abline(v = weekly_counts$Index[march_2020_index], col = "red", lwd = 2)
-
-# Add a label for March 2020
-text(
-  x = weekly_counts$Index[march_2020_index] - 80, 
-  y = 10, 
-  labels = "March 2020", 
-  pos = 3, 
-  col = "red"
-)
-```
-
-```{r,  fig.cap = "Los Angeles Crimes by Day of the Week", warning=FALSE, message=FALSE, fig.align = "center"}
-# Convert Date.Rptd to Date format (ignore time part)
-#crime_la$Date.Rptd
-crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)
-# Extract the day of the week
-crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")
-
-crime_la$DayOfWeek <- weekdays(crime_la$Date)
-# crime_la$DayOfWeek
-
-# Count occurrences of crimes per day
-crime_counts <- table(crime_la$DayOfWeek)
-# crime_counts
-
-# Order the days correctly
-day_order <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
-crime_counts <- data.frame(crime_counts[day_order])
-colnames(crime_counts) <- c("Day", "CrimeCount")
-
-crime_counts$thousands <- crime_counts$CrimeCount/1000
-
-# Plot bar chart with horizontal labels
-ggplot(crime_counts, aes(x = Day, y = thousands)) +
-  geom_col(fill = "blue4") +
-  labs(x = "Day of the Week",
-       y = "Number of Crimes (thousands)")
-```
-
-```{r,  warning=FALSE, message=FALSE, fig.cap="Percent of crime victims by gender, 2010 to present", fig.align = "center"}
-query <- dbSendQuery(con, "SELECT COUNT(*) As count, Modified_Vict_Sex As `Vict.Sex` FROM (
-    SELECT *,  
-        CASE 
-            WHEN `Vict.Sex` NOT IN ('M', 'F') THEN 'N'
-            ELSE `Vict.Sex`
-        END AS `Modified_Vict_Sex`
-    FROM crime_la c
-    WHERE `Vict.Sex` != '' AND `Vict.Sex` IS NOT NULL) s
-    GROUP BY Modified_Vict_Sex
-")
-gender_counts <- dbFetch(query,n=-1)
-dbClearResult(query)
-
-# Create a pie chart using base R
-color <- c("F" = "pink", "M" = "lightblue", "N" = "purple")
-pie(
-  gender_counts$count, 
-  labels = paste(gender_counts$Vict.Sex, 
-                 round(gender_counts$count / sum(gender_counts$count) * 100, 1), "%"), 
-  col = color[gender_counts$Vict.Sex]
-)
-```
-
-```{r,  warning=FALSE, message=FALSE, fig.align = "center"}
-zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
-query <- dbSendQuery(con, "SELECT * FROM zctas_in_la")
-LA_zctas <- dbFetch(query,n=-1)
-dbClearResult(query)
-
-colnames(LA_zctas) <- c("X", "ZCTA5CE20")
-LA_zctas$ZCTA5CE20 <- as.character(LA_zctas$ZCTA5CE20)
-
-# joining the zcta shape file with LA zcta reference file
-zcta_filtered <- zcta %>% inner_join(LA_zctas, by = "ZCTA5CE20")
-
-# getting LA median income using joined file
-query <- dbSendQuery(con, "SELECT ZCTA, S1903_C03_001E, year FROM cali_median_income")
-median_household_income <- dbFetch(query,n=-1)
-dbClearResult(query)
-
-zcta_with_data <- zcta_filtered %>% inner_join(median_household_income, join_by("ZCTA5CE20" == "ZCTA"))
-
-# remove NAs
-zcta_no_NAs <- zcta_with_data %>% filter(!is.na(S1903_C03_001E), S1903_C03_001E != "-")
-
-zcta_no_NAs$S1903_C03_001E <- as.integer(zcta_no_NAs$S1903_C03_001E)
-```
-
-```{r,  warning=FALSE, message=FALSE, fig.cap="Median household income by year and by ZIP Code Tabulation Area (ZCTA) in Los Angeles", fig.align = "center"}
-# quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
-qntls <- round(quantile(zcta_no_NAs$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0)
-
-# 2x3 plot layout for years 2018-2023, setting margins and outer margins, main title sizes
-par(mfrow = c(2, 3), mar = c(1, 1, 1, 5), oma = c(0, 0, 3, 0), cex.main = 1.5)
-
-# color palette for sequential data
-palette <- brewer.pal(7, "Reds")
-
-yrs <- 2018:2023
-
-for (yr in yrs) {
-  # get just that year's data
-  subset_data <- zcta_no_NAs[zcta_no_NAs$year == yr, ]
-
-  # creating bins based on the qtls
-  bins <- cut(subset_data$S1903_C03_001E, breaks = qntls, labels = FALSE, include.lowest = TRUE)
-
-  # Plot
-  plot(st_geometry(subset_data),
-       col = palette[bins],
-       border = 'grey',
-       axes = FALSE,
-       lwd = 0.5)
-  title(yr, line = -1)
-
-  # adding legend
-  legend("right", inset = c(-.75, 0), legend = qntls, fill = palette, title = "Median Income ($)", bty = "n", cex = 0.8, xpd = TRUE)
-}
-
-```
-
-```{r,  warning=FALSE, message=FALSE, fig.cap="Annual Median Income and Homicide Count by ZCTA in Los Angeles with top 1% income and top 1% homicide rates labelled", fig.align = "center"}
-zcta <- read_sf(dsn = "0_Data/Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
-query <- dbSendQuery(con, "SELECT i.*, c.*
-FROM cali_median_income i
-INNER JOIN crimes_with_zips c
-ON i.year = c.year AND CAST(i.ZCTA AS int) = c.ZCTA5CE20")
-crime_income <- dbFetch(query,n=-1)
-dbClearResult(query)
-
-crime_type <- crime_la %>% dplyr::select(year, Crm.Cd.Desc, DR_NO)
-
-crime_income_type <- crime_income[,-c(249, 247)] %>% inner_join(crime_type, join_by("year" == "year", "DR_NO" == "DR_NO"))
-
-homicide_income_by_year <- crime_income_type %>% 
-  filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>%
-  group_by(year, ZCTA5CE20) %>%
-  summarize(HomicideCount = n(),
-            cali_median_income = median(as.numeric(S1903_C03_001E))) %>%
-  ungroup()
-
-homicide_income_by_year <- homicide_income_by_year %>% filter(!is.na(HomicideCount) & !is.na(cali_median_income))
-
-homicide_income_by_year$label <- ifelse(
-  homicide_income_by_year$cali_median_income > quantile(homicide_income_by_year$cali_median_income, 0.99) | 
-  homicide_income_by_year$HomicideCount > quantile(homicide_income_by_year$HomicideCount, 0.99), 
-  homicide_income_by_year$ZCTA5CE20, 
-  "")
-
-ggplot(data = homicide_income_by_year, aes(x = cali_median_income, 
-                                        y = HomicideCount)) +
-  geom_point(aes(color = as.factor(year))) +
-  ggthemes::scale_color_stata() +
-  ggrepel::geom_text_repel(aes(label = label), size = 2.5, nudge_x = 2, nudge_y = 1) + 
-  labs(x = "Median Annual Income ($)",
-       y = "Annual Homicide Count",
-       color = "Year")
-```
-
-```{r,  warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Monthly Frequency of Evictions and Crimes in the City of Los Angeles, February 2023 to February 2025"}
-
-query <- dbSendQuery(con, "SELECT * FROM evictions_city_la")
-evictions_la <- dbFetch(query,n=-1)
-dbClearResult(query)
-# fix mistake
-evictions_la$Date.Received <- gsub("2205", "2025", evictions_la$Date.Received)
-
-# convert to date
-evictions_la$month_year <- format(as.Date(evictions_la$Date.Received, format="%m-%d-%Y"), "%Y-%m")
-
-# get unique months
-unique_months <- data.frame(month_year = sort(unique(evictions_la$month_year)), index = 1:length(unique(evictions_la$month_year)))
-
-evictions_la_indexed <- evictions_la %>% left_join(unique_months, by = "month_year")
-
-# create color scheme
-crime_la$month_year <- format(crime_la$Date, "%Y-%m")
-crime_la_monthly <- crime_la %>% 
-  group_by(month_year) %>% 
-  summarize(num_crimes = n())
-
-evictions_la_indexed_crimes <- evictions_la_indexed %>% left_join(crime_la_monthly, by = "month_year")
-
-avg_crimes <- mean(evictions_la_indexed_crimes$num_crimes)
-
-evictions_la_indexed_crimes$high_low_crime <- as.factor(ifelse(evictions_la_indexed_crimes$num_crimes > avg_crimes, "Above Average", "Below Average"))
-
-ggplot(data = evictions_la_indexed_crimes, aes(x = index)) + 
-  geom_histogram(aes(fill = high_low_crime), bins = nrow(unique_months)) +
-  scale_x_continuous(breaks=unique_months$index,
-        labels=unique_months$month_year) + 
-  labs(x = "Month and Year of Eviction Notice Receipt", y = "Number of Evictions", fill = "Crime Level") +
-  ggthemes::scale_fill_fivethirtyeight() + 
-  ggthemes::theme_stata(scheme = "s1mono") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
-        axis.title.x = element_text(vjust = -.75),
-        plot.title = element_text(size=12))
-```
-
-```{r,  warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Annual Number of Crimes by Los Angeles Police Department Bureau"}
-crime_la$year <- lubridate::year(crime_la$DateParsed)
-crime_by_area <- crime_la %>% group_by(AREA.NAME, year) %>% summarize(numberOfCrimes = n()) %>% ungroup()
-
-query <- dbSendQuery(con, "SELECT * FROM lapd_bureaus")
-bureaus <- dbFetch(query,n=-1)
-dbClearResult(query)
-
-bureaus$Division <- gsub("77th", "77th Street", bureaus$Division)
-bureaus$Division <- gsub("North", "N", bureaus$Division)
-bureaus$Division <- gsub("Neast", "Northeast", bureaus$Division)
-
-query <- dbSendQuery(con, "SELECT COUNT(*) FROM crime_la GROUP BY `AREA.NAME`, year")
-test <- dbFetch(query,)
-crime_by_area <- crime_by_area %>% left_join(bureaus, join_by("AREA.NAME" == "Division"))
-
-ggplot(crime_by_area, aes(y = numberOfCrimes, group = Bureau, x = Bureau)) + 
-        geom_boxplot(aes(fill = Bureau)) +
-  labs(y = "Annual Number of Crimes")
-```
-
-```{r,  warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Strength of Correlations Between the Number of Crimes and Different Law Enforcement and Socioeconomic Factors"}
-library(lubridate)
-# creating dataset for correlations
-
-# since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions. 
-query <- dbSendQuery(con, "SELECT * FROM evictions_county_la")
-LA_county_evictions <- dbFetch(query,n=-1)
-dbClearResult(query)
-
-# covid data
-query <- dbSendQuery(con, "
-  SELECT 
-    CAST(strftime('%Y', ep_date) AS INTEGER) AS year,
-    CAST(strftime('%m', ep_date) AS INTEGER) AS month,
-   CAST(strftime('%d', ep_date) AS INTEGER) AS date, 
-   cases_14day
-  FROM covid_cases
-")
-covid_14day <- dbFetch(query,n=-1)
-dbClearResult(query)
-# since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
-covid_14day$lagged <- covid_14day$date + 7
-
-covid_la_lagged <- covid_14day %>% filter(lagged %in% c(21, 35)) %>% arrange(month, year)
-
-covid_la_monthly <- covid_la_lagged %>% group_by(month, year) %>% summarize(cases = sum(cases_14day)) %>% ungroup() %>% arrange(year, month)
-
-covid_la_monthly$index <- 1:nrow(covid_la_monthly)
-
-# next, unemployment rate
-query <- dbSendQuery(con, "
-  SELECT 
-    u.month, 
-    u.Year, 
-    u.Value, 
-    l.police_budget
-  FROM unemployment_la AS u
-  INNER JOIN lapd_budget AS l ON u.Year = l.year
-")
-unemployment_lapd_budget <- dbFetch(query,n=-1)
-dbClearResult(query)
-
-crime_la_monthly$date <- ymd(paste0(crime_la_monthly$month_year, "-01"))
-crime_la_monthly$month <- month(crime_la_monthly$date)
-crime_la_monthly$year <- year(crime_la_monthly$date)
-
-correlations_df <- crime_la_monthly %>% inner_join(unemployment_lapd_budget, join_by("year" == "Year", "month" == "month")) %>% dplyr::select(month, year, Value, num_crimes, police_budget) %>% rename("unemployment_rate" = "Value")
-
-correlations_df2 <- correlations_df %>% inner_join(covid_la_monthly, by = c("month", "year"))
-
-correlations_df3 <- correlations_df2 %>% inner_join(LA_county_evictions, join_by("year" == "Year", "month" == "Month")) %>% rename("num_evictions" = "Count")
-
-correlations <- correlations_df3 %>% dplyr::select(year, month, unemployment_rate, num_crimes, cases, num_evictions, police_budget)
-
-Cor_mat <- round(cor(correlations), 2)
-
-# get only the upper triangle of variables 
-# (because a correlation heat map shows each correlation value twice)
-get_upper_tri <- function(cormat){
-  # lower.tri -> Returns a matrix of logicals, so set the values that are true to NA
-  cormat[lower.tri(cormat)]<- NA # set the correlation matrix lower triangle to NA
-  return(cormat) 
-}
-
-# reorder them by the strength of the correlation
-reorder_cormat <- function(cormat){
-  # Use correlation between variables as distance
-  dd <- as.dist((1-cormat)/2)
-  hc <- hclust(dd) # use hierarchical clustering to reorder
-  cormat <-cormat[hc$order, hc$order]
-}
-
-# use the functions above to remove the unnecessary parts
-Cor_mat<-reorder_cormat(Cor_mat)
-
-Cor_mat_upper<-get_upper_tri(Cor_mat)
-
-# using melt instead of pivot_longer because pivot_longer did not order correctly
-Cor_mat_upper<-reshape2::melt(Cor_mat_upper, na.rm=TRUE)
-
-# remove the central values 
-# which show the correlation for each variable with itself
-Cor_mat_upper<- Cor_mat_upper %>% filter(value != 1.00)
-
-long_Cor <- Cor_mat_upper %>%
-    mutate(Var1 = recode(Var1, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget ($ billions)"),
-           Var2 = recode(Var2, year = 'Year', month = 'Month', unemployment_rate =  'Unemployment Rate', num_crimes = "Number of Crimes", cases = "COVID Cases", num_evictions = "Number of Evictions", police_budget = "LAPD Budget \n ($ billions)"))
-
-ggplot(data = long_Cor, aes(x = Var1, y = Var2, fill = value)) +
-  geom_tile(color = "white") +
-  scale_fill_viridis_c(direction = -1,limit = c(-1, 1)) +
-  theme(axis.title.x =  element_blank(),
-        axis.title.y = element_blank()) +
-  geom_text(aes(label = value), color = "black", size = 4) + 
-   theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        plot.title = element_text(size=12)) +
-  labs(fill = "Correlation")
-```
-
-```{r,  warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Homicide Victims by Age and Gender in Los Angeles 2010-Present"}
-homicide_by_gender <- crime_la %>% filter(Crm.Cd.Desc == "CRIMINAL HOMICIDE") %>% dplyr::select(Vict.Sex, DateParsed, Vict.Age) %>% mutate(gender = factor(Vict.Sex, levels  = c("F", "M", "X"), labels = c("Female", "Male", "Other"))) %>% drop_na()
-
-levels(homicide_by_gender$gender) <- c("Female", "Male", "Other")
-ggplot(homicide_by_gender, aes(x = gender, y = Vict.Age, color = gender)) +
-  geom_jitter(alpha = 0.4, width = 0.2) +
-  labs(x = "Gender",
-       y = "Age") +
-  theme_minimal() +
-  scale_color_manual(c("M", "F", "N"), values = c("pink2", "blue", "purple")) + 
-  theme(legend.position = 'none')
-```
-
-```{r,  warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Frequency of Crimes in Los Angeles Throughout the Day"}
-crime_la$time <- sprintf("%04d",crime_la$TIME.OCC)
-crime_la$Hour <- as.numeric(substr(as.character(crime_la$time), start = 1, stop = 2))
-
-ggplot(crime_la, aes(x = Hour, y = 0, fill = after_stat(density))) +
-  ggridges::geom_density_ridges_gradient(scale = 1, alpha = 0.5) +
-  scale_fill_gradient(low = "lightblue", high = "darkblue") +
-  labs(x = "Hour (Military Time)",
-       y = "Density") +
-  theme_minimal() +
-  theme(legend.position = 'none')
-```
-
-```{r,  warning=FALSE, message=FALSE, fig.align = "center", fig.cap="Time series analyzing homicide rate trends over time by median income quartiles"}
-query <- dbSendQuery(con, "
-  SELECT i.S1903_C03_001E, x.year, i.ZCTA, x.LAT, x.LON
-  FROM (
-    SELECT c.LAT, c.LON, CAST(z.ZCTA5CE20 AS INTEGER) AS ZCTA5CE20, CAST(c.year AS INTEGER) AS year
-    FROM crime_la AS c
-    INNER JOIN crimes_with_zips AS z 
-      ON c.DR_NO = z.DR_NO 
-      AND c.LAT = z.LAT 
-      AND c.LON = z.LON 
-      AND c.X = z.X 
-      AND c.year = z.year
-    WHERE c.\"Crm.Cd.Desc\" = 'CRIMINAL HOMICIDE'
-  ) AS x INNER JOIN (SELECT m.S1903_C03_001E, CAST(m.year AS INTEGER) AS year, CAST(m.ZCTA AS INTEGER) AS ZCTA FROM cali_median_income m) AS i 
-    ON x.ZCTA5CE20 = i.ZCTA AND x.year = i.year
-")
-
-homicides <- dbFetch(query,n=-1)
-dbClearResult(query)
-
-# Form income quartiles
-homicides <- homicides %>% mutate(income_quartile = ntile(S1903_C03_001E, 4)) 
-homicide_trends <- homicides %>% group_by(year, income_quartile) %>% summarize(homicide_count = n(), 
-          ref = mean(S1903_C03_001E),
-          .groups = "drop")
-
-homicide_trends$income_quartile <- as.factor(homicide_trends$income_quartile)
-# removing NAs
-homicide_trends <- homicide_trends %>% filter(!is.na(homicide_count) & !is.na(income_quartile))
-
-# creating labels
-levels(homicide_trends$income_quartile) <- c("Low Income", "Lower-Middle Income", "Higher-Middle Income", "High Income")
-# Plot homicide trends over time by income quartile
-ggplot(homicide_trends, aes(x = year, y = homicide_count)) +
-  geom_area(aes(fill = as.factor(income_quartile)), alpha = 0.6, position = "identity") +
-  labs(
-    x = "Year",
-    y = "Number of Homicides",
-    fill = "Income Quartile"
-  ) +
-  theme_minimal() +
-  scale_fill_viridis_d(option = "plasma", direction = -1)
-```
-
-```{r,  warning=FALSE, message=FALSE, fig.align = "center"}
-library(broom)
-
-weekly_counts$CrimesLag <- c(NA, weekly_counts$Crimes[-nrow(weekly_counts)])
-
-ar1_trend_regression <- lm(Crimes ~ Index + CrimesLag, data = weekly_counts, na.action = na.omit)
-
-sjPlot::tab_model(ar1_trend_regression, dv.labels = "Table 1: Linear Regression Model Results Estimating Crimes Per Week", pred.labels = c("Intercept", 'Median Income'))
-```
-
-```{r,  warning=FALSE, message=FALSE, fig.align = "center"}
-lm_income_crimes <- lm(HomicideCount ~ cali_median_income, data = homicide_income_by_year)
-
-# summary(lm_income_crimes)
-
-sjPlot::tab_model(lm_income_crimes, dv.labels = "Table 2: Linear Regression Model Results Estimating Crimes Based on Median Income", pred.labels = c("Intercept", 'Median Income'))
-```
-
-```{r, warning=FALSE, message=FALSE,  fig.align = "center"}
-library(kableExtra)
-query <- dbSendQuery(con, "SELECT * FROM crimes_by_census_block")
-crimes_by_cb <- dbFetch(query,n=-1)
-dbClearResult(query)
-
-homicides_by_cb <- crimes_by_cb %>% filter(str_detect(Crm.Cd.Desc, "CRIMINAL HOMICIDE")) %>% group_by(BLKGRPCE, year) %>% summarize(HomicideCount = n()) %>% ungroup()
-
-# shapiro.test(homicides_by_cb$HomicideCount) # cannot use parametric tests. Strong evidence against normality.
-
-mean_homicides <- mean(homicides_by_cb$HomicideCount)
-var_homicides <- var(homicides_by_cb$HomicideCount)
-# since variance > mean, use negative binomial 
-library(MASS)
-model <- glm.nb(HomicideCount ~ as.factor(BLKGRPCE), data = homicides_by_cb)
-
-pretty_labels <- c("(Intercept)", paste("Census Block", 2:6))
-sjPlot::tab_model(model, pred.labels = pretty_labels, dv.labels = "Table 3: Annual Homicide Count by Census Block", show.est = TRUE)
-```
-
-```{r,  warning=FALSE, message=FALSE, fig.align = "center"}
-homicides_latlon_rounded <- homicides %>%
-  mutate(LAT_rounded = round(LAT, 2),
-         LON_rounded = round(LON, 2))
-
-homicides_lat_long_yr <- homicides_latlon_rounded %>% 
-  group_by(year, LAT_rounded, LON_rounded) %>%
-  summarize(HomicideCount = n()) %>% ungroup()
-
-linear_model1 <- lm(HomicideCount ~ LAT_rounded + LON_rounded, data = homicides_lat_long_yr)
-```
-
-```{r,  warning=FALSE, message=FALSE, fig.align = "center"}
-
-linear_model2 <- lm(HomicideCount ~ year, data = homicides_lat_long_yr)
-
-linear_model3 <- lm(HomicideCount ~ LAT_rounded + LON_rounded + year, data = homicides_lat_long_yr)
-
-linear_model4 <- lm(HomicideCount ~ LAT_rounded * LON_rounded * year, data = homicides_lat_long_yr)
-
-sjPlot::tab_model(linear_model1, linear_model2, linear_model3, linear_model4, pred.labels = c("(Intercept)", "Latitude", "Longitude", "Year", "Latitude:Longitude", "Latitude:Year", "Longitude:Year"), dv.labels = c("Model 1", "Model 2", "Model 3", "Model 4"), show.ci = FALSE, title =
-                    "Table 4: Geospatial Linear Regression Models Predicting Homicide Count")
-
-```

### 2025-05-01 2_Slides/SlidesFinal.qmd
Reorganized repo

--- 2025-04-16 Slides/SlidesFinal.qmd
@@
 title: "Crime in Los Angeles: Homicide Increases in the COVID-19 Era"
-subtitle: "Bianca Schutz, Kevin Hu, Kaitlyn Kirt, Marco Stine"
 format:
   revealjs:
-# embed-resources: true
+    theme: solarized
     slide-number: true
@@
     preview-links: auto
-    css: styles.css
     footer: Group 5 - STAT 405 - Bianca Schutz, Kevin Hu, Kaitlyn Kirt, Marco Stine - Crime in Los Angeles
+    title-slide-attributes:
+      data-background: "#0f172a"
+      data-transition: fade
 server: shiny
@@
 
-   grid.newpage()
-   grid.rect(gp = gpar(fill = "gray15", col = NA))  # No border, black fill
-   for (star in 1:nrow(df)) {
-        row <- as.list(df[star,])
-        grid.draw(create_star(scale_factor = row$scale, center = c(row$LONstd, row$LATstd))) 
-        grid.text(label = paste0(row$LAT, ", ", row$LON),
-    x = unit(row$LONstd, "npc"),
-    y = unit(row$LATstd + 0.05, "npc"),  # offset below star; adjust as needed
-    gp = gpar(col = "white", fontsize = 10))
-        }
-   grid.rect(x = 0.5, y = 0, width = 1, height = 0.125, just = "bottom", gp = gpar(col = "white", fill = "gray15"))
-
-        grid.text(label = "Average Homicides",
-    x = unit(.125, "npc"),
-    y = unit(0.0625, "npc"),  # offset below star; adjust as needed
-    gp = gpar(col = "white", fontsize = 14))
-
-      grid.draw(create_star(scale_factor = 1, center = c(.25, .0625)))
-
-      grid.draw(create_star(scale_factor = 2, center = c(.375, .0625)))
-
-      grid.draw(create_star(scale_factor = 3, center = c(.6, .0625)))
-
-      grid.draw(create_star(scale_factor = 4, center = c(.85, .0625)))
-
-      grid.text(label = as.character(labels[1]), x = unit(.275, "npc"),
-                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11))
-      grid.text(label = as.character(labels[2]), x = unit(.475, "npc"),
-                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11))
-      grid.text(label = as.character(labels[3]), x = unit(.725, "npc"),
-                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11))
-      grid.text(label = as.character(labels[4]), x = unit(.925, "npc"),
-                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11)) 
+   draw_spatial_killer(df, labels)
 } else {
       df <- selectedKillerData()
-      print(head(df))
       rows <- nrow(df)
-      print(rows)
       col_row <- if (rows <= 13) {
@@
       stars <- generate_star_centers(col_row[1], col_row[2])
-      print(nrow(stars))
       labels <- levels(standardize_scale(df$variable))
-      print(labels)
       df_sorted <- df[order(df$year), ]
@@
       stars$year <- df_sorted$year
-      print(stars)
 
-      grid.newpage()
-      grid.rect(gp = gpar(fill = "gray15", col = NA))  # No border, black fill
-
-      for (star in 1:nrow(stars)) {
-        row <- as.list(stars[star,])
-        grid.draw(create_star(scale_factor = row$scale, center = c(row$x, row$y)))
-
-        grid.text(label = as.character(row$year),
-    x = unit(row$x, "npc"),
-    y = unit(row$y - 0.05, "npc"),  # offset below star; adjust as needed
-    gp = gpar(col = "white", fontsize = 10))
+      draw_timeline_killer(stars, labels, input$dataset)
        }
-        grid.rect(x = 0.5, y = 0, width = 1, height = 0.125, just = "bottom", gp = gpar(col = "white", fill = "gray15"))
-
-        grid.text(label = paste(stringr::str_to_title(input$dataset), "Level"),
-    x = unit(.1, "npc"),
-    y = unit(0.0625, "npc"),  # offset below star; adjust as needed
-    gp = gpar(col = "white", fontsize = 14))
-
-      grid.draw(create_star(scale_factor = 1, center = c(.225, .0625)))
-
-      grid.draw(create_star(scale_factor = 2, center = c(.375, .0625)))
-
-      grid.draw(create_star(scale_factor = 3, center = c(.6, .0625)))
-
-      grid.draw(create_star(scale_factor = 4, center = c(.85, .0625)))
-
-      grid.text(label = as.character(labels[2]), x = unit(.275, "npc"),
-                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11))
-      grid.text(label = as.character(labels[3]), x = unit(.475, "npc"),
-                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11))
-      grid.text(label = as.character(labels[4]), x = unit(.725, "npc"),
-                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11))
-      grid.text(label = as.character(labels[5]), x = unit(.925, "npc"),
-                y = unit(0.0625, "npc"), gp = gpar(col = "white", fontsize = 11)) }
     })

### 2025-05-01 2_Slides/dependencies.R

@@ -1,0 +1,155 @@
+library(grid)
+library(RSQLite)
+library(dplyr)
+library(shiny)
+
+# killer plot pipeline functions
+standardize_scale <- function(column) {
+  breaks <- quantile(column, probs = seq(0, 1, 1/3))
+  cuts <- cut(column, breaks = c(-Inf, breaks, Inf))
+  lbls <- c(
+    paste0("< ", round(breaks[1])),
+    paste0("[", round(breaks[1]), ", ", round(breaks[2]), "]"),
+    paste0("[", round(breaks[2]), ", ", round(breaks[3]), "]"),
+    paste0("[", round(breaks[3]), ", ", round(breaks[4]), "]"),
+    paste0("> ", round(breaks[4]))
+  )
+  return(factor(cuts, levels = levels(cuts), labels = lbls))
+}
+
+generate_star_centers <- function(n_cols, n_rows, radius = 0.05 / 2, padding = c(0.05, .1)) {
+  # X spacing
+  usable_width <- 1 - 2 * padding[1]
+  step_x <- (usable_width - 2 * radius) / (n_cols - 1)
+  spaceX <- step_x - 2 * radius
+  x_start <- padding[1]
+  x <- seq(x_start, by = (2 * radius + spaceX), length.out = n_cols)
+  
+  # Y spacing with top and bottom padding
+  usable_height <- 1 - 2 * padding[2] - .1
+  
+  # Handle the case where n_rows is 1 to avoid NaN (division by zero)
+  if (n_rows > 1) {
+    step_y <- (usable_height - 2 * radius) / (n_rows - 1)
+    spaceY <- step_y - 2 * radius
+  } else {
+    step_y <- usable_height
+    spaceY <- 0
+  }
+  
+  y_start <- ifelse(n_rows > 1, padding[2] + .125, .5)
+  y <- seq(y_start, by = (2 * radius + spaceY), length.out = n_rows)
+  
+  # Create grid of star centers
+  stars <- expand.grid(x = x, y = y)
+  
+  # Track rows and columns to resort by row
+  stars$row <- rep(1:n_rows, times = n_cols)
+  stars$col <- rep(1:n_cols, each = n_rows)
+  stars <- stars[order(stars$col, stars$row), ]
+  
+  return(stars)
+}
+create_star <- function(scale_factor = 1, center = c(0.06, 0.06), fill = '#a44c5c', color = "#d3af37") {
[…]

### 2025-05-01 Slides/dependencies.R

@@ -1,71 +1,0 @@
-library(grid)
-library(RSQLite)
-library(dplyr)
-library(shiny)
-
-# killer plot pipeline functions
-standardize_scale <- function(column) {
-  breaks <- quantile(column, probs = seq(0, 1, 1/3))
-  cuts <- cut(column, breaks = c(-Inf, breaks, Inf))
-  lbls <- c(
-    paste0("< ", round(breaks[1])),
-    paste0("[", round(breaks[1]), ", ", round(breaks[2]), "]"),
-    paste0("[", round(breaks[2]), ", ", round(breaks[3]), "]"),
-    paste0("[", round(breaks[3]), ", ", round(breaks[4]), "]"),
-    paste0("> ", round(breaks[4]))
-  )
-  return(factor(cuts, levels = levels(cuts), labels = lbls))
-}
-
-generate_star_centers <- function(n_cols, n_rows, radius = 0.05 / 2, padding = c(0.05, .1)) {
-  # X spacing
-  usable_width <- 1 - 2 * padding[1]
-  step_x <- (usable_width - 2 * radius) / (n_cols - 1)
-  spaceX <- step_x - 2 * radius
-  x_start <- padding[1]
-  x <- seq(x_start, by = (2 * radius + spaceX), length.out = n_cols)
-  
-  # Y spacing with top and bottom padding
-  usable_height <- 1 - 2 * padding[2] - .1
-  
-  # Handle the case where n_rows is 1 to avoid NaN (division by zero)
-  if (n_rows > 1) {
-    step_y <- (usable_height - 2 * radius) / (n_rows - 1)
-    spaceY <- step_y - 2 * radius
-  } else {
-    step_y <- usable_height
-    spaceY <- 0
-  }
-  
-  y_start <- ifelse(n_rows > 1, padding[2] + .125, .5)
-  y <- seq(y_start, by = (2 * radius + spaceY), length.out = n_rows)
-  
-  # Create grid of star centers
-  stars <- expand.grid(x = x, y = y)
-  
-  # Track rows and columns to resort by row
-  stars$row <- rep(1:n_rows, times = n_cols)
-  stars$col <- rep(1:n_cols, each = n_rows)
-  stars <- stars[order(stars$col, stars$row), ]
-  
-  return(stars)
-}
-create_star <- function(scale_factor = 1, center = c(0.06, 0.06), fill = '#a44c5c', color = "#d3af37") {
[…]
